### 2022-09-17 Draft Ped Metrics Example.Rmd
adds draft explore

---
title: "DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report"
output: html_document
fig_caption: yes
bibliography: metrics.bib
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, message = FALSE)
```

```{r loaddata, cache = TRUE}
this_dir <- '~/Documents/non-motorized/'
source(paste0(this_dir,'functions/getCleanRoadCenterLine.R'))
library(tidyverse)
library(sf)

nice_comma <- function(x){
  format(round(as.numeric(x), 0), digits = 0, nsmall=0, big.mark=",") 
}
'%ni%' <- function(x,y)!('%in%'(x,y))

groupcl <- clean_roadcl %>%  
  st_set_geometry(., NULL) %>% 
  group_by(RUCODE, road_class) %>% 
  summarise(roads = n_distinct(FULL_NAME), 
            segments = n(), 
            length = sum(LENGTH, na.rm=TRUE))

outline <- sf::read_sf(paste0(this_dir,"data/outline"))
outline$plain = 1
kitsap <- ggplot() + 
  geom_sf(data = outline["plain"]) + 
  theme_void()

cities <- sf::read_sf(paste0(this_dir,"data/cities"))
uga <- sf::read_sf(paste0(this_dir,"data/uga"))
uga$isuga = 1
lamird <- sf::read_sf(paste0(this_dir,"data/lamird"))
lamird$islamird = 1
community <- sf::read_sf(paste0(this_dir,"data/community"))

# start plotting roads
# want always to see same color for same type of road, so
#['#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf']
# NOTE: Also make this a factor so that it is sorted properly!!
road_col <- scale_color_manual(values = c("Freeway / Expressway" = "#e41a1c",
                                          "Principal Arterial" = "#377eb8",
                                          "Minor Arterial" = '#4daf4a',
                                          "Major Collector" = "#984ea3",
                                          "Minor Collector" ='#ff7f00',
                                          "Local Access" = '#a65628',
                                          "Rural Local Access (e.g. easement or other type)" = '#f781bf',
                                          "Urban Local Access (e.g. easement or other type)" = '#ffff33'
                                          ))
clean_roadcl$road_class <- factor(clean_roadcl$road_class, levels = c("Freeway / Expressway",
                                                                      "Principal Arterial",
                                                                      "Minor Arterial",
                                                                      "Major Collector",
                                                                      "Minor Collector",
                                                                      "Local Access",
                                                                      "Rural Local Access (e.g. easement or other type)",
                                                                      "Urban Local Access (e.g. easement or other type)"))

# exclude Local Access EASEMENTS & Highway/Freeway
road_df <-st_as_sf(clean_roadcl) %>%
  filter(FCLASS %ni% c(100,110,1,2))

ncdf <- road_df[!lengths(st_intersects(road_df, cities)), ]
```

## Purpose

Kitsap County Pedestrian Facilities Metrics/Evaluation Report will:

1. Provide a complete evaluation of pedestrian facilities across the county down to the segment and side of road level
2. Present an aggregated quality rating for pedestrian facilities in specific areas
3. Target missing facilities critical to connecting communities or to connecting residential to local services, recreation, and shops
4. Show measure of progress (over time) of pedestrian facilities improvements (_starting with the second report_)

## This Draft

This draft example was prepared by the Kitsap County Non-Motorized Community Advisory Committee Special Committee on Pedestrian Metrics. It is NOT considered an authoritative source of these data and is NOT an official report of the county.

## Introduction

This report provides a complete evaluation of pedestrian facilities in Kitsap County comparing the actual facilities to a minimum adequate pedestrian facility level. The minimum adequate pedestrian facility level is either a paved shoulder or a sidewalk of a certain minimum width - depending on the character of the road.

The intended audience is the people of Kitsap County and law/decision makers determining how to target limited funds for transportation development.

Not all road segments will be evaluated. This report excludes road segments that are:

 - Inside of cities (Bremerton, Port Orchard, etc.)^[Draft excludes segments that intersect with the Kitsap County incorporated city limits polygons[@kcgis_cities]]
 - Local easements^[Draft assuming Function Class 100 or 110 in Road Centerline Data]
 - Highways^[Draft Assuming Function Class 2 in Road Centerline Data]
 - Speed limit > 50 MPH^[**these are not currently identified because the road centerline data does not include speed limit**]

```{r, cache = TRUE}
library(tidyverse)
library(sf)
kitsap_all_roads <- kitsap +
  geom_sf(data = clean_roadcl, aes(color = road_class)) + 
  road_col +
  theme_void() +
  labs(title = "Roads in Kitsap County",
       subtitle = paste0(nice_comma(sum(clean_roadcl$LENGTH)/5280),
                         " centerline miles"),
       #caption = "Data source: Kitsap GIS",
       color = "")

kitsap_roads_no_cities <- kitsap + 
  geom_sf(data = cities["NAME"], 
          aes(fill="NAME"),
          fill="#808080",
          alpha = 0.5,
          show.legend = FALSE) +
  geom_sf(data = ncdf, aes(color = road_class), show.legend = FALSE) + 
  road_col +
  theme_void() +
  labs(title = "No Cities, Easements, & Highways",
       subtitle = paste0(nice_comma(sum(ncdf$LENGTH)/5280),
                         " centerline miles"),
       caption = "Data source: Kitsap GIS")

library(patchwork)

kitsap_all_roads + kitsap_roads_no_cities +
  plot_layout(guides = 'collect') & theme(legend.position = 'bottom')

#detach("patchwork", unload=TRUE)
```

## Data and Methods

### Data

Kitsap County maintains data on roads, sidewalks, and shoulders. These data all use the Kitsap County Public Works Road Log Id to identify road segments. Combined, these data allow for assessment of the quality of pedestrian facilities on Kitsap County roads. In addition, Kitsap County maintains data on the geographic representations of the cities, UGAs, LAMIRDs, and some key points of interest (like schools and parks) within the County.[@kcgis]

### Method of Evaluation

Road segments are the basis for evaluation because they are the smallest unit in common across the data and can be grouped together to form trips. Each road segment is evaluated individually. For each road segment, its adequate pedestrian facility level is based on its classification and the Kitsap County Road Standards Table 3.3 and 3.4.[@kc_road_standards]^[revisions of road standards should be followed by review and reevaluation of these criteria]

Road segments that are explicitly excluded from the evaluation will be scored NA and not included in a summary of pedestrian metrics in an area (they will neither improve nor pull down the summary score).

Pedestrian facilities on road segments will be scored based on the sum of the score of the sides. A side of the road for a segment will be scored:

- 5 if it has a sidewalk or shoulder of adequate size (based on road classification & standards)
- 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
- 1 if the sidewalk or shoulder is present and is narrower than AASHTO
- 0 if none present or no data

Therefore, segments with sidewalk or shoulder of adequate size (based on road classification) on both sides of road would receive a 10. Segments with no data or with no sidewalk or shoulder would receive a 0.

Any given area score will be a weighted average of the scores of the segments within the area. Weighting is by centerline linear feet per segment. A simple example is an area composed of 4 segments.

|Segment|Segment Score | Segment Centerline Linear Feet |
|------|:-----|:-----|
|A|6|100|
|B|0|10|
|C|10|50|
|D|8|200|
|*Total*| *?* | *360*|

The Total Score would be the segment scores multiplied by their linear feet and then divided by the total linear feet of the area (360). For this example, we get a area score of 7.5 $$ 6 x 100 + 0 x 10 + 10 x 50 + 8 x 200 = 2700$$ $$2700/360 = 7.5$$

Interpretation of Scores for Segments and for Areas

|Score| Interpretation |
|------|:-----|
|10|Good|
|5 TO 9 | Adequate^[For a segment, this means one side is good or both sides meet AASHTO minimums]|
|3 AND 4 | Minimal^[For a segment, this means one side meets AASHTO minimums]|
|Less than 3 |Poor|

## Results

### Overall

```{r}
shoulders <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SHOULDERS")
sidewalks <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SIDEWALKS")
tmp_sw_ncdf <- st_filter(sidewalks, ncdf, .predicate = st_intersects) %>% 
  st_as_sf(.) %>%
  filter(WIDTH > 0, LENG_FT > 0) %>%
  mutate(LENGTH = LENG_FT,
         TYPE = paste0("Sidewalk: ",WIDTH, " ft"),
         RD_LOG_ID = as.integer(ROADLOGID)) %>%
  select(RD_LOG_ID,SEGMENT_ID = SEG_ID,SIDE_OF_RO, WIDTH, LENGTH, TYPE)
tmp_sh_ncdf <- st_filter(shoulders, ncdf, .predicate = st_intersects) %>% 
  st_as_sf(.) %>%
  filter(SHLDR_WIDT > 0, SHLDR_SURF == "ACP", DISTANCE_F > 0) %>%
  mutate(LENGTH = DISTANCE_F,
         TYPE = paste0("Paved Shoulder: ", SHLDR_WIDT," ft"),
         RD_LOG_ID = as.integer(ROAD_LOG_I)) %>%
  select(RD_LOG_ID,SEGMENT_ID = SEG_ID,SIDE_OF_RO, WIDTH = SHLDR_WIDT, LENGTH, TYPE)
tmp_sw_sh <- bind_rows(tmp_sw_ncdf, tmp_sh_ncdf)

tmp_sw_sh %>%
  st_set_geometry(., NULL) %>%
  filter(LENGTH > 0) %>%
  group_by(RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, TYPE) %>% # road_type) %>%
  summarise(miles = sum(LENGTH, na.rm=TRUE)/5280) %>%
  ungroup() %>%
  pivot_wider(names_from = SIDE_OF_RO, values_from = "TYPE")

tmp_sw_ncdf %>%  
  st_set_geometry(., NULL) %>%
  filter(LENGTH > 0) %>%
  group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
  summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)

tmp_sh_ncdf %>%  
  st_set_geometry(., NULL) %>%
  filter(LENGTH > 0) %>%
  group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
  summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
```

```{r}
kitsap +
  geom_sf(data = tmp_sw_sh, aes(color = as.factor(TYPE))) + 
  theme_void() +
  labs(title = 'Kitsap County Sidewalk Data',
       caption = "Data source: Kitsap GIS - data in development",
       color = "Pedestrian Facilities")
```

### UGAs

### LAMIRDs

### Pedestrian Generators

#### Schools

#### Libraries

#### Transit

## Conclusion

## References

### 2022-09-17 explore/manual_sw.R

@@ -1,0 +1,118 @@
+# manually coding sidewalks
+
+# https://www.wsdot.wa.gov/research/reports/fullreports/806.2.pdf
+
+source('~/Documents/non-motorized/functions/getCleanRoadCenterLine.R')
+library(tidyverse)
+library(sf)
+library(mapboxapi)
+#mapboxapi::mb_access_token(my_mapbox, #this will not work for you - need to use a mapbox access token
+#                           install = TRUE)
+
+library(leaflet)
+
+
+# work in batches
+curr <- tmp_sw %>%
+  filter(is.na(SW_LEFT), road_type == "Collector/Arterial" ) %>% #, 
+         #MUNI_L =="KI", MUNI_R == "KI") %>%
+  group_by(RD_LOG_ID) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  arrange(desc(n)) %>%
+  top_n(10, desc(n))
+
+# now, we plot it, and then fill in the tmp_sw table
+my_pal <- leaflet::colorNumeric("viridis",  t_roads$SEGMENT_ID, na.color = "transparent")
+
+#my_map <- leaflet() %>%
+#  addMapboxTiles(style_id = "satellite-v9",
+#                 username = "mapbox") %>%
+#  setView(lng = -122.6413,
+#          lat = 47.6477,
+#          zoom = 11)
+
+working_rd <- 56100 #curr$RD_LOG_ID[5]
+
+t_roads <- clean_roadcl %>%
+  filter(RD_LOG_ID %in% working_rd) %>%
+  st_transform(crs = 4326)
+
+
+
+my_map %>%
+  addPolygons(data = t_roads, #weight = 1,
+              label = ~paste(RD_LOG_ID, SEGMENT_ID),
+              fillOpacity = .5,
+              color = ~my_pal(t_roads$SEGMENT_ID),
+              labelOptions = labelOptions(noHide = T))
+
+
+#will code if has sidewalks "FULL", "PARTIAL", "NONE"
+#working_rd <- 11300
+tmp_sw$SW_LEFT[tmp_sw$RD_LOG_ID == working_rd] <- "PARTIAL"
+tmp_sw$SW_RIGHT[tmp_sw$RD_LOG_ID == working_rd] <- "NONE"
+tmp_sw$SW_RIGHT[tmp_sw$RD_LOG_ID == working_rd & tmp_sw$SEGMENT_ID == 10823] <- "FULL"
[…]

### 2022-09-17 explore/nm_kitsap.R

@@ -21,8 +21,31 @@ cities <- sf::read_sf("./data/cities")
 #shoulders <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SHOULDERS")
 #sidewalks <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SIDEWALKS")
 
-
-
+outline <- sf::read_sf("./data/outline")
+uga <- sf::read_sf("./data/uga")
+uga$isuga = 1
+outline$plain = 1
+plot(outline["plain"]) + plot(c_outline["NAME"])
+c_outline <- outline %>%
+  st_intersection(uga) %>%
+  st_intersection(cities)
+plot(c_outline["NAME"])
+ggplot() + 
+  geom_sf(data = outline["plain"]) + 
+  geom_sf(data = uga["isuga"], 
+          aes(fill="isuga"),
+          fill="#ccf2fe",
+          show.legend = FALSE) +
+  geom_sf(data = cities["NAME"], 
+          aes(fill="NAME"),
+          fill="#58D6FE",
+          show.legend = FALSE) +
+  geom_sf(data = schools, 
+          aes(fill="NAME"),
+          fill="#0000d8",
+          show.legend = FALSE) +
+  theme_void() 
+ggsave(filename = "cty.png",  bg = "transparent")
 
 mapbox_map <- leaflet() %>%
   addMapboxTiles(style_id = "light-v9",
@@ -58,7 +81,7 @@ pub_elem_schools <- schools %>%
 walking_isos <- mb_isochrone(
   pub_elem_schools,
   profile = "walking",
-  time = 30,
+  time = 15,
   id = "NAME" #this makes the value of id the name
 )
 st_crs(walking_isos)$epsg #4326
@@ -99,7 +122,7 @@ elem_school_facilities_arterials <- some_roads %>%
 
 # now try adding the other roads
 # this was before clean up of raods DO NOT USE
-kitsap_roads <- roadcl %>%
+kitsap_roads <- clean_roadcl_w_shoulders %>% #roadcl %>%
   st_transform(crs = 4326) %>%
   st_intersection(walking_isos) %>%
   mutate(road_type = recode(HIERARCHY,
@@ -139,7 +162,7 @@ getMapForSchool <- function(i) {
   school_iso <- mb_isochrone(
     pub_elem_schools[32,],
     profile = "walking",
-    time = 30,
+    time = 15,
     id = "NAME" #this makes the value of id the name
[…]

### 2022-09-17 explore/roads_for _pedestrian_metrics.R

@@ -1,0 +1,392 @@
+
+source('~/Documents/non-motorized/functions/getCleanRoadCenterLine.R')
+library(tidyverse)
+library(sf)
+
+nice_comma <- function(x){
+  format(round(as.numeric(x), 0), digits = 0, nsmall=0, big.mark=",") 
+}
+'%ni%' <- function(x,y)!('%in%'(x,y))
+
+groupcl <- clean_roadcl %>%  
+  st_set_geometry(., NULL) %>% 
+  group_by(RUCODE, road_class) %>% 
+  summarise(roads = n_distinct(FULL_NAME), 
+            segments = n(), 
+            length = sum(LENGTH, na.rm=TRUE))
+
+outline <- sf::read_sf("./data/outline")
+outline$plain = 1
+kitsap <- ggplot() + 
+  geom_sf(data = outline["plain"]) + 
+  theme_void()
+
+cities <- sf::read_sf("./data/cities")
+uga <- sf::read_sf("./data/uga")
+uga$isuga = 1
+
+
+# start plotting roads
+# want always to see same color for same type of road, so
+#['#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf']
+# NOTE: Also make this a factor so that it is sorted properly!!
+road_col <- scale_color_manual(values = c("Freeway / Expressway" = "#e41a1c",
+                                          "Principal Arterial" = "#377eb8",
+                                          "Minor Arterial" = '#4daf4a',
+                                          "Major Collector" = "#984ea3",
+                                          "Minor Collector" ='#ff7f00',
+                                          "Local Access" = '#a65628',
+                                          "Rural Local Access (e.g. easement or other type)" = '#f781bf',
+                                          "Urban Local Access (e.g. easement or other type)" = '#ffff33'
+                                          ))
+clean_roadcl$road_class <- factor(clean_roadcl$road_class, levels = c("Freeway / Expressway",
+                                                                      "Principal Arterial",
+                                                                      "Minor Arterial",
+                                                                      "Major Collector",
[…]

### 2022-09-17 functions/getMapForAddress.R

@@ -18,35 +18,69 @@ psrc_bikeped <- sf::read_sf("./data/ElmerGeo_BikePed_042021.gdb")
 transf_psrc_bikeped <- psrc_bikeped %>%
   st_transform(crs = 4326) 
 
+#"8196 WA-104, Kingston, WA 98346"
 
 #getMapByType(this_address = "10 Washington Ave, Bremerton, WA 98337", 
 #             walkorbike = "bike")
-getMapByType <- function(this_address, walkorbike = "walk") {
+# Silverdale Library
+getMapByType(title = "Silverdale Library", this_address = "3650 NW Anderson Hill Rd, Silverdale, WA 98383", walkorbike = "walk")
+# Cougar Valley Elementary
+getMapByType(title = "Cougar Valley Elementary", this_address = "13200 Olympic View Rd NW, Silverdale, WA 98383", walkorbike = "walk")
+getMapByType(title = "Fieldstone Memory Care", 
+             this_address = "11313 Clear Creek Rd NW, Silverdale, WA 98383", 
+             walkorbike = "walk")
+#others
+getMapByType(this_address = "8196 WA-104, Kingston, WA 98346", walkorbike = "walk")
+getMapByType(this_address="100 SW Lakeway Blvd, Port Orchard WA 98366", walkorbike = "walk")
+getMapByType(this_address= "330 NE Foster Rd, Bremerton WA 98311", walkorbike = "walk")
+getMapByType(this_address= "9100 Dickey Rd NW, Silverdale, WA 98383", walkorbike = "walk")
+getMapByType(this_address = "3999 NW Sunde Rd, Silverdale, WA 98383", walkorbike = "walk")
+getMapByType <- function(title = "", this_address, walkorbike = "walk") {
   # this address must be valid, or we will get an error
   profile <- ifelse(walkorbike == "bike", "cycling", "walking")
   word <- paste0(toupper(substr(profile, 1, 1)), substr(profile, 2, nchar(profile)))
   
-  this_title <- paste("15 Minute",word, "Range<br/>",this_address)
+  this_title <- paste(title, "<br/>15 Minute",word, "Range<br/>",this_address)
   this_15 <- mb_isochrone(this_address,
                time = 15,
                profile = profile)
-  some_roads <- transf_psrc_bikeped %>%
+  meters_in_mile <- 1609 #.34
+  this_1mile <- mb_isochrone(this_address,
+                             distance = meters_in_mile,
+                             profile = profile)
+  this_geo <- mb_geocode(this_address)
[…]

### 2022-09-22 child_single_geo.Rmd
creates child document for areas

# {{current_geo}}

```{r} 
library(broom)
this_geo <- st_as_sf(df) %>% filter(tomatch == "{{current_geo}}") 
this_roads <- st_intersection(road_df, st_buffer(this_geo,0))
this_sw_sh <- st_intersection(nc_sw_sh, st_buffer(this_geo,0))

this_test <- st_intersection(select(this_roads, RD_LOG_ID, FULL_NAME), 
                             st_buffer(this_sw_sh, 30))
#this_sw_sh <- st_as_sf(this_sw_sh) %>% 
#  mutate(f = ifelse("{{current_geo}}" != "Manchester" & RD_LOG_ID == 49435, 1, 0)) %>%
 # filter(f == 0)

this_miles <- format(round(as.numeric(sum(this_roads$LENGTH)/5280), 1), 
                           digits = 1, nsmall=0, big.mark=",")

ggplot() +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
  geom_sf(data = this_test, aes(color = as.factor(TYPE))) 
  geom_sf(data = st_as_sf(this_roads) %>% filter(RD_LOG_ID == 49435), 
          aes(fill = 1, color = 1)) +
  geom_sf(data = st_as_sf(this_sw_sh) %>% filter(RD_LOG_ID == 49435), 
          aes(fill = 2, color = 2))
```

Within this area, there are `r this_miles` centerline road miles. 
```{r }
road_sum <- this_roads %>%  
  st_set_geometry(., NULL) %>% 
  group_by(road_class) %>%
  summarise(miles = format(round(as.numeric(sum(LENGTH)/5280), 1), 
                           digits = 1, nsmall=0, big.mark=","))
colnames(road_sum) <- c("Road Class", "Miles")
knitr::kable(road_sum)
```

```{r }
p1 <- kitsap +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE))) + 
  theme_void() +
  labs(title = paste("Map for","{{current_geo}}"),
       caption = "Data source: Kitsap GIS - data in development",
       color = "Pedestrian Facilities")

p2 <- ggplot() +
  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = FALSE) +
  theme_void()

library(patchwork)

p1 + p2 +
  plot_layout(guides = 'collect') #+
  #labs(title = paste("Map for","{{current_geo}}"))

detach("package:patchwork", unload=TRUE)

```

### 2022-09-22 Draft Ped Metrics Example.Rmd
updates example metrics report

@@
 
+```{css, echo = FALSE}
+tfoot {
+  font-size: 80%;
+  font-style: italic;
+}
+```
+
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, message = FALSE)
+options(scipen=999)
+
+options(knitr.kable.NA = '') # hides NA in printed table
+
+nice_comma <- function(x, dec = 0){
+  format(round(as.numeric(x), dec), digits = NULL, nsmall=dec, big.mark=",") 
+}
+'%ni%' <- function(x,y)!('%in%'(x,y))
+
+this_dir <- '~/Documents/non-motorized/'
+
 ```
@@
 ```{r loaddata, cache = TRUE}
-this_dir <- '~/Documents/non-motorized/'
+
 source(paste0(this_dir,'functions/getCleanRoadCenterLine.R'))
@@
 library(sf)
-
-nice_comma <- function(x){
-  format(round(as.numeric(x), 0), digits = 0, nsmall=0, big.mark=",") 
-}
-'%ni%' <- function(x,y)!('%in%'(x,y))
-
-groupcl <- clean_roadcl %>%  
-  st_set_geometry(., NULL) %>% 
-  group_by(RUCODE, road_class) %>% 
-  summarise(roads = n_distinct(FULL_NAME), 
-            segments = n(), 
-            length = sum(LENGTH, na.rm=TRUE))
 
@@
 road_df <-st_as_sf(clean_roadcl) %>%
-  filter(FCLASS %ni% c(100,110,1,2))
+  filter(FCLASS %ni% c(100,110,1,2), RD_LOG_ID > 0)
 
 ncdf <- road_df[!lengths(st_intersects(road_df, cities)), ]
+#cdf <- st_intersection(road_df, cities)
+#ncdf2 <- st_as_sf(road_df) %>%
+#  mutate(rdseg = paste(SEGMENT_ID, RD_LOG_ID)) %>%
+#  filter(rdseg %ni% mutate(cdf,rdseg = paste(SEGMENT_ID, RD_LOG_ID))$rdseg)
+```
+
+```{r swshdata, cache = TRUE}
+library(tidyverse)
+library(sf)
+shoulders <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SHOULDERS")
+sidewalks <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SIDEWALKS")
+
+sidewalks_nc <- sidewalks[!lengths(st_intersects(sidewalks, cities)), ]%>% 
+  st_as_sf(.) %>%
+  filter(WIDTH > 0, LENG_FT > 0) %>%
+  mutate(LENGTH = LENG_FT,
+         TYPE = paste0("Sidewalk: ",WIDTH, " ft"),
+         RD_LOG_ID = as.integer(ROADLOGID),
+         SEGMENT_ID = as.integer(SEG_ID)) %>%
+  select(RD_LOG_ID,SEGMENT_ID,BMP,EMP,SIDE_OF_RO, WIDTH, LENGTH, TYPE)
+shoulders_nc <- shoulders[!lengths(st_intersects(shoulders, cities)), ] %>% 
+  st_as_sf(.) %>%
+  filter(SHLDR_WIDT > 0, SHLDR_SURF == "ACP", DISTANCE_F > 0) %>%
+  mutate(LENGTH = DISTANCE_F,
+         TYPE = paste0("Paved Shoulder: ", SHLDR_WIDT," ft"),
+         RD_LOG_ID = as.integer(ROAD_LOG_I),
+         SEGMENT_ID = as.integer(SEG_ID)) %>%
+  select(RD_LOG_ID,SEGMENT_ID,BMP,EMP,SIDE_OF_RO, WIDTH = SHLDR_WIDT, LENGTH, TYPE)
+nc_sw_sh <- bind_rows(sidewalks_nc, shoulders_nc)
+
 ```
@@
 
-#detach("patchwork", unload=TRUE)
+detach("package:patchwork", unload=TRUE)
 ```
@@
 
-### Data
+### Data {#datasection}
 
@@
 
+These are the roads that will be evaluated.
+```{r}
+library(tidyverse)
+library(sf)
+# groupcl <- ncdf %>% #clean_roadcl %>%  
+#   st_set_geometry(., NULL) %>% 
+#   group_by(RUCODE, road_class) %>% 
+#   summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
+#             segments = nice_comma(n()), 
+#             length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+#             .groups = "drop") %>%
+#   mutate(RUCODE = factor(RUCODE, 
+#                          levels = c(0,1,2), 
+#                          labels = c("Unknown", "Rural", "Small Urban")))
+# 
+# colnames(groupcl) <- c("Rural/Urban", "Road Classification", 
+#                        "Count of Roads", "Count of Segments", 
+#                        "Length in Miles") 
+groupcl <- ncdf %>% #clean_roadcl %>%  
+  st_set_geometry(., NULL) %>% 
+  group_by(road_class) %>% 
+  summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
+            segments = nice_comma(n()), 
+            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            .groups = "drop") 
+df <- ncdf %>% #clean_roadcl %>%  
+  st_set_geometry(., NULL) %>% 
+  group_by(road_class = "Total") %>% 
+  summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
+            segments = nice_comma(n()), 
+            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            .groups = "drop") 
+groupcl <- bind_rows(groupcl, df)
+
+colnames(groupcl) <- c("Road Classification", 
+                       "Count of Roads", "Count of Segments", 
+                       "Length in Miles") 
+knitr::kable(groupcl, 
+             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")
+```
+
+These are the sidewalk and shoulders also outside of Cities.
+Single sides shown.
+In the data, there are zero segments with sidewalks on both sides of the road.
+This is one example where we know the data are incomplete; for example, Silverdale Way is not in a city and is known to have sidewalks on both sides.^[For the specific example of Silverdale Way, there are three entries in the sidewalk data with the Road Log Id of 19515 that matches Silverdale Way, and they all indicate sidewalks on the left side of the road.
+Further note, all three have a segment ID of 0, so they will not match any of the segments in the roads data if we match on Road Log Id and Segment ID - an entirely different problem with the data that is not isolated to Silverdale Way.
+`r round(sum(nc_sw_sh$SEGMENT_ID ==0)/nrow(nc_sw_sh)*100,1)`% of the sidewalk data do not have segment ids.] For this draft example report it is okay; if the initial report is completed with the incomplete data, it will present a poorer overall condition of pedestrian facilities than is in place.
+
+```{r}
+df <- sidewalks_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+df2 <- sidewalks_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+
+df <- bind_rows(df, df2)
+
+knitr::kable(df, 
+             caption = "Miles of Sidewalks by Width of Sidewalk (single side)")
+```
+
+```{r}
+df <- shoulders_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+df2 <- shoulders_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+
+df <- bind_rows(df, df2)
+
+knitr::kable(df, 
+             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)")
+```
+
+```{r}
+# tmp_sw_sh %>%
+#   st_set_geometry(., NULL) %>%
+#   group_by(RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, TYPE) %>% # road_type) %>%
+#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280) %>%
+#   ungroup() %>%
+#   pivot_wider(names_from = SIDE_OF_RO, values_from = "TYPE")
+#%>%
+  #mutate(both = !is.na(RIGHT) & !is.na(LEFT)) %>%
+  #group_by(both) %>%
+  #summarise(n = n())
+
+# tmp_sw_ncdf %>%  
+#   st_set_geometry(., NULL) %>%
+#   filter(LENGTH > 0) %>%
+#   group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
+#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
+# 
+# tmp_sh_ncdf %>%  
+#   st_set_geometry(., NULL) %>%
+#   filter(LENGTH > 0) %>%
+#   group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
+#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
+```
+
+#### Combining the Data
+
+In order to do this evaluation, we must combine the roads data (which has information on the class and setting of roads) with the sidewalk and shoulder data (with information about the pedestrian facilities).
+However, the most obvious methods of combination fail.
+
+##### Matching on Road Log Id is good but insufficient
+
+All of the data sets appear to use the same set of road log IDs from public works.
+Unfortunately, knowing that there is a shoulder or sidewalk somewhere on a road does not get to how much of that road has the sidewalk or shoulder.
+
+This diagnostic and ugly plot is intended to show the coverage of road log id across the sidewalk and shoulder data sets compared to the raod data.
+While the sidewalk and shoulder data are more sparse, the range is consistent.
+
+```{r}
+sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
+                  aes(x = as.numeric(row.names(nc_sw_sh)), y = RD_LOG_ID)) + 
+  geom_point() +
+  labs(title = "Sidewalk and Shoulder Data",
+       x = "") +
+  expand_limits(y = c(0,100000))
+road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
+                  aes(x = as.numeric(row.names(ncdf)), y = RD_LOG_ID)) + 
+  geom_point() +
+  labs(title = "Roads Data",
+       x = "")
+
+library(patchwork)
+
+sw_plot + road_plot
+
+detach("package:patchwork", unload=TRUE)
+rm(sw_plot)
+rm(road_plot)
+```
+
+```{r}
+match_road_id <- st_set_geometry(sidewalks,NULL) %>%
+  mutate(RD_LOG_ID = as.integer(ROADLOGID)) %>%
+  left_join(select(st_set_geometry(clean_roadcl,NULL),RD_LOG_ID,FULL_NAME)) %>%
+  group_by(RD_LOG_ID, ROADNAME, SIDE_OF_RO, WIDTH) %>%
+  summarise(n = sum(!is.na(FULL_NAME)), .groups = "drop") %>% 
+  filter(n == 0, WIDTH > 0)
+```
+
+The most obvious smaller match is the segment id. However, 
+**NOTE: We cannot actually use the SEGMENT ID in this draft report because it does not have the same meaning in the sidewalk and shoulder data as it has in the road data.
+The following plots are diagnostic but show that the sidewalks and shoulders data do not have segment ids covering the same spectrum as the road data.
+
+```{r}
+sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
+                  aes(x = as.numeric(row.names(nc_sw_sh)), y = SEGMENT_ID)) + 
+  geom_point() +
+  labs(title = "Sidewalk and Shoulder Data",
+       x = "")+
+  expand_limits(y = c(0,20000))
+road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
+                  aes(x = as.numeric(row.names(ncdf)), y = SEGMENT_ID)) + 
+  geom_point() +
+  labs(title = "Roads Data",
+       x = "")
+
+library(patchwork)
+
+sw_plot + road_plot
+
+detach("package:patchwork", unload=TRUE)
+rm(sw_plot)
+rm(road_plot)
+```
+
+The segment ids are clearly not from the same space, but the road log ids are consistent.
+
+```{r}
+# combine the sidewalk and shoulder and roads data
+```
+
 ### Method of Evaluation
@@
 Road segments are the basis for evaluation because they are the smallest unit in common across the data and can be grouped together to form trips.
+Actually matching up the data for roads, shoulders, and sidewalks will prove to be a more complicated task than it appears at first blush.
+As noted in the [Data section](#datasection), the Segment IDs in the shoulder and sidewalk data do not cover the range of segments as in the road data.
+Here, we join based on the road log id and beginning and ending mileage point
+
 Each road segment is evaluated individually.
 For each road segment, its adequate pedestrian facility level is based on its classification and the Kitsap County Road Standards Table 3.3 and 3.4.[@kc_road_standards]^[revisions of road standards should be followed by review and reevaluation of these criteria]
+
+```{r tbl_standards}
+# Road Standards Tables 3.3 and 3.4
+library(tidyverse)
+library(kableExtra)
+road_standards_given <- tribble(
+  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, ~note,
+  "Urban", "Local Road", NA, 5, NA,NA,
+  "Urban", "Local Sub-collector", NA, 5, NA,"sub-collector is not defined in road data",
+  "Urban", "Arterial", NA, 6, NA,NA,
+  "Urban", "Collector", NA, 6, NA,NA,
+  "Rural", "Local Road", NA, NA, 3,NA,
+  "Rural", "Local Sub-collector", NA, NA, 4,"sub-collector is not defined in road data",
+  "Rural", "Collector", "ADT 400-750", NA, 3,"ADT not defined in road data",
+  "Rural", "Collector", "ADT 751-1000", NA, 4,"ADT not defined in road data",
+  "Rural", "Collector", "DHV 100-200", NA, 6,"DHV not defined in road data",
+  "Rural", "Collector", "DHV > 200", NA, 8,"DHV not defined in road data",
+  "Rural", "Minor Arterial", "DHV < 100", NA, 4,"DHV not defined in road data",
+  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,"DHV not defined in road data",
+  "Rural", "Minor Arterial", "DHV > 200", NA, 8,"DHV not defined in road data",
+  "Rural", "Principal Arterial", "DHV <200", NA, 6,"DHV not defined in road data",
+  "Rural", "Principal Arterial", "DHV > 200", NA, 8,"DHV not defined in road data",
+)
+
+knitr::kable(road_standards_given,
+             caption = "Summary of Tables 3.3 and 3.4 in Kitsap County Road Standards [@kc_road_standards]")  %>%
+   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT")
+```
+
+Because of all of the items that are not available in the road data at this time, this draft makes some assumptions to allow for an evaluation.
+When the information is known, a more accurate evaluation can be made.
+
+```{r tbl_standards_simple}
+# Road Standards Tables 3.3 and 3.4 translated for data we have
+library(tidyverse)
+library(kableExtra)
+road_standards <- tribble(
+  ~setting, ~class, ~sidewalk, ~shoulder,
+  "Urban", "Local Access", 5, NA,
+  "Urban", "Collector",6, NA,
+  "Urban", "Arterial", 6, NA,
+  "Rural", "Local Access", NA, 3,
+  "Rural", "Minor Collector",NA, 3,
+  "Rural", "Major Collector",NA, 4,
+  "Rural", "Minor Arterial", NA, 4,
+  "Rural", "Principal Arterial", NA, 6,
+)
+
+knitr::kable(road_standards,
+             caption = "Simplified Assumptions to Allow for Evaluation]")  %>%
+   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation")
+```
 
@@
 
-### Overall
+```{r compare_standards}
+# take the road cl data for the roads under evaluation
+# remove geometry
+# merge with standards to get requirement based on setting (RUCODE) and class (road_class)
+# merge with sidewalk and shoulder data to get what exists
 
-```{r}
-shoulders <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SHOULDERS")
-sidewalks <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SIDEWALKS")
-tmp_sw_ncdf <- st_filter(sidewalks, ncdf, .predicate = st_intersects) %>% 
-  st_as_sf(.) %>%
-  filter(WIDTH > 0, LENG_FT > 0) %>%
-  mutate(LENGTH = LENG_FT,
-         TYPE = paste0("Sidewalk: ",WIDTH, " ft"),
-         RD_LOG_ID = as.integer(ROADLOGID)) %>%
-  select(RD_LOG_ID,SEGMENT_ID = SEG_ID,SIDE_OF_RO, WIDTH, LENGTH, TYPE)
-tmp_sh_ncdf <- st_filter(shoulders, ncdf, .predicate = st_intersects) %>% 
-  st_as_sf(.) %>%
-  filter(SHLDR_WIDT > 0, SHLDR_SURF == "ACP", DISTANCE_F > 0) %>%
-  mutate(LENGTH = DISTANCE_F,
-         TYPE = paste0("Paved Shoulder: ", SHLDR_WIDT," ft"),
-         RD_LOG_ID = as.integer(ROAD_LOG_I)) %>%
-  select(RD_LOG_ID,SEGMENT_ID = SEG_ID,SIDE_OF_RO, WIDTH = SHLDR_WIDT, LENGTH, TYPE)
-tmp_sw_sh <- bind_rows(tmp_sw_ncdf, tmp_sh_ncdf)
+library(tidyverse)
+library(sf)
 
-tmp_sw_sh %>%
+road_standards <- road_standards %>%
+  rename(standard_sidewalk = sidewalk, 
+         standard_shoulder = shoulder, 
+         road_class = class)
+  # pivot_longer(cols=c(sidewalk, shoulder), 
+  #              names_to = "Standard Pedestrian Facility",
+  #              values_to = "Standard Width") %>% filter
+sw_ncdf <- sidewalks_nc %>%
+  mutate(sidewalk_right = ifelse(SIDE_OF_RO == "RIGHT", WIDTH, NA),
+         sidewalk_left = ifelse(SIDE_OF_RO == "LEFT", WIDTH, NA)) %>%
+  select(RD_LOG_ID, SEGMENT_ID, sidewalk_right, sidewalk_left)
+sh_ncdf <- shoulders_nc %>%
+  mutate(shoulder_right = ifelse(SIDE_OF_RO == "RIGHT", WIDTH, NA),
+         shoulder_left = ifelse(SIDE_OF_RO == "LEFT", WIDTH, NA)) %>%
+  select(RD_LOG_ID, SEGMENT_ID, shoulder_right, shoulder_left)
+sw_sh <- full_join(st_set_geometry(sw_ncdf, NULL), 
+                   st_set_geometry(sh_ncdf, NULL)
+                   )
+
+getSideScore <- function(standard_sidewalk, standard_shoulder,
+                         sidewalk, shoulder) {
+  ifelse(is.na(standard_sidewalk), # dealing with shoulder
+         case_when(
+         shoulder >= standard_shoulder ~ 5,
+         shoulder < standard_shoulder & shoulder > 4 ~ 3,
+         shoulder < standard_shoulder & shoulder > 0 ~ 1,
+         TRUE ~ 0
+         ),
+         case_when(
+         sidewalk >= standard_sidewalk ~ 5,
+         sidewalk < standard_sidewalk & sidewalk > 4 ~ 3,
+         sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
+         TRUE ~ 0
+         )
+  )
+}
+
+ncdf_w_all_nogeo <- ncdf %>% #clean_roadcl %>%  
   st_set_geometry(., NULL) %>% 
-  filter(LENGTH > 0) %>%
-  group_by(RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, TYPE) %>% # road_type) %>%
-  summarise(miles = sum(LENGTH, na.rm=TRUE)/5280) %>%
-  ungroup() %>%
-  pivot_wider(names_from = SIDE_OF_RO, values_from = "TYPE")
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
+  left_join(road_standards) %>%
+  left_join(sw_sh) %>%
+  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_left, shoulder_left),
+         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_right, shoulder_right),
+         score = score_left + score_right)
 
-tmp_sw_ncdf %>%  
-  st_set_geometry(., NULL) %>%
-  filter(LENGTH > 0) %>%
-  group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
-  summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
+# ncdf %>% 
+#   st_set_geometry(., NULL) %>% 
+#   select(RD_LOG_ID, 
+#          BMP = CRAB_BMP, 
+#          EMP = CRAB_EMP,
+#          LENGTH) %>%
+#   left_join(select(st_set_geometry(shoulders_nc, NULL), RD_LOG_ID = ROAD_LOG_I, 
+#                    BMP, EMP,
+#                    SIDE_OF_RO, SHLDR_WIDT, SHLDR_SURF)) %>%
+#   group_by(SHLDR_SURF, SHLDR_WIDT) %>%
+#   summarise(n = n(), LENGTH = sum(LENGTH, na.rm = TRUE)/5280)
+# 
+# ncdf %>% 
+#   st_set_geometry(., NULL) %>% 
+#   select(RD_LOG_ID, FULL_NAME,
+#          BMP = CRAB_BMP, 
+#          EMP = CRAB_EMP,
+#          LENGTH) %>%
+#   left_join(select(mutate(st_set_geometry(sidewalks_nc, NULL), RD_LOG_ID = as.integer(ROADLOGID)), 
+#                    RD_LOG_ID,
+#                    BMP, EMP,
+#                    SIDE_OF_RO, WIDTH)) %>%
+#   filter(!is.na(WIDTH) & WIDTH > 0) %>%
+#   group_by(RD_LOG_ID, FULL_NAME, `Sidewalk Width` = WIDTH) %>%
+#   summarise(n = n(), LENGTH = sum(LENGTH, na.rm = TRUE)/5280)
 
-tmp_sh_ncdf %>%  
-  st_set_geometry(., NULL) %>%
-  filter(LENGTH > 0) %>%
-  group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
-  summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
+knitr::kable(
+  ncdf_w_all_nogeo %>%
+  group_by(setting, road_class, score) %>% 
+  summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
+            segments = nice_comma(n()), 
+            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            .groups = "drop") 
+  )
 ```
 
-```{r}
-kitsap +
-  geom_sf(data = tmp_sw_sh, aes(color = as.factor(TYPE))) + 
-  theme_void() +
-  labs(title = 'Kitsap County Sidewalk Data',
-       caption = "Data source: Kitsap GIS - data in development",
-       color = "Pedestrian Facilities")
-```
+### Overall
 
@@
 ### LAMIRDs
+
+```{r plot_lamirds}
+kitsap +
+  #geom_sf(data = nc_sw_sh, aes(color = as.factor(TYPE))) + 
+  geom_sf(data = lamird, aes(fill = PLAN_AREA), alpha = 0.5) +
+  theme_void() +
+  labs(title = 'Kitsap County LAMIRDS',
+       caption = "Data source: Kitsap GIS - data in development",
+       color = "LAMIRDS")
+
+```
+
+```{r loop_lamirds,include=FALSE,echo=FALSE}
+the_lamirds <- unique(st_set_geometry(lamird, NULL)$PLAN_AREA)
+df <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)
+
+expanded_child <- lapply(
+    the_lamirds,
+    function(x) {
+      knitr::knit_expand(file = "child_single_geo.Rmd",
+                                   current_geo = x )
+
+    }
+    )
+
+parsed_child <- knitr::knit_child(text = unlist(expanded_child))
+
+```
+
+`r parsed_child`
 

### 2022-10-01 Draft Ped Metrics Example.Rmd
add palette for sidewalks and shoulders data

@@
 
+# NOTE: Also make this a factor so that it is sorted properly!!
+sw_sh_col <- scale_color_manual(values = c("Sidewalk: 3 ft"="#E41A1C",
+                                           "Sidewalk: 4 ft"="#874F6F",
+                                           "Sidewalk: 5 ft"="#3881B0",
+                                           "Sidewalk: 8 ft"="#449B75",
+                                           "Paved Shoulder: 1 ft"="#56A255",
+                                           "Paved Shoulder: 2 ft"="#7E6E85",
+                                           "Paved Shoulder: 3 ft"="#AC5782",
+                                           "Paved Shoulder: 4 ft"="#E3712B", 
+                                           "Paved Shoulder: 5 ft"="#FFA10D",
+                                           "Paved Shoulder: 6 ft"="#FFE528",
+                                           "Paved Shoulder: 7 ft"="#E1C62F",
+                                           "Paved Shoulder: 8 ft"="#B16C29",
+                                           "Paved Shoulder: 10 ft"="#C66764",
+                                           "Paved Shoulder: 12 ft"="#F17EB4",
+                                           "Paved Shoulder: 15 ft"="#CB8CAD",
+                                           "Paved Shoulder: 17 ft"="#999999"))
+nc_sw_sh$TYPE <- factor(nc_sw_sh$TYPE, levels = c("Sidewalk: 3 ft",
+                                                  "Sidewalk: 4 ft",
+                                                  "Sidewalk: 5 ft",
+                                                  "Sidewalk: 8 ft" ,
+                                                  "Paved Shoulder: 1 ft",
+                                                  "Paved Shoulder: 2 ft",
+                                                  "Paved Shoulder: 3 ft",
+                                                  "Paved Shoulder: 4 ft", 
+                                                  "Paved Shoulder: 5 ft",
+                                                  "Paved Shoulder: 6 ft",
+                                                  "Paved Shoulder: 7 ft",
+                                                  "Paved Shoulder: 8 ft",
+                                                  "Paved Shoulder: 10 ft",
+                                                  "Paved Shoulder: 12 ft",
+                                                  "Paved Shoulder: 15 ft",
+                                                  "Paved Shoulder: 17 ft"))
 ```

### 2022-10-03 child_single_geo.Rmd
cleans up geometry for single geo and makes map show left and right side of road

@@
 
-this_test <- st_intersection(select(this_roads, RD_LOG_ID, FULL_NAME), 
-                             st_buffer(this_sw_sh, 30))
-#this_sw_sh <- st_as_sf(this_sw_sh) %>% 
-#  mutate(f = ifelse("{{current_geo}}" != "Manchester" & RD_LOG_ID == 49435, 1, 0)) %>%
- # filter(f == 0)
+# identified a mislabeled RD_LOG_ID in the sidewalk data for Madrone (as Main)
+x <- st_as_sf(this_sw_sh) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
+x$RD_LOG_ID <- 43809
+# x will have 0 rows if we are not dealing with Manchester
 
-this_miles <- format(round(as.numeric(sum(this_roads$LENGTH)/5280), 1), 
-                           digits = 1, nsmall=0, big.mark=",")
+# comb will have repeating segments when there are sh/sw on both sides
+this_comb <- st_intersection(select(this_roads, RD_LOG_ID_R = RD_LOG_ID, 
+                                     FULL_NAME_R = FULL_NAME, 
+                                     road_class, 
+                                     SEGMENT_ID_R = SEGMENT_ID), 
+                             st_buffer(bind_rows(this_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
+  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
+  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
+  mutate(distance = st_length(.))
 
-ggplot() +
-  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
-  geom_sf(data = this_test, aes(color = as.factor(TYPE))) 
-  geom_sf(data = st_as_sf(this_roads) %>% filter(RD_LOG_ID == 49435), 
-          aes(fill = 1, color = 1)) +
-  geom_sf(data = st_as_sf(this_sw_sh) %>% filter(RD_LOG_ID == 49435), 
-          aes(fill = 2, color = 2))
+# this_centerline_miles is CENTERLINE MILES TOTAL
+this_centerline_miles <- format(round(as.numeric(sum(st_length(this_roads))/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=",")
+# this_sw/sh_miles is RIGHT LEFT MILES with SW or SH TOTAL
+this_sw_sh_miles <- format(round(as.numeric(sum(st_length(this_sw_sh))/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=",")
+# this_sw/sh_miles is RIGHT LEFT MILES where we could match SW/SH to a road segment
+this_comb_miles <- format(round(as.numeric(sum(st_length(this_comb))/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=",")
+
+# ggplot() +
+#   geom_sf(data = this_geo, aes(fill = 1), alpha = 0.2, lwd = 0, show.legend = FALSE) + 
+#   geom_sf(data = st_buffer(bind_rows(this_sw_sh, x), 50, endCapStyle="FLAT"),
+#           fill = "#4E22EE", alpha = 0.5) +
+#   geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
+#   geom_sf(data = this_test2, color = "black") +
+#   geom_sf(data = this_sw_sh, aes(color = as.factor(RD_LOG_ID)), lwd = 0)
+
+          +
+
+r <- ggplot() +
+  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
+  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
+  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
+  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
+  sw_sh_col +
+  labs(title = "Right",
+       color = "Pedestrian Facilities") +
+  theme_void()
+
+l <- ggplot() +
+  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
+  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
+  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
+  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
+  sw_sh_col +
+  labs(title = "Left",
+       color = "Pedestrian Facilities") +
+  theme_void()
+
 ```
 
-Within this area, there are `r this_miles` centerline road miles. 
+Within this area, there are `r this_centerline_miles` centerline road miles; `r this_sw_sh_miles` miles of recorded sidewalk or shoulder (either side of road); and `r this_comb_miles` miles where we could match the recorded sidewalk or shoulder to a road segment (either side of road)^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]
+
 ```{r }
 road_sum <- this_roads %>%
+  mutate(distance = st_length(.)) %>%  
   st_set_geometry(., NULL) %>% 
   group_by(road_class) %>%
-  summarise(miles = format(round(as.numeric(sum(LENGTH)/5280), 1), 
-                           digits = 1, nsmall=0, big.mark=","))
-colnames(road_sum) <- c("Road Class", "Miles")
-knitr::kable(road_sum)
+  summarise(t_miles = sum(distance)/5280,
+            miles = format(round(as.numeric(sum(distance)/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=","),
+            .groups = "drop")
+df <- select(raod_sum, -t_miles)
+colnames(df) <- c("Road Class", "Miles")
+knitr::kable(df)
+```
+
+And we still need to get this into segment scores, but here are the miles with sidewalks or shoulders.
+
+```{r}
+df <- this_comb %>% st_set_geometry(., NULL) %>% 
+  group_by(road_class, SIDE_OF_RO, TYPE) %>%
+  summarize(miles = sum(distance)/5280,
+            .groups = "drop")
+knitr::kable(df)
 ```
@@
   geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
-  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE))) + 
+  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = "line") + 
+  sw_sh_col +
   theme_void() +
@@
 
-p1 + p2 +
-  plot_layout(guides = 'collect') #+
-  #labs(title = paste("Map for","{{current_geo}}"))
+(p1 + #l + r + #p2 +
+  plot_layout(guides = 'collect') +
+  labs(title = paste("Map for","{{current_geo}}"),
+       caption = "Data source: Kitsap GIS - data in development",
+       color = "Pedestrian Facilities") & 
+    theme(legend.position = "bottom") ) +
+  (l + theme(legend.position = "none") ) +
+  (r + theme(legend.position = "none") )
 

### 2022-10-03 Draft Ped Metrics Example.Rmd
change length in main document to st_length()

@@
 
-nice_comma <- function(x, dec = 0){
+nice_comma <- function(x, dec = 1){
   format(round(as.numeric(x), dec), digits = NULL, nsmall=dec, big.mark=",") 
@@
 road_df <-st_as_sf(clean_roadcl) %>%
-  filter(FCLASS %ni% c(100,110,1,2), RD_LOG_ID > 0)
+  filter(FCLASS %ni% c(100,110,1,2), RD_LOG_ID > 0) %>%
+  mutate(LENGTH = st_length(.))
 
-ncdf <- road_df[!lengths(st_intersects(road_df, cities)), ]
+ncdf <- road_df[!lengths(st_intersects(road_df, cities)), ] %>%
+  mutate(LENGTH = st_length(.))
+
 #cdf <- st_intersection(road_df, cities)
@@
   filter(WIDTH > 0, LENG_FT > 0) %>%
-  mutate(LENGTH = LENG_FT,
+  mutate(LENGTH = st_length(.),
          TYPE = paste0("Sidewalk: ",WIDTH, " ft"),
@@
   filter(SHLDR_WIDT > 0, SHLDR_SURF == "ACP", DISTANCE_F > 0) %>%
-  mutate(LENGTH = DISTANCE_F,
+  mutate(LENGTH = st_length(.),
          TYPE = paste0("Paved Shoulder: ", SHLDR_WIDT," ft"),
@@
   labs(title = "Roads in Kitsap County",
-       subtitle = paste0(nice_comma(sum(clean_roadcl$LENGTH)/5280),
+       subtitle = paste0(nice_comma(sum(st_length(clean_roadcl))/5280),
                          " centerline miles"),
@@
   labs(title = "No Cities, Easements, & Highways",
-       subtitle = paste0(nice_comma(sum(ncdf$LENGTH)/5280),
+       subtitle = paste0(nice_comma(sum(st_length(ncdf))/5280),
                          " centerline miles"),
@@
 All of the data sets appear to use the same set of road log IDs from public works.
-Unfortunately, knowing that there is a shoulder or sidewalk somewhere on a road does not get to how much of that road has the sidewalk or shoulder.
+Unfortunately, knowing that there is a shoulder or sidewalk somewhere on a road does not get to how much of that road has the sidewalk or shoulder.^[As roads sometimes change names and road log IDs, there are also some breaks between the road data and sidewalk/shoulder data.
+This was noted for Madrone Ave in Manchester - given the same road log ID as Main St when the road bends.]
 

### 2022-10-03 Draft Ped Metrics Example.Rmd
replace df risk with loop_df to avoid overwrite with temporary df

@@
 the_lamirds <- unique(st_set_geometry(lamird, NULL)$PLAN_AREA)
-df <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)
+loop_df <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)
 

### 2022-10-03 child_single_geo.Rmd
replace df risk with loop_df to avoid overwrite with temporary df

@@
 library(broom)
-this_geo <- st_as_sf(df) %>% filter(tomatch == "{{current_geo}}") 
+this_geo <- st_as_sf(loop_df) %>% filter(tomatch == "{{current_geo}}") 
 this_roads <- st_intersection(road_df, st_buffer(this_geo,0))

### 2022-10-03 Draft Ped Metrics Example.Rmd
converts st_length() to integer to avoid class of units

@@
   filter(FCLASS %ni% c(100,110,1,2), RD_LOG_ID > 0) %>%
-  mutate(LENGTH = st_length(.))
+  mutate(LENGTH = as.integer(st_length(.)))
 
 ncdf <- road_df[!lengths(st_intersects(road_df, cities)), ] %>%
-  mutate(LENGTH = st_length(.))
+  mutate(LENGTH = as.integer(st_length(.)))
 
@@
   filter(WIDTH > 0, LENG_FT > 0) %>%
-  mutate(LENGTH = st_length(.),
+  mutate(LENGTH = as.integer(st_length(.)),
          TYPE = paste0("Sidewalk: ",WIDTH, " ft"),
@@
   filter(SHLDR_WIDT > 0, SHLDR_SURF == "ACP", DISTANCE_F > 0) %>%
-  mutate(LENGTH = st_length(.),
+  mutate(LENGTH = as.integer(st_length(.)),
          TYPE = paste0("Paved Shoulder: ", SHLDR_WIDT," ft"),

### 2022-10-03 child_single_geo.Rmd
remove extraneous + and move graphing to last chunk
converts st_length() to integer to avoid class of units

--- 2022-10-03 child_single_geo.Rmd
@@
 library(broom)
-this_geo <- st_as_sf(df) %>% filter(tomatch == "{{current_geo}}") 
+this_geo <- st_as_sf(loop_df) %>% filter(tomatch == current_geo) #"{{current_geo}}") 
 this_roads <- st_intersection(road_df, st_buffer(this_geo,0))
@@
 # identified a mislabeled RD_LOG_ID in the sidewalk data for Madrone (as Main)
-x <- st_as_sf(this_sw_sh) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
+x <- st_as_sf(this_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
 x$RD_LOG_ID <- 43809
@@
 #   geom_sf(data = this_sw_sh, aes(color = as.factor(RD_LOG_ID)), lwd = 0)
-
-          +
-
-r <- ggplot() +
-  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
-  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
-  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
-  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
-  sw_sh_col +
-  labs(title = "Right",
-       color = "Pedestrian Facilities") +
-  theme_void()
-
-l <- ggplot() +
-  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
-  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
-  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
-  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
-  sw_sh_col +
-  labs(title = "Left",
-       color = "Pedestrian Facilities") +
-  theme_void()
 
@@
 
+r <- ggplot() +
+  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
+  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
+  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
+  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
+  sw_sh_col +
+  labs(title = "Right",
+       color = "Pedestrian Facilities") +
+  theme_void()
+
+l <- ggplot() +
+  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.4, lwd = 0, show.legend = FALSE) +
+  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
+  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
+  geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
+  sw_sh_col +
+  labs(title = "Left",
+       color = "Pedestrian Facilities") +
+  theme_void()
+
 p2 <- ggplot() +

### 2022-10-03 child_single_geo.Rmd
correct placeholders

@@
 library(broom)
-this_geo <- st_as_sf(loop_df) %>% filter(tomatch == current_geo) #"{{current_geo}}") 
+this_geo <- st_as_sf(loop_df) %>% filter(tomatch == "{{current_geo}}") 
 this_roads <- st_intersection(road_df, st_buffer(this_geo,0))

### 2022-10-03 Draft Ped Metrics Example.Rmd
adapt for UGAs and remove cities from UGAs

@@
 
+We can also look at regions within the County.
+For these, we identify the roads and any sidewalks or shoulders.^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]
+
 ### UGAs
+
+```{r plot_ugas}
+uga_no_city <- st_as_sf(uga) %>%
+  filter(substring(GMA_JURISD, 1, 4) != "City")
+kitsap +
+  geom_sf(data = uga_no_city, aes(fill = GMA_JURISD), alpha = 0.5) +
+  theme_void() +
+  labs(title = 'Kitsap County UGAs',
+       caption = "Data source: Kitsap GIS - data in development",
+       color = "UGAS")
+
+```
+
+```{r loop_ugas,include=FALSE,echo=FALSE}
+the_ugas <- unique(st_set_geometry(uga_no_city, NULL)$GMA_JURISD)
+loop_df <- st_as_sf(uga_no_city) %>% mutate(tomatch = GMA_JURISD)
+
+expanded_child <- lapply(
+    the_ugas,
+    function(x) {
+      knitr::knit_expand(file = "child_single_geo.Rmd",
+                                   current_geo = x )
+
+    }
+    )
+
+parsed_child <- knitr::knit_child(text = unlist(expanded_child))
+
+```
+
+Kitsap County has `r length(the_ugas)` UGAs - Urban Growth Areas that are outside of cities.
+Within the Urban Growth Areas, most design and planning standards are similar to what would be expected in an urban environment for new development.
+
+`r parsed_child`
 
@@
 ```
+Kitsap County has `r length(the_lamirds)` LAMIRDS - Local Area of More Intense Rural Development.
 

### 2022-10-03 child_single_geo.Rmd
adapt for UGAs and remove cities from UGAs

--- 2022-10-03 child_single_geo.Rmd
@@
 ```{r} 
-library(broom)
+#library(broom)
+
 this_geo <- st_as_sf(loop_df) %>% filter(tomatch == "{{current_geo}}") 
@@
 
+# need to handle the case where this_sw_sh or this_roads has 0 rows
+has_roads <- nrow(this_roads) != 0
+has_sw_sh <- nrow(this_sw_sh) != 0
+has_both <- has_roads & has_sw_sh
+# do not evaluate most of this child if does not
+```
+
+```{r, eval=!has_both}
+cat("This area does not have either known roads or known sidewalks or shoulders")
+```
+
+```{r, eval=has_sw_sh & !has_roads}
+this_centerline_miles <- 0 # this_centerline_miles is CENTERLINE MILES TOTAL
+# this_sw/sh_miles is RIGHT LEFT MILES with SW or SH TOTAL
+this_sw_sh_miles <- format(round(as.numeric(sum(st_length(this_sw_sh))/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=",")
+# this_comb_miles is RIGHT LEFT MILES where we could match SW/SH to a road segment
+this_comb_miles <- 0
+```
+```{r, eval=!has_sw_sh & has_roads}
+# this_centerline_miles is CENTERLINE MILES TOTAL
+this_centerline_miles <- format(round(as.numeric(sum(st_length(this_roads))/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=",")
+# this_sw/sh_miles is RIGHT LEFT MILES with SW or SH TOTAL
+this_sw_sh_miles <- 0
+# this_sw/sh_miles is RIGHT LEFT MILES where we could match SW/SH to a road segment
+this_comb_miles <- 0
+```
+
+```{r, eval=has_both}
 # identified a mislabeled RD_LOG_ID in the sidewalk data for Madrone (as Main)
-x <- st_as_sf(this_sw_sh) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
+x <- st_as_sf(this_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
 x$RD_LOG_ID <- 43809
@@
 
-          +
+```
+
+Within this area, there are `r this_centerline_miles` centerline road miles; `r this_sw_sh_miles` miles of recorded sidewalk or shoulder (either side of road); and `r this_comb_miles` miles where we could match the recorded sidewalk or shoulder to a road segment (either side of road).
+
+```{r, eval=has_both}
+
+road_sum <- this_roads %>%
+  mutate(distance = st_length(.)) %>%  
+  st_set_geometry(., NULL) %>% 
+  group_by(road_class) %>%
+  summarise(t_miles = sum(distance)/5280,
+            miles = format(round(as.numeric(sum(distance)/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=","),
+            .groups = "drop")
+df <- select(road_sum, -t_miles)
+colnames(df) <- c("Road Class", "Miles")
+knitr::kable(df)
+```
+
+And we still need to get this into segment scores, but here are the miles with sidewalks or shoulders.
+
+```{r, eval=has_both}
+df <- this_comb %>% st_set_geometry(., NULL) %>% 
+  group_by(road_class, SIDE_OF_RO, TYPE) %>%
+  summarize(s_miles = sum(distance)/5280,
+            miles = format(round(as.numeric(sum(distance)/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=","),
+            .groups = "drop")
+knitr::kable(select(df,-s_miles))
+```
+
+```{r, eval=has_both}
+p1 <- kitsap +
+  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
+  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
+  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = "line") + 
+  sw_sh_col +
+  theme_void() +
+  labs(title = paste("Map for","{{current_geo}}"),
+       caption = "Data source: Kitsap GIS - data in development",
+       color = "Pedestrian Facilities")
 
@@
 
-```
-
-Within this area, there are `r this_centerline_miles` centerline road miles; `r this_sw_sh_miles` miles of recorded sidewalk or shoulder (either side of road); and `r this_comb_miles` miles where we could match the recorded sidewalk or shoulder to a road segment (either side of road)^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]
-
-```{r }
-road_sum <- this_roads %>%
-  mutate(distance = st_length(.)) %>%  
-  st_set_geometry(., NULL) %>% 
-  group_by(road_class) %>%
-  summarise(t_miles = sum(distance)/5280,
-            miles = format(round(as.numeric(sum(distance)/5280), 1), 
-                           digits = 1, nsmall=1, big.mark=","),
-            .groups = "drop")
-df <- select(raod_sum, -t_miles)
-colnames(df) <- c("Road Class", "Miles")
-knitr::kable(df)
-```
-
-And we still need to get this into segment scores, but here are the miles with sidewalks or shoulders.
-
-```{r}
-df <- this_comb %>% st_set_geometry(., NULL) %>% 
-  group_by(road_class, SIDE_OF_RO, TYPE) %>%
-  summarize(miles = sum(distance)/5280,
-            .groups = "drop")
-knitr::kable(df)
-```
-
-```{r }
-p1 <- kitsap +
-  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
-  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
-  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = "line") + 
-  sw_sh_col +
-  theme_void() +
-  labs(title = paste("Map for","{{current_geo}}"),
-       caption = "Data source: Kitsap GIS - data in development",
-       color = "Pedestrian Facilities")
-
 p2 <- ggplot() +

### 2022-10-03 child_single_geo.Rmd
adds next steps
fixes plain text conditional text

@@
 
-```{r, eval=!has_both}
-cat("This area does not have either known roads or known sidewalks or shoulders")
+```{r, results = 'asis', eval=!has_roads}
+cat("This area does not have either known roads.")
+```
+
+```{r, results = 'asis', eval=!has_sw_sh}
+cat("This area does not have known sidewalks or shoulders.")
 ```
@@
 ```
+
+```{r}
+# next step
+# take the comb data, get the distance for each segment
+# combine that with this roads
+# where no sw sh = 0
+# determine if meets criteria
+
+```

### 2022-10-04 Draft Ped Metrics Example.Rmd
calculates scores properly for whole county

@@
 
+# need to make sure that this is right
 ncdf <- road_df[!lengths(st_intersects(road_df, cities)), ] %>%
@@
   group_by(road_class) %>% 
-  summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
-            segments = nice_comma(n()), 
+  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
+            segments = nice_comma(n(),0), 
             length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
@@
   group_by(road_class = "Total") %>% 
-  summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
-            segments = nice_comma(n()), 
+  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
+            segments = nice_comma(n(),0), 
             length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
@@
 Single sides shown.
-In the data, there are zero segments with sidewalks on both sides of the road.
 This is one example where we know the data are incomplete; for example, Silverdale Way is not in a city and is known to have sidewalks on both sides.^[For the specific example of Silverdale Way, there are three entries in the sidewalk data with the Road Log Id of 19515 that matches Silverdale Way, and they all indicate sidewalks on the left side of the road.
@@
 
-The segment ids are clearly not from the same space, but the road log ids are consistent.
+The segment ids are clearly not from the same space, but the road log ids are consistent.^[Mostly.] However, we cannot combine on road log id alone - because the roads change character and the type of pedestrian facilities over space.
+We combine instead on geometry.
 
-```{r}
+For now, we are making a space around a sidewalk or shoulder geometry, a buffer, to capture the nearby area where the road would be, and seeing if there are any road segments that intersect.
+If the road segment intersects and has the same road log id, we keep it.
+
+```{r combine-rd-sw-sh, echo = TRUE}
 # combine the sidewalk and shoulder and roads data
+comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
+                                     SEGMENT_ID_R = SEGMENT_ID,
+                                     FULL_NAME_R = FULL_NAME, 
+                                     LENGTH_R = LENGTH,
+                                     road_class, RUCODE), 
+                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
+  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
+  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
+  mutate(distance = as.integer(st_length(.)))
+#distance is the length of the matched road segment and sidewalk or shoulder 
+comb %>% st_set_geometry(., NULL) %>% 
+  group_by(road_class, SIDE_OF_RO, TYPE) %>%
+  summarize(s_miles = sum(distance)/5280,
+            miles = format(round(as.numeric(sum(distance)/5280), 1), 
+                           digits = 1, nsmall=1, big.mark=","),
+            .groups = "drop")
 ```
@@
   "Urban", "Local Access", 5, NA,
-  "Urban", "Collector",6, NA,
-  "Urban", "Arterial", 6, NA,
+  "Urban", "Minor Collector",6, NA,
+  "Urban", "Major Collector",6, NA,
+  "Urban", "Minor Arterial", 6, NA,
+  "Urban", "Principal Arterial", 6, NA,
   "Rural", "Local Access", NA, 3,
@@
 
+comb_stand <- comb %>% 
+  st_set_geometry(., NULL) %>% 
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown")),
+         ped_type = sub("\\:.*", "", TYPE),
+         ped_w = as.numeric(sub(".*?([0-9]+).*", "\\1", TYPE)),
+         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
+  #select(RD_LOG_ID = RD_LOG_ID_R, SEGMENT_ID = SEGMENT_ID_R, FULL_NAME = FULL_NAME_R,
+  #       SEGMENT_LENGTH = LENGTH_R, SW_SH_LENGTH = LENGTH, SIDE_OF_RO,
+  #       distance, road_class, setting, nn, ped_w) %>%
+  pivot_wider(id_cols = -SIDE_OF_RO, #c(RD_LOG_ID, SEGMENT_ID, FULL_NAME), 
+              names_from = nn, 
+              values_from = ped_w) %>%
+  left_join(road_standards)
+
+comb_score <- comb_stand %>%
+  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_left, shoulder_left),
+         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_right, shoulder_right),
+         score = score_left + score_right)
+
 ncdf_w_all_nogeo <- ncdf %>% #clean_roadcl %>%  
@@
 knitr::kable(
-  ncdf_w_all_nogeo %>%
+ # ncdf_w_all_nogeo %>%
+  comb_score %>%
   group_by(setting, road_class, score) %>% 
-  summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
+  summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
             segments = nice_comma(n()), 
             length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
             .groups = "drop") 

### 2022-10-04 Draft Ped Metrics Example.Rmd
combined scores

@@
          ped_type = sub("\\:.*", "", TYPE),
-         ped_w = as.numeric(sub(".*?([0-9]+).*", "\\1", TYPE)),
+         #ped_w = as.numeric(sub(".*?([0-9]+).*", "\\1", TYPE)),
          nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
@@
   #       distance, road_class, setting, nn, ped_w) %>%
-  pivot_wider(id_cols = -SIDE_OF_RO, #c(RD_LOG_ID, SEGMENT_ID, FULL_NAME), 
+  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), #c(RD_LOG_ID, SEGMENT_ID, FULL_NAME), 
               names_from = nn, 
-              values_from = ped_w) %>%
+              values_from = WIDTH,
+              values_fn = mean) %>%
   left_join(road_standards)
@@
 
-ncdf_w_all_nogeo <- ncdf %>% #clean_roadcl %>%  
-  st_set_geometry(., NULL) %>% 
-  mutate(setting = ifelse(RUCODE == 1, "Rural", 
-                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
-  left_join(road_standards) %>%
-  left_join(sw_sh) %>%
-  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
-                                   sidewalk_left, shoulder_left),
-         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
-                                   sidewalk_right, shoulder_right),
-         score = score_left + score_right)
-
-# ncdf %>% 
-#   st_set_geometry(., NULL) %>% 
-#   select(RD_LOG_ID, 
-#          BMP = CRAB_BMP, 
-#          EMP = CRAB_EMP,
-#          LENGTH) %>%
-#   left_join(select(st_set_geometry(shoulders_nc, NULL), RD_LOG_ID = ROAD_LOG_I, 
-#                    BMP, EMP,
-#                    SIDE_OF_RO, SHLDR_WIDT, SHLDR_SURF)) %>%
-#   group_by(SHLDR_SURF, SHLDR_WIDT) %>%
-#   summarise(n = n(), LENGTH = sum(LENGTH, na.rm = TRUE)/5280)
-# 
-# ncdf %>% 
-#   st_set_geometry(., NULL) %>% 
-#   select(RD_LOG_ID, FULL_NAME,
-#          BMP = CRAB_BMP, 
-#          EMP = CRAB_EMP,
-#          LENGTH) %>%
-#   left_join(select(mutate(st_set_geometry(sidewalks_nc, NULL), RD_LOG_ID = as.integer(ROADLOGID)), 
-#                    RD_LOG_ID,
-#                    BMP, EMP,
-#                    SIDE_OF_RO, WIDTH)) %>%
-#   filter(!is.na(WIDTH) & WIDTH > 0) %>%
-#   group_by(RD_LOG_ID, FULL_NAME, `Sidewalk Width` = WIDTH) %>%
-#   summarise(n = n(), LENGTH = sum(LENGTH, na.rm = TRUE)/5280)
-
 knitr::kable(
@@
             segments = nice_comma(n()), 
-            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            #length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),

### 2022-10-05 Draft Ped Metrics Example.Rmd
changes the main breakdowns to non city county roads

@@
 cities <- sf::read_sf(paste0(this_dir,"data/cities"))
+
+# try to make a kitsap without cities
+county_no_cities <- st_difference(outline, st_union(cities))
+
 uga <- sf::read_sf(paste0(this_dir,"data/uga"))
@@
 # need to make sure that this is right
-ncdf <- road_df[!lengths(st_intersects(road_df, cities)), ] %>%
+
+ncdf <- st_intersection(road_df, st_buffer(county_no_cities,0)) %>%
   mutate(LENGTH = as.integer(st_length(.)))
 
-#cdf <- st_intersection(road_df, cities)
-#ncdf2 <- st_as_sf(road_df) %>%
-#  mutate(rdseg = paste(SEGMENT_ID, RD_LOG_ID)) %>%
-#  filter(rdseg %ni% mutate(cdf,rdseg = paste(SEGMENT_ID, RD_LOG_ID))$rdseg)
 ```
@@
 
-sidewalks_nc <- sidewalks[!lengths(st_intersects(sidewalks, cities)), ]%>% 
+sidewalks_nc <- st_intersection(sidewalks, st_buffer(county_no_cities,0)) %>%
   st_as_sf(.) %>%
@@
   select(RD_LOG_ID,SEGMENT_ID,BMP,EMP,SIDE_OF_RO, WIDTH, LENGTH, TYPE)
-shoulders_nc <- shoulders[!lengths(st_intersects(shoulders, cities)), ] %>% 
+shoulders_nc <- st_intersection(shoulders, st_buffer(county_no_cities,0)) %>%
   st_as_sf(.) %>%
@@
 knitr::kable(
- # ncdf_w_all_nogeo %>%
   comb_score %>%
@@
             segments = nice_comma(n()), 
-            #length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),

### 2022-10-06 Draft Ped Metrics Example.Rmd
add weighted scores for summary

@@
 # combine the sidewalk and shoulder and roads data
+x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
+x$RD_LOG_ID <- 43809
 comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
@@
 knitr::kable(road_standards,
-             caption = "Simplified Assumptions to Allow for Evaluation]")  %>%
+             caption = "Simplified Assumptions to Allow for Evaluation")  %>%
    footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation")
@@
 
-Interpretation of Scores for Segments and for Areas
+```{r}
+score_interpretation <- tribble(
+  ~score_at_least, ~interpretation,
+  10, "Good",
+  5, "Adequate",
+  3, "Minimum",
+  0, "Poor",
+)
 
-|Score| Interpretation |
-|------|:-----|
-|10|Good|
-|5 TO 9 | Adequate^[For a segment, this means one side is good or both sides meet AASHTO minimums]|
-|3 AND 4 | Minimal^[For a segment, this means one side meets AASHTO minimums]|
-|Less than 3 |Poor|
+getScoreInterpretation <- function(x){
+  first(score_interpretation$interpretation[
+        score_interpretation$score_at_least <= x]
+    )
+}
+
+knitr::kable(score_interpretation,
+             caption = "Interpretation of Scores for Segments and for Areas")  %>%
+   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums")
+```
 
 ## Results
+First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.
+
+### Overall
 
@@
          road_class = class)
-  # pivot_longer(cols=c(sidewalk, shoulder), 
-  #              names_to = "Standard Pedestrian Facility",
-  #              values_to = "Standard Width") %>% filter
-sw_ncdf <- sidewalks_nc %>%
-  mutate(sidewalk_right = ifelse(SIDE_OF_RO == "RIGHT", WIDTH, NA),
-         sidewalk_left = ifelse(SIDE_OF_RO == "LEFT", WIDTH, NA)) %>%
-  select(RD_LOG_ID, SEGMENT_ID, sidewalk_right, sidewalk_left)
-sh_ncdf <- shoulders_nc %>%
-  mutate(shoulder_right = ifelse(SIDE_OF_RO == "RIGHT", WIDTH, NA),
-         shoulder_left = ifelse(SIDE_OF_RO == "LEFT", WIDTH, NA)) %>%
-  select(RD_LOG_ID, SEGMENT_ID, shoulder_right, shoulder_left)
-sw_sh <- full_join(st_set_geometry(sw_ncdf, NULL), 
-                   st_set_geometry(sh_ncdf, NULL)
-                   )
 
@@
          ped_type = sub("\\:.*", "", TYPE),
-         #ped_w = as.numeric(sub(".*?([0-9]+).*", "\\1", TYPE)),
          nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
-  #select(RD_LOG_ID = RD_LOG_ID_R, SEGMENT_ID = SEGMENT_ID_R, FULL_NAME = FULL_NAME_R,
-  #       SEGMENT_LENGTH = LENGTH_R, SW_SH_LENGTH = LENGTH, SIDE_OF_RO,
-  #       distance, road_class, setting, nn, ped_w) %>%
-  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), #c(RD_LOG_ID, SEGMENT_ID, FULL_NAME), 
+  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), 
               names_from = nn, 
@@
 
+```
+
+```{r}
+comb_score2 <- st_set_geometry(ncdf, NULL) %>% 
+  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
+         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
+  left_join(comb_score) %>%
+  mutate(the_score = ifelse(is.na(score), 0, score),
+         the_length = ifelse(is.na(distance), LENGTH, distance),
+         w_score = the_score*the_length)
+
+comb_total_score <- sum(comb_score2$w_score)/sum(comb_score2$the_length)
+
+comb_score_by_setting <- comb_score2 %>%
+  group_by(setting) %>%
+  summarise(score = sum(w_score)/sum(the_length),
+            .groups = "drop") %>%
+  rowwise() %>%
+  mutate(interpretation = getScoreInterpretation(score))
+
+comb_score_by_setting_class <- comb_score2 %>%
+  group_by(setting, road_class) %>%
+  summarise(score = sum(w_score)/sum(the_length),
+            .groups = "drop") %>%
+  rowwise() %>%
+  mutate(interpretation = getScoreInterpretation(score))
+```
+
+The overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, `r getScoreInterpretation(comb_total_score)`.
+
+```{r}
+knitr::kable(comb_score_by_setting_class)
+```
+
+More detailed information
+```{r}
 knitr::kable(
-  comb_score %>%
+  comb_score2 %>%
   group_by(setting, road_class, score) %>% 
@@
             segments = nice_comma(n()), 
+            #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
@@
 
-### Overall
+```{r}
+ggplot(filter(comb_score2, setting != "Unknown"), 
+       aes(x = as.factor(score), y = the_length/5280))+
+  geom_col() +
+  labs(x = "score based on data available", 
+       y = "centerline miles of road") +
+  theme_minimal() +
+  facet_grid(setting ~ road_class)
+```
+
+If we remove the segments where we do not have information about sidewalks and shoulders at all, we can see that for those, we are generally finding higher scores for rural roads - with paved shoulders.
+
+```{r}
+ggplot(filter(comb_score2, setting != "Unknown", !is.na(score)), 
+       aes(x = as.factor(score), y = LENGTH/5280))+
+  geom_col() +
+  labs(x = "score based on data available", 
+       y = "centerline miles of road") +
+  theme_minimal() +
+  facet_grid(setting ~ road_class)
+```
+
+Some of the segments with score of 0 are because they have shoulders when they should have sidewalks.
+Here's a few example segments.
+If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.
+
+```{r}
+
+roads_0 <- comb_score2 %>%
+  filter(score == 0) %>%
+  select(setting, road_class, FULL_NAME_R, sidewalk_right:score)
+
+knitr::kable(head(roads_0))
+
+```
+
+```{r}
+roads_10 <- comb_score2 %>%
+  filter(score == 10) %>%
+  group_by(setting, road_class, "Road Name" = FULL_NAME_R) %>%
+  summarise(segments = nice_comma(n()), 
+            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
+            .groups = "drop")
+
+knitr::kable(roads_10)
+```
 

### 2022-10-06 child_single_geo.Rmd
add weighted scores for summary

@@
 
+```{r eval=has_both}
+this_comb_stand <- this_comb %>% 
+  st_set_geometry(., NULL) %>% 
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown")),
+         ped_type = sub("\\:.*", "", TYPE),
+         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
+  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), 
+              names_from = nn, 
+              values_from = WIDTH,
+              values_fn = mean) %>%
+  left_join(road_standards)
+
+this_comb_score <- this_comb_stand %>%
+  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_left, shoulder_left),
+         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_right, shoulder_right),
+         score = score_left + score_right)
+this_comb_score2 <- st_set_geometry(this_roads, NULL) %>% 
+  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
+         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
+  left_join(this_comb_score) %>%
+  mutate(the_score = ifelse(is.na(score), 0, score),
+         the_length = ifelse(is.na(distance), LENGTH, distance),
+         w_score = the_score*the_length)
+
+this_comb_total_score <- sum(this_comb_score2$w_score)/sum(this_comb_score2$the_length)
+```
+
 ```{r}

### 2022-10-06 Draft Ped Metrics Example.Rmd
defend against missing columns

@@
   mutate(distance = as.integer(st_length(.)))
+
 #distance is the length of the matched road segment and sidewalk or shoulder 
-comb %>% st_set_geometry(., NULL) %>% 
-  group_by(road_class, SIDE_OF_RO, TYPE) %>%
-  summarize(s_miles = sum(distance)/5280,
-            miles = format(round(as.numeric(sum(distance)/5280), 1), 
-                           digits = 1, nsmall=1, big.mark=","),
-            .groups = "drop")
+#comb %>% st_set_geometry(., NULL) %>% 
+#  group_by(road_class, SIDE_OF_RO, TYPE) %>%
+#  summarize(s_miles = sum(distance)/5280,
+#            miles = format(round(as.numeric(sum(distance)/5280), 1), 
+#                           digits = 1, nsmall=1, big.mark=","),
+#            .groups = "drop")
 ```
@@
 
+For pedestrian generators, we identify an area that is within 1 mile of the facility and evaluate roads therein in the same manner as the defined county areas, like UGAs and LAMIRDS.
+
 #### Schools
@@
 #### Libraries
+
+```{r}
+libraries <- community %>%
+  filter(TYPE == "PUBLIC LIBRARY")
+```
+
+Kitsap County has `r nrow(libraries)`
 

### 2022-10-06 child_single_geo.Rmd
defend against missing columns

@@
 this_comb <- st_intersection(select(this_roads, RD_LOG_ID_R = RD_LOG_ID, 
+                                     SEGMENT_ID_R = SEGMENT_ID,
                                      FULL_NAME_R = FULL_NAME, 
-                                     road_class, 
-                                     SEGMENT_ID_R = SEGMENT_ID), 
+                                     LENGTH_R = LENGTH,
+                                     road_class, RUCODE), 
                              st_buffer(bind_rows(this_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
@@
               values_fn = mean) %>%
-  left_join(road_standards)
+  left_join(road_standards) %>%
+  rowwise() %>% # now make sure that have all the columns
+  mutate(sidewalk_left = ifelse("sidewalk_left" %in% names(.), sidewalk_left, NA),
+         sidewalk_right = ifelse("sidewalk_right" %in% names(.), sidewalk_right, NA),
+         shoulder_left = ifelse("shoulder_left" %in% names(.), shoulder_left, NA),
+         shoulder_right = ifelse("shoulder_right" %in% names(.), shoulder_right, NA)
+         )
 
@@
 
+The overall score for pedestrian facilities in {{current_geo}} is `r nice_comma(this_comb_total_score)`, `r getScoreInterpretation(this_comb_total_score)`.
+
 ```{r}

### 2022-10-06 child_single_geo.Rmd
try kable material styling

@@
 colnames(df) <- c("Road Class", "Miles")
-knitr::kable(df)
+knitr::kable(df) %>%
+  kable_material(c("striped", "hover"))
 ```
@@
             .groups = "drop")
-knitr::kable(select(df,-s_miles))
+knitr::kable(select(df,-s_miles)) %>%
+  kable_material(c("striped", "hover"))
 ```
@@
 
-The overall score for pedestrian facilities in {{current_geo}} is `r nice_comma(this_comb_total_score)`, `r getScoreInterpretation(this_comb_total_score)`.
+**The overall score for pedestrian facilities in {{current_geo}} is `r nice_comma(this_comb_total_score)`, `r getScoreInterpretation(this_comb_total_score)`.**
 

### 2022-10-06 Draft Ped Metrics Example.Rmd
clean up tables

@@
 this_dir <- '~/Documents/non-motorized/'
+
+library(kableExtra)
 
@@
 library(sf)
-# groupcl <- ncdf %>% #clean_roadcl %>%  
-#   st_set_geometry(., NULL) %>% 
-#   group_by(RUCODE, road_class) %>% 
-#   summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
-#             segments = nice_comma(n()), 
-#             length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-#             .groups = "drop") %>%
-#   mutate(RUCODE = factor(RUCODE, 
-#                          levels = c(0,1,2), 
-#                          labels = c("Unknown", "Rural", "Small Urban")))
-# 
-# colnames(groupcl) <- c("Rural/Urban", "Road Classification", 
-#                        "Count of Roads", "Count of Segments", 
-#                        "Length in Miles") 
-groupcl <- ncdf %>% #clean_roadcl %>%  
+
+groupcl <- ncdf %>%  
   st_set_geometry(., NULL) %>% 
@@
 knitr::kable(groupcl, 
-             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")
+             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")  %>%
+  kable_material(c("striped", "hover"))
 ```
@@
 knitr::kable(df, 
-             caption = "Miles of Sidewalks by Width of Sidewalk (single side)")
+             caption = "Miles of Sidewalks by Width of Sidewalk (single side)") %>%
+  kable_material(c("striped", "hover"))
 ```
@@
 knitr::kable(df, 
-             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)")
+             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)") %>%
+  kable_material(c("striped", "hover"))
 ```
@@
 
-```{r combine-rd-sw-sh, echo = TRUE}
+```{r combine-rd-sw-sh}
 # combine the sidewalk and shoulder and roads data
@@
 
-#distance is the length of the matched road segment and sidewalk or shoulder 
-#comb %>% st_set_geometry(., NULL) %>% 
-#  group_by(road_class, SIDE_OF_RO, TYPE) %>%
-#  summarize(s_miles = sum(distance)/5280,
-#            miles = format(round(as.numeric(sum(distance)/5280), 1), 
-#                           digits = 1, nsmall=1, big.mark=","),
-#            .groups = "drop")
 ```
@@
 road_standards_given <- tribble(
-  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, ~note,
-  "Urban", "Local Road", NA, 5, NA,NA,
-  "Urban", "Local Sub-collector", NA, 5, NA,"sub-collector is not defined in road data",
-  "Urban", "Arterial", NA, 6, NA,NA,
-  "Urban", "Collector", NA, 6, NA,NA,
-  "Rural", "Local Road", NA, NA, 3,NA,
-  "Rural", "Local Sub-collector", NA, NA, 4,"sub-collector is not defined in road data",
-  "Rural", "Collector", "ADT 400-750", NA, 3,"ADT not defined in road data",
-  "Rural", "Collector", "ADT 751-1000", NA, 4,"ADT not defined in road data",
-  "Rural", "Collector", "DHV 100-200", NA, 6,"DHV not defined in road data",
-  "Rural", "Collector", "DHV > 200", NA, 8,"DHV not defined in road data",
-  "Rural", "Minor Arterial", "DHV < 100", NA, 4,"DHV not defined in road data",
-  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,"DHV not defined in road data",
-  "Rural", "Minor Arterial", "DHV > 200", NA, 8,"DHV not defined in road data",
-  "Rural", "Principal Arterial", "DHV <200", NA, 6,"DHV not defined in road data",
-  "Rural", "Principal Arterial", "DHV > 200", NA, 8,"DHV not defined in road data",
+  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, 
+  "Urban", "Local Road", NA, 5, NA,
+  "Urban", "Local Sub-collector", NA, 5, NA,
+  "Urban", "Arterial", NA, 6, NA,
+  "Urban", "Collector", NA, 6, NA,
+  "Rural", "Local Road", NA, NA, 3,
+  "Rural", "Local Sub-collector", NA, NA, 4,
+  "Rural", "Collector", "ADT 400-750", NA, 3,
+  "Rural", "Collector", "ADT 751-1000", NA, 4,
+  "Rural", "Collector", "DHV 100-200", NA, 6,
+  "Rural", "Collector", "DHV > 200", NA, 8,
+  "Rural", "Minor Arterial", "DHV < 100", NA, 4,
+  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,
+  "Rural", "Minor Arterial", "DHV > 200", NA, 8,
+  "Rural", "Principal Arterial", "DHV <200", NA, 6,
+  "Rural", "Principal Arterial", "DHV > 200", NA, 8,
 )
@@
              caption = "Summary of Tables 3.3 and 3.4 in Kitsap County Road Standards [@kc_road_standards]")  %>%
-   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT")
+   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT.
+Note: sub-collector, ADT, and DHV are not defined in the publicly available road data.") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
 ```
@@
              caption = "Simplified Assumptions to Allow for Evaluation")  %>%
-   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation")
+   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
 ```
@@
 
-|Segment|Segment Score | Segment Centerline Linear Feet |
-|------|:-----|:-----|
-|A|6|100|
-|B|0|10|
-|C|10|50|
-|D|8|200|
-|*Total*| *?* | *360*|
+```{r}
+example_tbl <- tribble(
+  ~Segment, ~"Segment Score", ~"Segment Centerline Linear Feet",
+  "A", 6, 100,
+  "B", 0, 10,
+  "C",10,50,
+  "D",8,200,
+  "Total",7.5 ,360,
+)
+
+knitr::kable(example_tbl) %>%
+kable_styling(bootstrap_options = "striped", full_width = F, position = "float_left")
+```
 
@@
              caption = "Interpretation of Scores for Segments and for Areas")  %>%
-   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums")
+   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
 ```
@@
 
+Looking by the same breakdown that we have for standards, we can see where the scores are highest and lowest.
+
 ```{r}
-knitr::kable(comb_score_by_setting_class)
+knitr::kable(comb_score_by_setting_class) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
 ```
 
-More detailed information
 ```{r}
-knitr::kable(
-  comb_score2 %>%
-  group_by(setting, road_class, score) %>% 
-  summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
-            segments = nice_comma(n()), 
-            #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
-            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
-            .groups = "drop") 
-  )
+# More detailed information
+# knitr::kable(
+#   comb_score2 %>%
+#   group_by(setting, road_class, score) %>% 
+#   summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
+#             segments = nice_comma(n()), 
+#             #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
+#             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
+#             .groups = "drop") 
+#   )
 ```
@@
 
-knitr::kable(head(roads_0))
+knitr::kable(head(roads_0)) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
 
@@
 
+Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road.
+A few details for roads where at least one segment has a 10 are shown here.
 ```{r}
@@
 
-knitr::kable(roads_10)
+knitr::kable(roads_10) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
 ```

### 2022-10-06 Draft Ped Metrics Example.Rmd
store scores of breakdowns and builds isochrones

@@
 library(sf)
+
 kitsap_all_roads <- kitsap +
@@
 
+Some road segments are missing entirely from the centerline road data - for example, the portion of Rigetop that goes from the end of Myhre and curves back to Silverdale Way.
+These missing road segments are mostly complete streets with wide sidewalks.
+Here, showing a map with all roads with the name *RIDGETOP*, you can only see the part going from Silverdale Way to Myhre.
+These segments, and any others like them, missing from the data decrease the overall pedestrian facilities score.
+
+```{r}
+kitsap +
+  geom_sf(data = filter(clean_roadcl, STREETNAME == "RIDGETOP"), aes(color = road_class)) + 
+  road_col +
+  theme_void() +
+  labs(title = "STREETNAME == RIDGETOP",
+       color = "")
+```
+
+## Smaller Breakdowns of the County
+
 We can also look at regions within the County.
 For these, we identify the roads and any sidewalks or shoulders.^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]
+
+```{r}
+score_breakdowns <- tibble(Area = "Kitsap County", Score = comb_total_score)
+```
 
@@
 loop_df <- st_as_sf(uga_no_city) %>% mutate(tomatch = GMA_JURISD)
+
+score_breakdowns <- score_breakdowns %>%
+  bind_rows(tibble(Area = the_ugas, Score = NA))
 
@@
 
+score_breakdowns <- score_breakdowns %>%
+  bind_rows(tibble(Area = the_lamirds, Score = NA))
+
 expanded_child <- lapply(
@@
   filter(TYPE == "PUBLIC LIBRARY")
+
+getWalkingPolygonInCounty <- function(locdf, idcol) {
+  # finds the items in locdf that are in the target area
+  # then finds the polygon around them
+  # then limits that again to the area in the target (because the polygon may go into a city)
+  fdf <- st_intersection(locdf, st_buffer(county_no_cities,0)) 
+  mb_isochrone(fdf,
+               time = 30, "walking", id_column=idcol)
+  #fdf <- st_transform(fdf, 9003)
+  #st_intersection(fdf, st_buffer(county_no_cities,0)) 
+}
+library(mapboxapi)
+lib_30min <- mb_isochrone(libraries,
+               time = 30, "walking", id_column="NAME")
 ```
 
-Kitsap County has `r nrow(libraries)`
+Kitsap County has `r nrow(libraries)` Public libraries.
+
+```{r plot_libraries}
+
+lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0)) %>%
+  mutate(LENGTH = as.integer(st_length(.)))
+lib2 <- getWalkingPolygonInCounty(libraries, "NAME")
+
+kitsap +
+  #geom_sf(data = libraries, aes(color = NAME)) +
+  geom_sf(data = lib2, aes(fill = id), 
+          lwd = 0, alpha = 0.3, show.legend = FALSE) 
+  geom_sf(data = lib_circles, aes(fill = NAME), 
+          lwd = 0, alpha = 0.3, show.legend = FALSE) +
+  geom_sf(data = st_buffer(libraries, 5280), aes(fill = NAME), 
+          lwd = 0, alpha = 0.3, show.legend = FALSE) +
+  theme_void() +
+  labs(title = 'Kitsap County libraries',
+       caption = "Data source: Kitsap GIS - data in development",
+       color = "Libraries")
+
+```
 
@@
 
+Using the data publicly available and draft criteria (special committee investigating pedestrian metrics), the overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, which is interpreted (see methods) as `r getScoreInterpretation(comb_total_score)`.
+
+Smaller breakdowns of the county fare differently.
+
+```{r}
+score_breakdowns %>%
+  rowwise() %>%
+  mutate(Interpretation = getScoreInterpretation(Score))
+
+knitr::kable(score_breakdowns) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+However, there are a couple of known data irregularities, such as:
+
+ * missing part of Ridgetop Blvd in Silverdale where sidewalks are known to exist
+ * unmatched road log identifiers between road and sidewalk and shoulder data
+
+ And there are likely unknown data irregularities.
+
+In addition, many roads have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**
+
 ## References

### 2022-10-06 Draft Ped Metrics Example.Rmd
make sure only nc roads

@@
   #st_intersection(fdf, st_buffer(county_no_cities,0)) 
+  # doesn't do the clip off of the last part
 }

### 2022-10-06 child_single_geo.Rmd
store scores of breakdowns and builds isochrones
make sure only nc roads

@@
 this_geo <- st_as_sf(loop_df) %>% filter(tomatch == "{{current_geo}}") 
-this_roads <- st_intersection(road_df, st_buffer(this_geo,0))
+this_roads <- st_intersection(ncdf, st_buffer(this_geo,0))
 this_sw_sh <- st_intersection(nc_sw_sh, st_buffer(this_geo,0))
@@
 ```{r}
+# add the score to the score breakdowns
+if (!exists(deparse(substitute(this_comb_total_score)))){
+  this_comb_total_score <- 0
+}
+score_breakdowns$Score[score_breakdowns$Area == "{{current_geo}}"] <- this_comb_total_score
+```
+
+```{r}
 # next step

### 2022-10-06 Draft Ped Metrics Example.Rmd
adds library loop

--- 2022-10-06 Draft Ped Metrics Example.Rmd
@@
   #st_intersection(fdf, st_buffer(county_no_cities,0)) 
+  # doesn't do the clip off of the last part
 }
 library(mapboxapi)
-lib_30min <- mb_isochrone(libraries,
-               time = 30, "walking", id_column="NAME")
+
+lib2 <- getWalkingPolygonInCounty(libraries, "NAME")
 ```
 
-Kitsap County has `r nrow(libraries)` Public libraries.
+Kitsap County has `r nrow(libraries)` Public libraries; however, only `r nrow(lib2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
 
@@
 
-lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0)) %>%
-  mutate(LENGTH = as.integer(st_length(.)))
-lib2 <- getWalkingPolygonInCounty(libraries, "NAME")
+#lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0))
 
 kitsap +
-  #geom_sf(data = libraries, aes(color = NAME)) +
   geom_sf(data = lib2, aes(fill = id), 
-          lwd = 0, alpha = 0.3, show.legend = FALSE) 
-  geom_sf(data = lib_circles, aes(fill = NAME), 
-          lwd = 0, alpha = 0.3, show.legend = FALSE) +
-  geom_sf(data = st_buffer(libraries, 5280), aes(fill = NAME), 
-          lwd = 0, alpha = 0.3, show.legend = FALSE) +
+          lwd = 0, alpha = 0.3) +
+  geom_sf(data = libraries, aes(color = 1), color = "black", show.legend = FALSE) +
   theme_void() +
@@
        caption = "Data source: Kitsap GIS - data in development",
-       color = "Libraries")
+       fill = "Libraries")
 
 ```
+
+```{r loop_libraries,include=FALSE,echo=FALSE}
+the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
+loop_df <- st_as_sf(lib2) %>% mutate(tomatch = id)
+
+score_breakdowns <- score_breakdowns %>%
+  bind_rows(tibble(Area = the_libraries, Score = NA))
+
+expanded_child <- lapply(
+    the_libraries,
+    function(x) {
+      knitr::knit_expand(file = "child_single_geo.Rmd",
+                                   current_geo = x )
+    })
+
+parsed_child <- knitr::knit_child(text = unlist(expanded_child))
+```
+
+`r parsed_child`
 

### 2022-10-07 child_single_geo.Rmd
handles different crs

@@
 ```{r}
-#library(broom)
 
 this_geo <- st_as_sf(loop_df) %>% filter(tomatch == "{{current_geo}}") 
+if (!is.na(st_crs(ncdf)$epsg) & !is.na(st_crs(this_geo)$epsg) & 
+    st_crs(ncdf)$epsg != st_crs(this_geo)$epsg) {
+  ncdf <- st_transform(ncdf, st_crs(this_geo)$epsg)
+  nc_sw_sh <- st_transform(ncdf, st_crs(this_geo)$epsg)
+}
 this_roads <- st_intersection(ncdf, st_buffer(this_geo,0))

### 2022-10-07 Draft Ped Metrics Example.Rmd
handles different crs and adds transit

@@
 community <- sf::read_sf(paste0(this_dir,"data/community"))
+schools <- sf::read_sf(paste0(this_dir,"data/schools"))
 
@@
 
-#### Schools
-
-#### Libraries
-
 ```{r}
-libraries <- community %>%
-  filter(TYPE == "PUBLIC LIBRARY")
-
 getWalkingPolygonInCounty <- function(locdf, idcol) {
@@
   # then limits that again to the area in the target (because the polygon may go into a city)
+  if (!is.na(st_crs(locdf)$epsg)) { # the locdf has a epsg and cnc is NA
+    cnc <- st_transform(county_no_cities, st_crs(locdf)$epsg)
+    fdf <- st_intersection(locdf, st_buffer(cnc,0)) 
+  } else {
     fdf <- st_intersection(locdf, st_buffer(county_no_cities,0)) 
-  mb_isochrone(fdf,
+  }
+  fdf2 <- mapboxapi::mb_isochrone(fdf,
                                   time = 30, "walking", id_column=idcol)
-  #fdf <- st_transform(fdf, 9003)
-  #st_intersection(fdf, st_buffer(county_no_cities,0)) 
-  # doesn't do the clip off of the last part
+  cnc <- st_transform(county_no_cities, st_crs(fdf2)$epsg)
+  st_intersection(fdf2, st_buffer(cnc,0)) 
 }
-library(mapboxapi)
+```
+
+#### Schools
+```{r}
+libraries <- community %>%
+  filter(TYPE == "PUBLIC LIBRARY")
 
@@
   labs(title = 'Kitsap County libraries',
+       subtitle = "Libraries not labeled with color are not in the unincorporated county",
+       caption = "Data source: Kitsap GIS - data in development",
+       fill = "Libraries")
+
+```
+
+```{r loop_libraries,include=FALSE,echo=FALSE}
+the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
+loop_df <- st_as_sf(lib2) %>% mutate(tomatch = id)
+
+score_breakdowns <- score_breakdowns %>%
+  bind_rows(tibble(Area = the_libraries, Score = NA))
+
+expanded_child <- lapply(
+    the_libraries,
+    function(x) {
+      knitr::knit_expand(file = "child_single_geo.Rmd",
+                                   current_geo = x )
+    })
+
+parsed_child <- knitr::knit_child(text = unlist(expanded_child))
+```
+
+`r parsed_child`
+
+#### Libraries
+
+```{r}
+libraries <- community %>%
+  filter(TYPE == "PUBLIC LIBRARY")
+
+lib2 <- getWalkingPolygonInCounty(libraries, "NAME")
+```
+
+Kitsap County has `r nrow(libraries)` Public libraries; however, only `r nrow(lib2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
+
+```{r plot_libraries}
+
+#lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0))
+
+kitsap +
+  geom_sf(data = lib2, aes(fill = id), 
+          lwd = 0, alpha = 0.3) +
+  geom_sf(data = libraries, aes(color = 1), color = "black", show.legend = FALSE) +
+  theme_void() +
+  labs(title = 'Kitsap County libraries',
+       subtitle = "Libraries not labeled with color are not in the unincorporated county",
        caption = "Data source: Kitsap GIS - data in development",
@@
 
+According to Kitsap Transit, "Transit Centers and Transfer Points are locations where Kitsap Transit buses and ferries meet up to allow the safe transfer of passengers from one route to another.
+"[@ktransit]^[Transit Center location data were manually copied from the Kitsap Transit Website because they are not one of the identified data sets on the Kitsap County GIS site.
+Future users will need to ensure that these addresses are correct.]
+
+```{r}
+transit_cp <- tribble(
+  ~NAME, ~ADDRESS,
+  "Bremerton Transportation Center (Ferry Terminal)","10 Washington Ave, Bremerton 98337",
+  "Bainbridge Island Ferry Terminal","270 Olympic Drive SE, Bainbridge 98110",
+  "Wheaton Way Transit Center","3915 Wheaton Way, Bremerton 98310",
+  "West Bremerton Transit Center","540 Bruenn Ave, Bremerton 98312",
+  "Silverdale Transit Center","Greaves Way & Kitsap Mall Blvd, Silverdale 98383",
+  "North Viking Transit Center","21992 Viking Ave NW, Poulsbo 98370",
+  "Kingston Ferry Terminal","11264 State Route 104, Kingston 98346",
+  "Port Orchard Ferry Dock","73 Sidney Ave, Port Orchard 98366",
+  "Southworth Ferry Terminal","11564 SE State Hwy. 160, Southworth 98386",
+  "Annapolis Ferry Dock","1076 Bay Street, Port Orchard 98366",
+  "Hwy. 305 & Suquamish Way","16003 WA-305, Poulsbo, Washington 98370",
+  "McWilliams Park & Ride","1601 NE McWilliams Rd, Bremerton, Washington 98311",
+  "Suquamish Park & Ride","18829 Division Ave NE, Suquamish, Washington 98392",
+  "Port Orchard Wal-Mart","3497 Bethel Rd SE, Port Orchard 98366",
+  "George's Corner Park & Ride","27618 Hansville Rd NE, Kingston 98346",
+  "Gateway Fellowship Park & Ride","18901 8th Ave NE, Poulsbo 98370",
+  "Miller Bay & Indianola","23404 Miller Bay Rd NE Poulsbo, Washington 98370",
+  "St. Gabriel's Church","1150 Mitchell Ave SE, Port Orchard 98366"
+)
+# geocode the addresses into geocode = c(lon, lat)
+transit_cp <- transit_cp %>% #tmaptools::geocode_OSM(transit_cp$ADDRESS)
+  rowwise() %>%
+  mutate(geocode = list(set_names(mapboxapi::mb_geocode(ADDRESS),
+                        c("lon","lat")))) %>%
+  unnest_wider(geocode)
+
+# convert the lat lon into a point geometry
+transit_cp <- st_as_sf(transit_cp, 
+                      coords = c(x = "lon", y = "lat"), 
+                      crs = 4326) 
+transit_cp2 <- getWalkingPolygonInCounty(transit_cp, "NAME")
+```
+
+Kitsap County has `r nrow(transit_cp)` Public libraries; however, only `r nrow(transit_cp2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
+
+```{r plot_transit}
+kitsap +
+  geom_sf(data = transit_cp2, aes(fill = id), 
+          lwd = 0, alpha = 0.3) +
+  geom_sf(data = transit_cp, aes(color = 1), color = "black", show.legend = FALSE) +
+  theme_void() +
+  labs(title = 'Kitsap County Transit Centers and Transfer Points',
+       subtitle = "Libraries not labeled with color are not in the unincorporated county",
+       caption = "Data source: Kitsap Transit Website[@ktransit]",
+       fill = "Transit Centers")
+
+```
+
+```{r loop_transit}
+the_transit <- unique(st_set_geometry(transit_cp2, NULL)$id)
+loop_df <- st_as_sf(transit_cp2) %>% mutate(tomatch = id)
+
+score_breakdowns <- score_breakdowns %>%
+  bind_rows(tibble(Area = the_transit, Score = NA))
+
+expanded_child <- lapply(
+    the_transit,
+    function(x) {
+      knitr::knit_expand(file = "child_single_geo.Rmd",
+                                   current_geo = x )
+    })
+
+parsed_child <- knitr::knit_child(text = unlist(expanded_child))
+```
+
+`r parsed_child`
+
 ## Conclusion

### 2022-10-07 Draft Ped Metrics Example.Rmd
adds public schools

@@
 ```{r}
-libraries <- community %>%
-  filter(TYPE == "PUBLIC LIBRARY")
-
-lib2 <- getWalkingPolygonInCounty(libraries, "NAME")
-```
-
-Kitsap County has `r nrow(libraries)` Public libraries; however, only `r nrow(lib2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
-
-```{r plot_libraries}
-
-#lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0))
-
+pub_schools <- schools %>%
+  filter(TYPE == "PUBLIC")
+
+pub_schools2 <- getWalkingPolygonInCounty(pub_schools, "NAME")
+```
+
+Kitsap County has `r nrow(pub_schools)` Public schools; however, only `r nrow(pub_schools2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
+
+```{r plot_pubschools}
 kitsap +
-  geom_sf(data = lib2, aes(fill = id), 
+  geom_sf(data = pub_schools2, aes(fill = id), 
           lwd = 0, alpha = 0.3) +
-  geom_sf(data = libraries, aes(color = 1), color = "black", show.legend = FALSE) +
+  geom_sf(data = pub_schools, aes(color = 1), color = "black", show.legend = FALSE) +
   theme_void() +
-  labs(title = 'Kitsap County libraries',
-       subtitle = "Libraries not labeled with color are not in the unincorporated county",
+  labs(title = 'Kitsap County Public Schools',
+       subtitle = "Schools not labeled with color are not in the unincorporated county",
        caption = "Data source: Kitsap GIS - data in development",
-       fill = "Libraries")
-
-```
-
-```{r loop_libraries,include=FALSE,echo=FALSE}
-the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
-loop_df <- st_as_sf(lib2) %>% mutate(tomatch = id)
+       fill = "Public Schools")
+
+```
+
+```{r loop_pubschools,include=FALSE,echo=FALSE}
+the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
+loop_df <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)
 
 score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_libraries, Score = NA))
+  bind_rows(tibble(Area = the_schools, Score = NA))
 
 expanded_child <- lapply(
-    the_libraries,
+    the_schools,
     function(x) {
@@
   labs(title = 'Kitsap County Transit Centers and Transfer Points',
-       subtitle = "Libraries not labeled with color are not in the unincorporated county",
+       subtitle = "Centers not labeled with color are not in the unincorporated county",
        caption = "Data source: Kitsap Transit Website[@ktransit]",

### 2022-10-08 child_single_geo.Rmd
stores score_breakdowns, fixes loop printing

--- 2022-10-06 child_single_geo.Rmd
@@
 ```{r}
-#library(broom)
 
 this_geo <- st_as_sf(loop_df) %>% filter(tomatch == "{{current_geo}}") 
+if (!is.na(st_crs(this_geo)$epsg)) {
+  ncdf <- st_transform(ncdf, st_crs(this_geo)$epsg)
+  nc_sw_sh <- st_transform(nc_sw_sh, st_crs(this_geo)$epsg)
+}
 this_roads <- st_intersection(ncdf, st_buffer(this_geo,0))
@@
 # do not evaluate most of this child if does not
+this_comb_total_score <-0
 ```
@@
 this_comb_miles <- 0
+
 ```

### 2022-10-08 child_single_geo.Rmd
updates zero and ten tables

@@
                            digits = 1, nsmall=1, big.mark=",")
+this_msg <- ifelse(this_comb_miles > this_sw_sh_miles, 
+                   "There are duplicate segments matching.", 
+                   "")
 
@@
 Within this area, there are `r this_centerline_miles` centerline road miles; `r this_sw_sh_miles` miles of recorded sidewalk or shoulder (either side of road); and `r this_comb_miles` miles where we could match the recorded sidewalk or shoulder to a road segment (either side of road).
-
-```{r, eval=has_both}
-
-road_sum <- this_roads %>%
-  mutate(distance = st_length(.)) %>%  
-  st_set_geometry(., NULL) %>% 
-  group_by(road_class) %>%
-  summarise(t_miles = sum(distance)/5280,
-            miles = format(round(as.numeric(sum(distance)/5280), 1), 
-                           digits = 1, nsmall=1, big.mark=","),
-            .groups = "drop")
-df <- select(road_sum, -t_miles)
-colnames(df) <- c("Road Class", "Miles")
-knitr::kable(df) %>%
-  kable_material(c("striped", "hover"))
-```
-
-And we still need to get this into segment scores, but here are the miles with sidewalks or shoulders.
-
-```{r, eval=has_both}
-df <- this_comb %>% st_set_geometry(., NULL) %>% 
-  group_by(road_class, SIDE_OF_RO, TYPE) %>%
-  summarize(s_miles = sum(distance)/5280,
-            miles = format(round(as.numeric(sum(distance)/5280), 1), 
-                           digits = 1, nsmall=1, big.mark=","),
-            .groups = "drop")
-knitr::kable(select(df,-s_miles)) %>%
-  kable_material(c("striped", "hover"))
-```
+`r this_msg`
 
@@
   geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
-  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
   geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
@@
   geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
-  #geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(linetype = as.factor(RD_LOG_ID)), color = "black") +
   geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
@@
 
-p2 <- ggplot() +
-  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
-  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
-  geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = FALSE) +
-  theme_void()
-
 library(patchwork)
 
-(p1 + #l + r + #p2 +
+(p1 + 
   plot_layout(guides = 'collect') +
@@
 
+A few segments with a score of 0 are shown here. 
+```{r eval=has_both}
+
+roads_0 <- this_comb_score %>%
+  filter(score == 0) %>%
+  select(setting, road_class, FULL_NAME_R, 
+         sidewalk_right, sidewalk_left,
+         shoulder_right, shoulder_left,
+         standard_sidewalk, standard_shoulder)
+
+knitr::kable(head(roads_0) %>%
+               rename(Setting = setting,
+                      Class = road_class,
+                      Name = FULL_NAME_R,
+                      `sidewalk R` = sidewalk_right,
+                      `sidewalk L` = sidewalk_left,
+                      `shoulder R` = shoulder_right,
+                      `shoulder L` = shoulder_left,
+                      `standard sidewalk` = standard_sidewalk,
+                      `standard shoulder` = standard_shoulder)) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+
+```
+
 ```{r}
@@
 ```
-
-```{r}
-# next step
-# take the comb data, get the distance for each segment
-# combine that with this roads
-# where no sw sh = 0
-# determine if meets criteria
-
-```

### 2022-10-08 Draft-Ped-Metrics-Example.Rmd
change scoring to allow for sidewalks when shoulder is standard

--- 2022-10-07 Draft Ped Metrics Example.Rmd
@@
 
-- 5 if it has a sidewalk or shoulder of adequate size (based on road classification & standards)
+- 5 if it has a sidewalk or shoulder of adequate size (based on road classification & standards) or a sidewalk is present where a shoulder is the standard
 - 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
-- 1 if the sidewalk or shoulder is present and is narrower than AASHTO
+- 1 if the sidewalk or shoulder is present and is narrower than AASHTO (**NOTE to subcommittee - do we think we should use a 1 if there is a shoulder when there should be a sidewalk?**)
 - 0 if none present or no data
@@
   3, "Minimum",
-  0, "Poor",
+  1, "Poor",
+  0, "Nothing"
 )
@@
          case_when(
+           !is.na(sidewalk) ~ 5, # there is a sidewalk present when a shoulder is the standard
            shoulder >= standard_shoulder ~ 5,
@@
 Some of the segments with score of 0 are because they have shoulders when they should have sidewalks.
+If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.
 Here's a few example segments.
-If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.
 
@@
   filter(score == 0) %>%
-  select(setting, road_class, FULL_NAME_R, sidewalk_right:score)
-
-knitr::kable(head(roads_0)) %>%
+  select(setting, road_class, FULL_NAME_R, sidewalk_right:standard_sidewalk)
+
+knitr::kable(head(roads_0) %>%
+               rename(Setting = setting,
+                      Class = road_class,
+                      Name = FULL_NAME_R,
+                      `sidewalk R` = sidewalk_right,
+                      `sidewalk L` = sidewalk_left,
+                      `shoulder R` = shoulder_right,
+                      `shoulder L` = shoulder_left,
+                      `standard sidewalk` = standard_sidewalk)) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
@@
   filter(score == 10) %>%
-  group_by(setting, road_class, "Road Name" = FULL_NAME_R) %>%
-  summarise(segments = nice_comma(n()), 
+  group_by(setting, road_class, FULL_NAME_R) %>%
+  summarise(segments = nice_comma(n(),0), 
             road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
-            .groups = "drop")
+            .groups = "drop") %>%
+  rename(Setting = setting,
+         Class = road_class,
+         Name = FULL_NAME_R,
+         `Segments w 10` = segments,
+         `Segment Length` = road_len,
+         `Ped Fac Length` = dist)
 
@@
 ```
+
+### Commissioner Districts
+
+```{r plot_commdists}
+commdists <- sf::read_sf(paste0(this_dir,"data/commdist"))
+commdists <- st_intersection(commdists, st_buffer(county_no_cities,0)) 
+commdists <- st_as_sf(commdists) %>% 
+  mutate(NAME = ifelse(DISTRICT == "1", "1 North",
+                ifelse(DISTRICT == "2", "2 South", "3 Central")))
+
+kitsap +
+  geom_sf(data = commdists, aes(fill = NAME), alpha = 0.5) +
+  theme_void() +
+  labs(title = 'Kitsap County Commissioner Districts',
+       caption = "Data source: Kitsap GIS - data in development",
+       color = "DIMS")
+
+```
+
+```{r loop_commdists,include=FALSE,echo=FALSE}
+the_commdists <- unique(st_set_geometry(commdists, NULL)$NAME)
+loop_df <- st_as_sf(commdists) %>% mutate(tomatch = NAME)
+
+score_breakdowns <- score_breakdowns %>%
+  bind_rows(tibble(Area = the_commdists, Score = NA))
+
+expanded_child <- lapply(
+    the_commdists,
+    function(x) {
+      knitr::knit_expand(file = "child_single_geo.Rmd",
+                                   current_geo = x )
+
+    }
+    )
+
+parsed_child <- knitr::knit_child(text = unlist(expanded_child))
+
+```
+
+Kitsap County has `r length(the_commdists)` Commissioner Districts.
+
+`r parsed_child`
 
@@
 
-```{r loop_libraries,include=FALSE,echo=FALSE}
+```{r loop_libraries,include=FALSE,echo=FALSE, cache = TRUE}
 the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
@@
 
-```{r loop_transit}
-the_transit <- unique(st_set_geometry(transit_cp2, NULL)$id)
+```{r loop_transit,include=FALSE,echo=FALSE, cache = TRUE}
+the_transits <- unique(st_set_geometry(transit_cp2, NULL)$id)
 loop_df <- st_as_sf(transit_cp2) %>% mutate(tomatch = id)
@@
 score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_transit, Score = NA))
+  bind_rows(tibble(Area = the_transits, Score = NA))
 
 expanded_child <- lapply(
-    the_transit,
+    the_transits,
     function(x) {
@@
 ```{r}
-score_breakdowns %>%
+
+score_breakdowns <- score_breakdowns %>%
   rowwise() %>%
   mutate(Interpretation = getScoreInterpretation(Score))
-
-knitr::kable(score_breakdowns) %>%
+write_csv(score_breakdowns, paste0(this_dir,"score_breakdowns.csv"))
+
+getScoreBreakdownForArea <- function(the_area) {
+  knitr::kable(score_breakdowns %>% 
+                 filter(Area %in% eval(as.symbol(the_area))) %>%
+                 mutate(Score = nice_comma(Score,2) 
+                                           ),
+             caption = paste("Scores for",
+                             str_to_title(sub("the_"," ",the_area))
+                             )
+             ) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+}
+
+```
+
+```{r, echo=FALSE, results='asis'}
+score_breakdowns <- read_csv(paste0(this_dir,"score_breakdowns.csv"))
+res <- lapply(c("the_commdists","the_ugas", "the_lamirds", "the_schools", "the_libraries", "the_transits"), function(x) {
+  knitr::knit_child(text = c(
+    '## `r gsub("_"," ",x)`',
+    '',
+    '```{r}',
+    'getScoreBreakdownForArea(x)',
+    '```',
+    ''
+  ), envir = environment(), quiet = TRUE)
+})
+cat(unlist(res), sep = '\n')
+
 ```
@@
 
-In addition, many roads have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**
+In addition, many road segments have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**
+
+### Notes to the special committee drafting this
+We have a few to dos on this before we move forward.
+
+1. Be sure that our criteria make sense to us
+2. Consider how the scores should be interpreted - does the interpretation fit?
+3. Consider if we are missing something critical
 

### 2022-10-08 child_single_geo.Rmd
handles adding points when they exist

@@
 this_geo <- st_as_sf(loop_df) %>% filter(tomatch == "{{current_geo}}")
+
 if (!is.na(st_crs(this_geo)$epsg)) {
@@
 ```{r, eval=has_both}
+ # should we test for points?
+the_points <- c()
+if (exists("the_schools")) { the_points <- append(the_points, "the_schools")} 
+if (exists("the_libraries")) { the_points <- append(the_points, "the_libraries")} 
+if (exists("the_transits")) { the_points <- append(the_points, "the_transits")} 
+check_points <- ifelse(length(the_points) > 0, T, F)
+has_point = FALSE # start with false
+```
+
+```{r, eval=has_both & check_points}
+  # see if this geo has a point (is it an area around a place, like a school)
+
+if (length(which(eval(as.symbol(the_points)) == this_geo$tomatch)) !=0) {
+  # we need to figure out which one
+  has_point = TRUE
+  tr <- which(eval(as.symbol(the_points)) == this_geo$tomatch)
+  if (tr <= length(the_schools)) { # it is a school
+    this_point <- st_as_sf(pub_schools) %>%
+      filter(NAME == this_geo$tomatch)
+  } else if (tr > length(the_schools) & 
+             tr <= length(the_schools + length(the_libraries))) { # library
+    this_point <- st_as_sf(libraries) %>%
+      filter(NAME == this_geo$tomatch)
+  } else { #transit
+    this_point <- st_as_sf(transit_cp) %>%
+      filter(NAME == this_geo$tomatch)
+  }
+}
+
+```
+
+```{r, eval=has_both, fig.align = "center", out.width = '100%'}
 p1 <- kitsap +
@@
   geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
+  {if(has_point) geom_sf(data = this_point, 
+                         aes(color = 1), color = "black", show.legend = FALSE)} +
   sw_sh_col +
@@
   geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
+  {if(has_point) geom_sf(data = this_point, 
+                         aes(color = 1), color = "black", show.legend = FALSE)} +
   sw_sh_col +

### 2022-10-08 Draft-Ped-Metrics-Example.Rmd
names cached chunk

@@
 
-```{r, cache = TRUE}
+```{r kitsapmap, cache = TRUE}
 library(tidyverse)
@@
   1, "Poor",
+  0.001, "Very Poor",
   0, "Nothing"

### 2022-10-09 Draft-Ped-Metrics-Example.Rmd
adjusts map plot sizing and corrects the points

@@
 title: "DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report"
-output: html_document
+date: "`r Sys.Date()`"
+bibliography: metrics.bib
+output: 
+  html_document:
+    toc: true
+    toc_depth: 3
     fig_caption: yes
-bibliography: metrics.bib
-date: "`r Sys.Date()`"
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, 
+                      message = FALSE, 
+                      warning = FALSE, 
+                      message = FALSE,
+                      fig.align = 'center')
 options(scipen=999)
@@
 
-## Purpose
+# Purpose
 
@@
 
-## Introduction
+# Introduction
 
@@
 ```{r kitsapmap, cache = TRUE}
+
+#| fig.align = "center"
+#| fig.width = 12
+#| fig.height = 8
+#| out.width = '100%'
+
 library(tidyverse)
@@
 
-## Data and Methods
-
-### Data {#datasection}
+# Data and Methods
+
+## Data {#datasection}
 
@@
 
-#### Combining the Data
+### Combining the Data
 
@@
 
-##### Matching on Road Log Id is good but insufficient
+#### Matching on Road Log Id is good but insufficient
 
@@
 
-### Method of Evaluation
+## Method of Evaluation
 
@@
 Because of all of the items that are not available in the road data at this time, this draft makes some assumptions to allow for an evaluation.
-When the information is known, a more accurate evaluation can be made.
+When the information is known, a more accurate evaluation can be made.^[Roads with unknown setting - RUCODE is 0, we match them to the Urban road standards.
+This may be incorrect.]
 
@@
 
-## Results
+# Results
 First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.
 
-### Overall
+## Overall
 
@@
 
+Regions are already defined by an area.
+For pedestrian generators, we identify an area that is within 30 minute walking distance of the facility and evaluate roads therein.^[30 minute walking distance calculated using the mapboxapi function `mb_isochrone`.
+Isochrones use map data from Mapbox and OpenStreetMap and their data sources.
+To learn more, visit https://www.mapbox.com/about/maps/ and http://www.openstreetmap.org/copyright.[@mapbox]]
+
+These regions and pedestrian generators are:
+ * Commissioner Districts
+ * UGAs
+ * LAMIRDs
+ * Schools
+ * Libraries
+ * Transit Centers
+
 ```{r}
@@
 `r parsed_child`
-
-### Pedestrian Generators
-
-For pedestrian generators, we identify an area that is within 1 mile of the facility and evaluate roads therein in the same manner as the defined county areas, like UGAs and LAMIRDS.
 
@@
 
-#### Schools
+### Schools
 ```{r}
@@
 ```{r plot_pubschools}
+#| fig.width = 8,
+#| fig.height = 8,
+#| out.width = '100%',
+#| dpi = 96
+
 kitsap +
@@
        subtitle = "Schools not labeled with color are not in the unincorporated county",
-       caption = "Data source: Kitsap GIS - data in development",
+       caption = "Data source: Kitsap GIS - data in development\n
+                  Walking distance polygons from Mapbox",
        fill = "Public Schools")
@@
 
-```{r loop_pubschools,include=FALSE,echo=FALSE}
+```{r loop_pubschools,include=FALSE,echo=FALSE, cache=TRUE}
 the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
@@
 
-#### Libraries
+### Libraries
 
@@
 ```{r plot_libraries}
-
-#lib_circles <- st_intersection(st_buffer(libraries, 5280), st_buffer(county_no_cities,0))
-
 kitsap +
@@
        subtitle = "Libraries not labeled with color are not in the unincorporated county",
-       caption = "Data source: Kitsap GIS - data in development",
+       caption = "Data source: Kitsap GIS - data in development\n
+                  Walking distance polygons from Mapbox",
        fill = "Libraries")
@@
 
-#### Transit
+### Transit
 
@@
        subtitle = "Centers not labeled with color are not in the unincorporated county",
-       caption = "Data source: Kitsap Transit Website[@ktransit]",
+       caption = "Data source: Kitsap Transit Website\n
+                  Walking distance polygons from Mapbox",
        fill = "Transit Centers")
@@
   knitr::knit_child(text = c(
-    '## `r gsub("_"," ",x)`',
+    '## `r str_to_title(gsub("_"," ",x))`',
     '',

### 2022-10-09 child_single_geo.Rmd
adjusts map plot sizing and corrects the points

@@
-# {{current_geo}}
+#### {{current_geo}}
 
@@
 # do not evaluate most of this child if does not
+
+# reset
 this_comb_total_score <-0
+has0 <- F
+has10 <- F
+this0 <- ""
+this10 <- ""
 ```
@@
   } else if (tr > length(the_schools) & 
-             tr <= length(the_schools + length(the_libraries))) { # library
+             tr <= length(the_schools) + length(the_libraries)) { # library
     this_point <- st_as_sf(libraries) %>%
@@
 
-```{r, eval=has_both, fig.align = "center", out.width = '100%'}
+```{r, eval=has_both}
+
+#| fig.width = 30,
+#| fig.height = 8,
+#| out.width = '100%',
+#| dpi = 96
+
 p1 <- kitsap +
@@
   theme_void() +
-  labs(title = paste("Map for","{{current_geo}}"),
-       caption = "Data source: Kitsap GIS - data in development",
+  labs(title = "{{current_geo}}",#paste("Map for","{{current_geo}}"),
        color = "Pedestrian Facilities")
@@
        color = "Pedestrian Facilities") +
-  theme_void()
+  theme_void() +
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
        color = "Pedestrian Facilities") +
-  theme_void()
+  theme_void() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# calculate the number of rows needed to display the legend
+this_rows <- max(4,ceiling(n_distinct(as.character(this_sw_sh$TYPE))/3))
 
@@
 
-(p1 + 
+((p1 + 
   plot_layout(guides = 'collect') +
-  labs(title = paste("Map for","{{current_geo}}"),
-       caption = "Data source: Kitsap GIS - data in development",
-       color = "Pedestrian Facilities") & 
-    theme(legend.position = "bottom") ) +
+  labs(color = "Pedestrian Facilities") & 
+    theme(legend.position = "bottom") &
+    guides(color=guide_legend(nrow=this_rows,byrow=TRUE))) +
   (l + theme(legend.position = "none") ) +
-  (r + theme(legend.position = "none") )
+  (r + theme(legend.position = "none") ) )
 
@@
 this_comb_total_score <- sum(this_comb_score2$w_score)/sum(this_comb_score2$the_length)
+
+# do any with sw sh have score 0?
+has0 <- any(this_comb_score$score == 0)
+# do any with sw sh have score 10? 
+has10 <- any(this_comb_score$score == 10)
+this0 <- ifelse(has0, "A few segments where sidewalks or shoulders were matched to a road segment, but the resulting score was 0 are shown here. ","")
+this10 <- ifelse(has10, "A few good (score 10) segments are shown here. ","")
 ```
@@
 
-A few segments with a score of 0 are shown here. 
-```{r eval=has_both}
+`r this0`
+
+```{r eval=has_both & has0}
 
@@
 
+`r this10`
+
+```{r eval=has_both & has10}
+
+roads_10 <- this_comb_score %>%
+  filter(score == 10) %>%
+  select(setting, road_class, FULL_NAME_R, 
+         sidewalk_right, sidewalk_left,
+         shoulder_right, shoulder_left,
+         standard_sidewalk, standard_shoulder)
+
+knitr::kable(head(roads_10) %>%
+               rename(Setting = setting,
+                      Class = road_class,
+                      Name = FULL_NAME_R,
+                      `sidewalk R` = sidewalk_right,
+                      `sidewalk L` = sidewalk_left,
+                      `shoulder R` = shoulder_right,
+                      `shoulder L` = shoulder_left,
+                      `standard sidewalk` = standard_sidewalk,
+                      `standard shoulder` = standard_shoulder)) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+
+```
 ```{r}

### 2022-10-10 child_region.Rmd
shift the regions to loops and nest child

[comment]: # creates a section for a smaller geo passed with details
[comment]: # loops over a number i referencing the geo in breakdowns tibble
[comment]: # called from Draft-Ped-Metrics-Example.Rmd

```{r bdsetup-{{i}}}
# create intro text
i <- as.integer({{i}})
has_points <- !is.na(breakdowns$pointdf[i])

b_name <- breakdowns$Name[i]
if(has_points){ b_pointdf <- eval(as.symbol(breakdowns$pointdf[i])) }
b_polydf <- eval(as.symbol(breakdowns$polydf[i]))
b_items <- eval(as.symbol(breakdowns$items[i]))
b_idcol <- breakdowns$idcolpoly[i]

loop_df <- st_as_sf(eval(b_polydf)) %>% mutate(tomatch = eval(as.symbol(b_idcol)))
score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = b_items, Score = NA))
```

### `r b_name`

```{r bdtext-{{i}}}
intro_text <- paste0("Kitsap County has ",
                    ifelse(!has_points, # poly only 
                           paste(nrow(b_polydf), 
                                 b_name)
                           , # now the ones with points
                           paste0(nrow(b_pointdf), 
                                  " ",
                                  b_name, 
                                  "; however, only ",
                                  nrow(b_polydf),
                                  " are outside of city limits. Here they are shown with a 30 minute walking distance")
                           ),".")

title_text <- paste("Kitsap County",b_name)
subtitle_text <- ifelse(b_name %ni% c("Commissioner Districts","LAMIRDs"),
                        "only showing those in the unincorporated county","")
caption_text <- paste("Data source:",breakdowns$data_source[i])

```

`r intro_text`

```{r bdmap-{{i}}}
#| fig.width = 8,
#| fig.height = 8,
#| out.width = '100%',
#| dpi = 96
#| 
kitsap +
  geom_sf(data = b_polydf, aes(fill = eval(as.symbol(b_idcol))), 
          lwd = 0, alpha = 0.6) +
  {if (has_points) geom_sf(data = b_pointdf, aes(color = 1), 
                           color = "black", show.legend = FALSE) } +
  theme_void() +
  labs(title = title_text,
       subtitle = subtitle_text,
       caption = caption_text,
       fill = b_name)

```

```{r bdloop-{{i}}, include=FALSE, echo=FALSE, cache = TRUE}

expanded_child <- lapply(
    b_items,
    function(x) {
      knitr::knit_expand(file = "child_single_geo_in_region.Rmd",
                                current_geo = x )
    })

parsed_child <- knitr::knit_child(text = unlist(expanded_child))
```

`r parsed_child`

### 2022-10-10 child_single_geo_in_region.Rmd
shift the regions to loops and nest child

--- 2022-10-09 child_single_geo.Rmd
@@
 ```{r, results = 'asis', eval=!has_roads}
-cat("This area does not have either known roads.")
+cat("This area does not have known roads.")
 ```
@@
 # identified a mislabeled RD_LOG_ID in the sidewalk data for Madrone (as Main)
-x <- st_as_sf(this_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
+x <- st_as_sf(this_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 500)
 x$RD_LOG_ID <- 43809
@@
                    "")
-
-# ggplot() +
-#   geom_sf(data = this_geo, aes(fill = 1), alpha = 0.2, lwd = 0, show.legend = FALSE) + 
-#   geom_sf(data = st_buffer(bind_rows(this_sw_sh, x), 50, endCapStyle="FLAT"),
-#           fill = "#4E22EE", alpha = 0.5) +
-#   geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) + 
-#   geom_sf(data = this_test2, color = "black") +
-#   geom_sf(data = this_sw_sh, aes(color = as.factor(RD_LOG_ID)), lwd = 0)
-
 ```
@@
 
-```{r, eval=has_both}
- # should we test for points?
-the_points <- c()
-if (exists("the_schools")) { the_points <- append(the_points, "the_schools")} 
-if (exists("the_libraries")) { the_points <- append(the_points, "the_libraries")} 
-if (exists("the_transits")) { the_points <- append(the_points, "the_transits")} 
-check_points <- ifelse(length(the_points) > 0, T, F)
-has_point = FALSE # start with false
-```
-
-```{r, eval=has_both & check_points}
+```{r, eval=has_both & has_points}
   # see if this geo has a point (is it an area around a place, like a school)
-
-if (length(which(eval(as.symbol(the_points)) == this_geo$tomatch)) !=0) {
-  # we need to figure out which one
-  has_point = TRUE
-  tr <- which(eval(as.symbol(the_points)) == this_geo$tomatch)
-  if (tr <= length(the_schools)) { # it is a school
-    this_point <- st_as_sf(pub_schools) %>%
+this_point <- st_as_sf(b_pointdf) %>%
       filter(NAME == this_geo$tomatch)
-  } else if (tr > length(the_schools) & 
-             tr <= length(the_schools) + length(the_libraries)) { # library
-    this_point <- st_as_sf(libraries) %>%
-      filter(NAME == this_geo$tomatch)
-  } else { #transit
-    this_point <- st_as_sf(transit_cp) %>%
-      filter(NAME == this_geo$tomatch)
-  }
-}
-
 ```
@@
 p1 <- kitsap +
-  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.5, show.legend = FALSE) +
-  geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
+  geom_sf(data = this_geo, aes(fill = 1), lwd = 0, alpha = 0.8, show.legend = FALSE) +
+  #geom_sf(data = this_roads, aes(color = 1), color = "#999999", alpha = 0.5) +
   geom_sf(data = this_sw_sh, aes(color = as.factor(TYPE)), show.legend = "line") + 
@@
   labs(title = "{{current_geo}}",#paste("Map for","{{current_geo}}"),
-       color = "Pedestrian Facilities")
+       color = "")
 
@@
   geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "RIGHT"), aes(color = as.factor(TYPE))) +
-  {if(has_point) geom_sf(data = this_point, 
+  {if(has_points) geom_sf(data = this_point, 
                          aes(color = 1), color = "black", show.legend = FALSE)} +
@@
   labs(title = "Right",
-       color = "Pedestrian Facilities") +
+       color = "") +
   theme_void() +
@@
   geom_sf(data = filter(this_sw_sh, SIDE_OF_RO == "LEFT"), aes(color = as.factor(TYPE))) +
-  {if(has_point) geom_sf(data = this_point, 
+  {if(has_points) geom_sf(data = this_point, 
                          aes(color = 1), color = "black", show.legend = FALSE)} +
@@
   labs(title = "Left",
-       color = "Pedestrian Facilities") +
+       color = "") +
   theme_void() +
@@
   plot_layout(guides = 'collect') +
-  labs(color = "Pedestrian Facilities") & 
+  labs(color = "") & 
     theme(legend.position = "bottom") &

### 2022-10-10 Draft-Ped-Metrics-Example.Rmd
handle shifted children and clean up score tables

@@
 library(kableExtra)
-
+library(tidyverse)
+library(sf)
 ```
@@
 kitsap <- ggplot() + 
-  geom_sf(data = outline["plain"]) + 
+  geom_sf(data = outline["plain"], alpha = 0.5) + 
   theme_void()
@@
 
-# Purpose
-
-Kitsap County Pedestrian Facilities Metrics/Evaluation Report will:
-
-1. Provide a complete evaluation of pedestrian facilities across the county down to the segment and side of road level
-2. Present an aggregated quality rating for pedestrian facilities in specific areas
-3. Target missing facilities critical to connecting communities or to connecting residential to local services, recreation, and shops
-4. Show measure of progress (over time) of pedestrian facilities improvements (_starting with the second report_)
-
-## This Draft
-
-This draft example was prepared by the Kitsap County Non-Motorized Community Advisory Committee Special Committee on Pedestrian Metrics.
-It is NOT considered an authoritative source of these data and is NOT an official report of the county.
-
-# Introduction
-
-This report provides a complete evaluation of pedestrian facilities in Kitsap County comparing the actual facilities to a minimum adequate pedestrian facility level.
-The minimum adequate pedestrian facility level is either a paved shoulder or a sidewalk of a certain minimum width - depending on the character of the road.
-
-The intended audience is the people of Kitsap County and law/decision makers determining how to target limited funds for transportation development.
-
-Not all road segments will be evaluated. This report excludes road segments that are:
-
- - Inside of cities (Bremerton, Port Orchard, etc.)^[Draft excludes segments that intersect with the Kitsap County incorporated city limits polygons[@kcgis_cities]]
- - Local easements^[Draft assuming Function Class 100 or 110 in Road Centerline Data]
- - Highways^[Draft Assuming Function Class 2 in Road Centerline Data]
- - Speed limit > 50 MPH^[**these are not currently identified because the road centerline data does not include speed limit**]
-
-```{r kitsapmap, cache = TRUE}
-
-#| fig.align = "center"
-#| fig.width = 12
-#| fig.height = 8
-#| out.width = '100%'
-
-library(tidyverse)
-library(sf)
-
-kitsap_all_roads <- kitsap +
-  geom_sf(data = clean_roadcl, aes(color = road_class)) + 
-  road_col +
-  theme_void() +
-  labs(title = "Roads in Kitsap County",
-       subtitle = paste0(nice_comma(sum(st_length(clean_roadcl))/5280),
-                         " centerline miles"),
-       #caption = "Data source: Kitsap GIS",
-       color = "")
-
-kitsap_roads_no_cities <- kitsap + 
-  geom_sf(data = cities["NAME"], 
-          aes(fill="NAME"),
-          fill="#808080",
-          alpha = 0.5,
-          show.legend = FALSE) +
-  geom_sf(data = ncdf, aes(color = road_class), show.legend = FALSE) + 
-  road_col +
-  theme_void() +
-  labs(title = "No Cities, Easements, & Highways",
-       subtitle = paste0(nice_comma(sum(st_length(ncdf))/5280),
-                         " centerline miles"),
-       caption = "Data source: Kitsap GIS")
-
-library(patchwork)
-
-kitsap_all_roads + kitsap_roads_no_cities +
-  plot_layout(guides = 'collect') & theme(legend.position = 'bottom')
-
-detach("package:patchwork", unload=TRUE)
-```
-
-# Data and Methods
-
-## Data {#datasection}
-
-Kitsap County maintains data on roads, sidewalks, and shoulders.
-These data all use the Kitsap County Public Works Road Log Id to identify road segments.
-Combined, these data allow for assessment of the quality of pedestrian facilities on Kitsap County roads.
-In addition, Kitsap County maintains data on the geographic representations of the cities, UGAs, LAMIRDs, and some key points of interest (like schools and parks) within the County.[@kcgis]
-
-These are the roads that will be evaluated.
-```{r}
-library(tidyverse)
-library(sf)
-
-groupcl <- ncdf %>%  
-  st_set_geometry(., NULL) %>% 
-  group_by(road_class) %>% 
-  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
-            segments = nice_comma(n(),0), 
-            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
-            .groups = "drop") 
-df <- ncdf %>% #clean_roadcl %>%  
-  st_set_geometry(., NULL) %>% 
-  group_by(road_class = "Total") %>% 
-  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
-            segments = nice_comma(n(),0), 
-            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
-            .groups = "drop") 
-groupcl <- bind_rows(groupcl, df)
-
-colnames(groupcl) <- c("Road Classification", 
-                       "Count of Roads", "Count of Segments", 
-                       "Length in Miles") 
-knitr::kable(groupcl, 
-             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")  %>%
-  kable_material(c("striped", "hover"))
-```
-
-These are the sidewalk and shoulders also outside of Cities.
-Single sides shown.
-This is one example where we know the data are incomplete; for example, Silverdale Way is not in a city and is known to have sidewalks on both sides.^[For the specific example of Silverdale Way, there are three entries in the sidewalk data with the Road Log Id of 19515 that matches Silverdale Way, and they all indicate sidewalks on the left side of the road.
-Further note, all three have a segment ID of 0, so they will not match any of the segments in the roads data if we match on Road Log Id and Segment ID - an entirely different problem with the data that is not isolated to Silverdale Way.
-`r round(sum(nc_sw_sh$SEGMENT_ID ==0)/nrow(nc_sw_sh)*100,1)`% of the sidewalk data do not have segment ids.] For this draft example report it is okay; if the initial report is completed with the incomplete data, it will present a poorer overall condition of pedestrian facilities than is in place.
-
-```{r}
-df <- sidewalks_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-df2 <- sidewalks_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-
-df <- bind_rows(df, df2)
-
-knitr::kable(df, 
-             caption = "Miles of Sidewalks by Width of Sidewalk (single side)") %>%
-  kable_material(c("striped", "hover"))
-```
-
-```{r}
-df <- shoulders_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-df2 <- shoulders_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-
-df <- bind_rows(df, df2)
-
-knitr::kable(df, 
-             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)") %>%
-  kable_material(c("striped", "hover"))
-```
-
-```{r}
-# tmp_sw_sh %>%
-#   st_set_geometry(., NULL) %>%
-#   group_by(RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, TYPE) %>% # road_type) %>%
-#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280) %>%
-#   ungroup() %>%
-#   pivot_wider(names_from = SIDE_OF_RO, values_from = "TYPE")
-#%>%
-  #mutate(both = !is.na(RIGHT) & !is.na(LEFT)) %>%
-  #group_by(both) %>%
-  #summarise(n = n())
-
-# tmp_sw_ncdf %>%  
-#   st_set_geometry(., NULL) %>%
-#   filter(LENGTH > 0) %>%
-#   group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
-#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
-# 
-# tmp_sh_ncdf %>%  
-#   st_set_geometry(., NULL) %>%
-#   filter(LENGTH > 0) %>%
-#   group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
-#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
-```
-
-### Combining the Data
-
-In order to do this evaluation, we must combine the roads data (which has information on the class and setting of roads) with the sidewalk and shoulder data (with information about the pedestrian facilities).
-However, the most obvious methods of combination fail.
-
-#### Matching on Road Log Id is good but insufficient
-
-All of the data sets appear to use the same set of road log IDs from public works.
-Unfortunately, knowing that there is a shoulder or sidewalk somewhere on a road does not get to how much of that road has the sidewalk or shoulder.^[As roads sometimes change names and road log IDs, there are also some breaks between the road data and sidewalk/shoulder data.
-This was noted for Madrone Ave in Manchester - given the same road log ID as Main St when the road bends.]
-
-This diagnostic and ugly plot is intended to show the coverage of road log id across the sidewalk and shoulder data sets compared to the raod data.
-While the sidewalk and shoulder data are more sparse, the range is consistent.
-
-```{r}
-sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
-                  aes(x = as.numeric(row.names(nc_sw_sh)), y = RD_LOG_ID)) + 
-  geom_point() +
-  labs(title = "Sidewalk and Shoulder Data",
-       x = "") +
-  expand_limits(y = c(0,100000))
-road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
-                  aes(x = as.numeric(row.names(ncdf)), y = RD_LOG_ID)) + 
-  geom_point() +
-  labs(title = "Roads Data",
-       x = "")
-
-library(patchwork)
-
-sw_plot + road_plot
-
-detach("package:patchwork", unload=TRUE)
-rm(sw_plot)
-rm(road_plot)
-```
-
-```{r}
-match_road_id <- st_set_geometry(sidewalks,NULL) %>%
-  mutate(RD_LOG_ID = as.integer(ROADLOGID)) %>%
-  left_join(select(st_set_geometry(clean_roadcl,NULL),RD_LOG_ID,FULL_NAME)) %>%
-  group_by(RD_LOG_ID, ROADNAME, SIDE_OF_RO, WIDTH) %>%
-  summarise(n = sum(!is.na(FULL_NAME)), .groups = "drop") %>% 
-  filter(n == 0, WIDTH > 0)
-```
-
-The most obvious smaller match is the segment id. However, 
-**NOTE: We cannot actually use the SEGMENT ID in this draft report because it does not have the same meaning in the sidewalk and shoulder data as it has in the road data.
-The following plots are diagnostic but show that the sidewalks and shoulders data do not have segment ids covering the same spectrum as the road data.
-
-```{r}
-sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
-                  aes(x = as.numeric(row.names(nc_sw_sh)), y = SEGMENT_ID)) + 
-  geom_point() +
-  labs(title = "Sidewalk and Shoulder Data",
-       x = "")+
-  expand_limits(y = c(0,20000))
-road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
-                  aes(x = as.numeric(row.names(ncdf)), y = SEGMENT_ID)) + 
-  geom_point() +
-  labs(title = "Roads Data",
-       x = "")
-
-library(patchwork)
-
-sw_plot + road_plot
-
-detach("package:patchwork", unload=TRUE)
-rm(sw_plot)
-rm(road_plot)
-```
-
-The segment ids are clearly not from the same space, but the road log ids are consistent.^[Mostly.] However, we cannot combine on road log id alone - because the roads change character and the type of pedestrian facilities over space.
-We combine instead on geometry.
-
-For now, we are making a space around a sidewalk or shoulder geometry, a buffer, to capture the nearby area where the road would be, and seeing if there are any road segments that intersect.
-If the road segment intersects and has the same road log id, we keep it.
-
-```{r combine-rd-sw-sh}
-# combine the sidewalk and shoulder and roads data
-x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
-x$RD_LOG_ID <- 43809
-comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
-                                     SEGMENT_ID_R = SEGMENT_ID,
-                                     FULL_NAME_R = FULL_NAME, 
-                                     LENGTH_R = LENGTH,
-                                     road_class, RUCODE), 
-                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
-  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
-  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
-  mutate(distance = as.integer(st_length(.)))
-
-```
-
-## Method of Evaluation
-
-Road segments are the basis for evaluation because they are the smallest unit in common across the data and can be grouped together to form trips.
-Actually matching up the data for roads, shoulders, and sidewalks will prove to be a more complicated task than it appears at first blush.
-As noted in the [Data section](#datasection), the Segment IDs in the shoulder and sidewalk data do not cover the range of segments as in the road data.
-Here, we join based on the road log id and beginning and ending mileage point
-
-Each road segment is evaluated individually.
-For each road segment, its adequate pedestrian facility level is based on its classification and the Kitsap County Road Standards Table 3.3 and 3.4.[@kc_road_standards]^[revisions of road standards should be followed by review and reevaluation of these criteria]
-
-```{r tbl_standards}
-# Road Standards Tables 3.3 and 3.4
-library(tidyverse)
-library(kableExtra)
-road_standards_given <- tribble(
-  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, 
-  "Urban", "Local Road", NA, 5, NA,
-  "Urban", "Local Sub-collector", NA, 5, NA,
-  "Urban", "Arterial", NA, 6, NA,
-  "Urban", "Collector", NA, 6, NA,
-  "Rural", "Local Road", NA, NA, 3,
-  "Rural", "Local Sub-collector", NA, NA, 4,
-  "Rural", "Collector", "ADT 400-750", NA, 3,
-  "Rural", "Collector", "ADT 751-1000", NA, 4,
-  "Rural", "Collector", "DHV 100-200", NA, 6,
-  "Rural", "Collector", "DHV > 200", NA, 8,
-  "Rural", "Minor Arterial", "DHV < 100", NA, 4,
-  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,
-  "Rural", "Minor Arterial", "DHV > 200", NA, 8,
-  "Rural", "Principal Arterial", "DHV <200", NA, 6,
-  "Rural", "Principal Arterial", "DHV > 200", NA, 8,
-)
-
-knitr::kable(road_standards_given,
-             caption = "Summary of Tables 3.3 and 3.4 in Kitsap County Road Standards [@kc_road_standards]")  %>%
-   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT.
-Note: sub-collector, ADT, and DHV are not defined in the publicly available road data.") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-```
-
-Because of all of the items that are not available in the road data at this time, this draft makes some assumptions to allow for an evaluation.
-When the information is known, a more accurate evaluation can be made.^[Roads with unknown setting - RUCODE is 0, we match them to the Urban road standards.
-This may be incorrect.]
-
-```{r tbl_standards_simple}
-# Road Standards Tables 3.3 and 3.4 translated for data we have
-library(tidyverse)
-library(kableExtra)
-road_standards <- tribble(
-  ~setting, ~class, ~sidewalk, ~shoulder,
-  "Urban", "Local Access", 5, NA,
-  "Urban", "Minor Collector",6, NA,
-  "Urban", "Major Collector",6, NA,
-  "Urban", "Minor Arterial", 6, NA,
-  "Urban", "Principal Arterial", 6, NA,
-  "Rural", "Local Access", NA, 3,
-  "Rural", "Minor Collector",NA, 3,
-  "Rural", "Major Collector",NA, 4,
-  "Rural", "Minor Arterial", NA, 4,
-  "Rural", "Principal Arterial", NA, 6,
-)
-
-knitr::kable(road_standards,
-             caption = "Simplified Assumptions to Allow for Evaluation")  %>%
-   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-```
-
-Road segments that are explicitly excluded from the evaluation will be scored NA and not included in a summary of pedestrian metrics in an area (they will neither improve nor pull down the summary score).
-
-Pedestrian facilities on road segments will be scored based on the sum of the score of the sides. A side of the road for a segment will be scored:
-
-- 5 if it has a sidewalk or shoulder of adequate size (based on road classification & standards) or a sidewalk is present where a shoulder is the standard
-- 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
-- 1 if the sidewalk or shoulder is present and is narrower than AASHTO (**NOTE to subcommittee - do we think we should use a 1 if there is a shoulder when there should be a sidewalk?**)
-- 0 if none present or no data
-
-Therefore, segments with sidewalk or shoulder of adequate size (based on road classification) on both sides of road would receive a 10.
-Segments with no data or with no sidewalk or shoulder would receive a 0.
-
-Any given area score will be a weighted average of the scores of the segments within the area. Weighting is by centerline linear feet per segment. A simple example is an area composed of 4 segments.
-
-```{r}
-example_tbl <- tribble(
-  ~Segment, ~"Segment Score", ~"Segment Centerline Linear Feet",
-  "A", 6, 100,
-  "B", 0, 10,
-  "C",10,50,
-  "D",8,200,
-  "Total",7.5 ,360,
-)
-
-knitr::kable(example_tbl) %>%
-kable_styling(bootstrap_options = "striped", full_width = F, position = "float_left")
-```
-
-The Total Score would be the segment scores multiplied by their linear feet and then divided by the total linear feet of the area (360).
-For this example, we get a area score of 7.5 $$ 6 x 100 + 0 x 10 + 10 x 50 + 8 x 200 = 2700$$ $$2700/360 = 7.5$$
-
-```{r}
-score_interpretation <- tribble(
-  ~score_at_least, ~interpretation,
-  10, "Good",
-  5, "Adequate",
-  3, "Minimum",
-  1, "Poor",
-  0.001, "Very Poor",
-  0, "Nothing"
-)
-
-getScoreInterpretation <- function(x){
-  first(score_interpretation$interpretation[
-        score_interpretation$score_at_least <= x]
-    )
-}
-
-knitr::kable(score_interpretation,
-             caption = "Interpretation of Scores for Segments and for Areas")  %>%
-   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-```
-
-# Results
-First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.
-
-## Overall
-
-```{r compare_standards}
-# take the road cl data for the roads under evaluation
-# remove geometry
-# merge with standards to get requirement based on setting (RUCODE) and class (road_class)
-# merge with sidewalk and shoulder data to get what exists
-
-library(tidyverse)
-library(sf)
-
-road_standards <- road_standards %>%
-  rename(standard_sidewalk = sidewalk, 
-         standard_shoulder = shoulder, 
-         road_class = class)
-
-getSideScore <- function(standard_sidewalk, standard_shoulder,
-                         sidewalk, shoulder) {
-  ifelse(is.na(standard_sidewalk), # dealing with shoulder
-         case_when(
-           !is.na(sidewalk) ~ 5, # there is a sidewalk present when a shoulder is the standard
-           shoulder >= standard_shoulder ~ 5,
-           shoulder < standard_shoulder & shoulder > 4 ~ 3,
-           shoulder < standard_shoulder & shoulder > 0 ~ 1,
-           TRUE ~ 0
-         ),
-         case_when(
-           sidewalk >= standard_sidewalk ~ 5,
-           sidewalk < standard_sidewalk & sidewalk > 4 ~ 3,
-           sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
-           TRUE ~ 0
-         )
-  )
-}
-
-comb_stand <- comb %>% 
-  st_set_geometry(., NULL) %>% 
-  mutate(setting = ifelse(RUCODE == 1, "Rural", 
-                          ifelse(RUCODE == 2, "Urban", "Unknown")),
-         ped_type = sub("\\:.*", "", TYPE),
-         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
-  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), 
-              names_from = nn, 
-              values_from = WIDTH,
-              values_fn = mean) %>%
-  left_join(road_standards)
-
-comb_score <- comb_stand %>%
-  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
-                                   sidewalk_left, shoulder_left),
-         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
-                                   sidewalk_right, shoulder_right),
-         score = score_left + score_right)
-
-```
-
-```{r}
-comb_score2 <- st_set_geometry(ncdf, NULL) %>% 
-  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
-         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
-  mutate(setting = ifelse(RUCODE == 1, "Rural", 
-                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
-  left_join(comb_score) %>%
-  mutate(the_score = ifelse(is.na(score), 0, score),
-         the_length = ifelse(is.na(distance), LENGTH, distance),
-         w_score = the_score*the_length)
-
-comb_total_score <- sum(comb_score2$w_score)/sum(comb_score2$the_length)
-
-comb_score_by_setting <- comb_score2 %>%
-  group_by(setting) %>%
-  summarise(score = sum(w_score)/sum(the_length),
-            .groups = "drop") %>%
-  rowwise() %>%
-  mutate(interpretation = getScoreInterpretation(score))
-
-comb_score_by_setting_class <- comb_score2 %>%
-  group_by(setting, road_class) %>%
-  summarise(score = sum(w_score)/sum(the_length),
-            .groups = "drop") %>%
-  rowwise() %>%
-  mutate(interpretation = getScoreInterpretation(score))
-```
-
-The overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, `r getScoreInterpretation(comb_total_score)`.
-
-Looking by the same breakdown that we have for standards, we can see where the scores are highest and lowest.
-
-```{r}
-knitr::kable(comb_score_by_setting_class) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-```
-
-```{r}
-# More detailed information
-# knitr::kable(
-#   comb_score2 %>%
-#   group_by(setting, road_class, score) %>% 
-#   summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
-#             segments = nice_comma(n()), 
-#             #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
-#             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
-#             .groups = "drop") 
-#   )
-```
-
-```{r}
-ggplot(filter(comb_score2, setting != "Unknown"), 
-       aes(x = as.factor(score), y = the_length/5280))+
-  geom_col() +
-  labs(x = "score based on data available", 
-       y = "centerline miles of road") +
-  theme_minimal() +
-  facet_grid(setting ~ road_class)
-```
-
-If we remove the segments where we do not have information about sidewalks and shoulders at all, we can see that for those, we are generally finding higher scores for rural roads - with paved shoulders.
-
-```{r}
-ggplot(filter(comb_score2, setting != "Unknown", !is.na(score)), 
-       aes(x = as.factor(score), y = LENGTH/5280))+
-  geom_col() +
-  labs(x = "score based on data available", 
-       y = "centerline miles of road") +
-  theme_minimal() +
-  facet_grid(setting ~ road_class)
-```
-
-Some of the segments with score of 0 are because they have shoulders when they should have sidewalks.
-If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.
-Here's a few example segments.
-
-```{r}
-
-roads_0 <- comb_score2 %>%
-  filter(score == 0) %>%
-  select(setting, road_class, FULL_NAME_R, sidewalk_right:standard_sidewalk)
-
-knitr::kable(head(roads_0) %>%
-               rename(Setting = setting,
-                      Class = road_class,
-                      Name = FULL_NAME_R,
-                      `sidewalk R` = sidewalk_right,
-                      `sidewalk L` = sidewalk_left,
-                      `shoulder R` = shoulder_right,
-                      `shoulder L` = shoulder_left,
-                      `standard sidewalk` = standard_sidewalk)) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-
-```
-
-Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road.
-A few details for roads where at least one segment has a 10 are shown here.
-```{r}
-roads_10 <- comb_score2 %>%
-  filter(score == 10) %>%
-  group_by(setting, road_class, FULL_NAME_R) %>%
-  summarise(segments = nice_comma(n(),0), 
-            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
-            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
-            .groups = "drop") %>%
-  rename(Setting = setting,
-         Class = road_class,
-         Name = FULL_NAME_R,
-         `Segments w 10` = segments,
-         `Segment Length` = road_len,
-         `Ped Fac Length` = dist)
-
-knitr::kable(roads_10) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-```
-
-Some road segments are missing entirely from the centerline road data - for example, the portion of Rigetop that goes from the end of Myhre and curves back to Silverdale Way.
-These missing road segments are mostly complete streets with wide sidewalks.
-Here, showing a map with all roads with the name *RIDGETOP*, you can only see the part going from Silverdale Way to Myhre.
-These segments, and any others like them, missing from the data decrease the overall pedestrian facilities score.
-
-```{r}
-kitsap +
-  geom_sf(data = filter(clean_roadcl, STREETNAME == "RIDGETOP"), aes(color = road_class)) + 
-  road_col +
-  theme_void() +
-  labs(title = "STREETNAME == RIDGETOP",
-       color = "")
-```
-
-## Smaller Breakdowns of the County
-
-We can also look at regions within the County.
-For these, we identify the roads and any sidewalks or shoulders.^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]
-
-Regions are already defined by an area.
-For pedestrian generators, we identify an area that is within 30 minute walking distance of the facility and evaluate roads therein.^[30 minute walking distance calculated using the mapboxapi function `mb_isochrone`.
-Isochrones use map data from Mapbox and OpenStreetMap and their data sources.
-To learn more, visit https://www.mapbox.com/about/maps/ and http://www.openstreetmap.org/copyright.[@mapbox]]
-
-These regions and pedestrian generators are:
- * Commissioner Districts
- * UGAs
- * LAMIRDs
- * Schools
- * Libraries
- * Transit Centers
-
-```{r}
-score_breakdowns <- tibble(Area = "Kitsap County", Score = comb_total_score)
-```
-
-### Commissioner Districts
-
-```{r plot_commdists}
-commdists <- sf::read_sf(paste0(this_dir,"data/commdist"))
-commdists <- st_intersection(commdists, st_buffer(county_no_cities,0)) 
-commdists <- st_as_sf(commdists) %>% 
-  mutate(NAME = ifelse(DISTRICT == "1", "1 North",
-                ifelse(DISTRICT == "2", "2 South", "3 Central")))
-
-kitsap +
-  geom_sf(data = commdists, aes(fill = NAME), alpha = 0.5) +
-  theme_void() +
-  labs(title = 'Kitsap County Commissioner Districts',
-       caption = "Data source: Kitsap GIS - data in development",
-       color = "DIMS")
-
-```
-
-```{r loop_commdists,include=FALSE,echo=FALSE}
-the_commdists <- unique(st_set_geometry(commdists, NULL)$NAME)
-loop_df <- st_as_sf(commdists) %>% mutate(tomatch = NAME)
-
-score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_commdists, Score = NA))
-
-expanded_child <- lapply(
-    the_commdists,
-    function(x) {
-      knitr::knit_expand(file = "child_single_geo.Rmd",
-                                   current_geo = x )
-
-    }
-    )
-
-parsed_child <- knitr::knit_child(text = unlist(expanded_child))
-
-```
-
-Kitsap County has `r length(the_commdists)` Commissioner Districts.
-
-`r parsed_child`
-
-### UGAs
-
-```{r plot_ugas}
-uga_no_city <- st_as_sf(uga) %>%
-  filter(substring(GMA_JURISD, 1, 4) != "City")
-kitsap +
-  geom_sf(data = uga_no_city, aes(fill = GMA_JURISD), alpha = 0.5) +
-  theme_void() +
-  labs(title = 'Kitsap County UGAs',
-       caption = "Data source: Kitsap GIS - data in development",
-       color = "UGAS")
-
-```
-
-```{r loop_ugas,include=FALSE,echo=FALSE}
-the_ugas <- unique(st_set_geometry(uga_no_city, NULL)$GMA_JURISD)
-loop_df <- st_as_sf(uga_no_city) %>% mutate(tomatch = GMA_JURISD)
-
-score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_ugas, Score = NA))
-
-expanded_child <- lapply(
-    the_ugas,
-    function(x) {
-      knitr::knit_expand(file = "child_single_geo.Rmd",
-                                   current_geo = x )
-
-    }
-    )
-
-parsed_child <- knitr::knit_child(text = unlist(expanded_child))
-
-```
-
-Kitsap County has `r length(the_ugas)` UGAs - Urban Growth Areas that are outside of cities.
-Within the Urban Growth Areas, most design and planning standards are similar to what would be expected in an urban environment for new development.
-
-`r parsed_child`
-
-### LAMIRDs
-
-```{r plot_lamirds}
-kitsap +
-  #geom_sf(data = nc_sw_sh, aes(color = as.factor(TYPE))) + 
-  geom_sf(data = lamird, aes(fill = PLAN_AREA), alpha = 0.5) +
-  theme_void() +
-  labs(title = 'Kitsap County LAMIRDS',
-       caption = "Data source: Kitsap GIS - data in development",
-       color = "LAMIRDS")
-
-```
-
-```{r loop_lamirds,include=FALSE,echo=FALSE}
-the_lamirds <- unique(st_set_geometry(lamird, NULL)$PLAN_AREA)
-loop_df <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)
-
-score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_lamirds, Score = NA))
-
-expanded_child <- lapply(
-    the_lamirds,
-    function(x) {
-      knitr::knit_expand(file = "child_single_geo.Rmd",
-                                   current_geo = x )
-
-    }
-    )
-
-parsed_child <- knitr::knit_child(text = unlist(expanded_child))
-
-```
-Kitsap County has `r length(the_lamirds)` LAMIRDS - Local Area of More Intense Rural Development.
-
-`r parsed_child`
-
-```{r}
+```{r gwfunction}
 getWalkingPolygonInCounty <- function(locdf, idcol) {
@@
 
-### Schools
-```{r}
+```{r smallerdata}
+# this section sets up the data that will be used for subdivisions of the county
+# get any data not set up yet
+commdists <- sf::read_sf(paste0(this_dir,"data/commdist"))
+commdists <- st_intersection(commdists, st_buffer(county_no_cities,0)) 
+commdists <- st_as_sf(commdists) %>% 
+  mutate(NAME = ifelse(DISTRICT == "1", "1 North",
+                ifelse(DISTRICT == "2", "2 South", "3 Central")))
+the_commdists <- unique(st_set_geometry(commdists, NULL)$NAME)
+loop_commdists <- st_as_sf(commdists) %>% mutate(tomatch = NAME)
+
+uga_no_city <- st_as_sf(uga) %>%
+  filter(substring(GMA_JURISD, 1, 4) != "City")
+the_ugas <- unique(st_set_geometry(uga_no_city, NULL)$GMA_JURISD)
+loop_ugas <- st_as_sf(uga_no_city) %>% mutate(tomatch = GMA_JURISD)
+
+the_lamirds <- unique(st_set_geometry(lamird, NULL)$PLAN_AREA)
+loop_lamirds <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)
+
 pub_schools <- schools %>%
   filter(TYPE == "PUBLIC")
-
 pub_schools2 <- getWalkingPolygonInCounty(pub_schools, "NAME")
-```
-
-Kitsap County has `r nrow(pub_schools)` Public schools; however, only `r nrow(pub_schools2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
-
-```{r plot_pubschools}
-#| fig.width = 8,
-#| fig.height = 8,
-#| out.width = '100%',
-#| dpi = 96
-
-kitsap +
-  geom_sf(data = pub_schools2, aes(fill = id), 
-          lwd = 0, alpha = 0.3) +
-  geom_sf(data = pub_schools, aes(color = 1), color = "black", show.legend = FALSE) +
-  theme_void() +
-  labs(title = 'Kitsap County Public Schools',
-       subtitle = "Schools not labeled with color are not in the unincorporated county",
-       caption = "Data source: Kitsap GIS - data in development\n
-                  Walking distance polygons from Mapbox",
-       fill = "Public Schools")
-
-```
-
-```{r loop_pubschools,include=FALSE,echo=FALSE, cache=TRUE}
 the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
-loop_df <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)
-
-score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_schools, Score = NA))
-
-expanded_child <- lapply(
-    the_schools,
-    function(x) {
-      knitr::knit_expand(file = "child_single_geo.Rmd",
-                                   current_geo = x )
-    })
-
-parsed_child <- knitr::knit_child(text = unlist(expanded_child))
-```
-
-`r parsed_child`
-
-### Libraries
-
-```{r}
+loop_schools <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)
+
 libraries <- community %>%
   filter(TYPE == "PUBLIC LIBRARY")
-
-lib2 <- getWalkingPolygonInCounty(libraries, "NAME")
-```
-
-Kitsap County has `r nrow(libraries)` Public libraries; however, only `r nrow(lib2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
-
-```{r plot_libraries}
-kitsap +
-  geom_sf(data = lib2, aes(fill = id), 
-          lwd = 0, alpha = 0.3) +
-  geom_sf(data = libraries, aes(color = 1), color = "black", show.legend = FALSE) +
-  theme_void() +
-  labs(title = 'Kitsap County libraries',
-       subtitle = "Libraries not labeled with color are not in the unincorporated county",
-       caption = "Data source: Kitsap GIS - data in development\n
-                  Walking distance polygons from Mapbox",
-       fill = "Libraries")
-
-```
-
-```{r loop_libraries,include=FALSE,echo=FALSE, cache = TRUE}
-the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
-loop_df <- st_as_sf(lib2) %>% mutate(tomatch = id)
-
-score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_libraries, Score = NA))
-
-expanded_child <- lapply(
-    the_libraries,
-    function(x) {
-      knitr::knit_expand(file = "child_single_geo.Rmd",
-                                   current_geo = x )
-    })
-
-parsed_child <- knitr::knit_child(text = unlist(expanded_child))
-```
-
-`r parsed_child`
-
-### Transit
-
-According to Kitsap Transit, "Transit Centers and Transfer Points are locations where Kitsap Transit buses and ferries meet up to allow the safe transfer of passengers from one route to another.
-"[@ktransit]^[Transit Center location data were manually copied from the Kitsap Transit Website because they are not one of the identified data sets on the Kitsap County GIS site.
-Future users will need to ensure that these addresses are correct.]
-
-```{r}
+libraries2 <- getWalkingPolygonInCounty(libraries, "NAME")
+the_libraries <- unique(st_set_geometry(libraries2, NULL)$id)
+loop_libraries <- st_as_sf(libraries2) %>% mutate(tomatch = id)
+
 transit_cp <- tribble(
@@
 # convert the lat lon into a point geometry
-transit_cp <- st_as_sf(transit_cp, 
+transits <- st_as_sf(transit_cp, 
                       coords = c(x = "lon", y = "lat"), 
                       crs = 4326) 
-transit_cp2 <- getWalkingPolygonInCounty(transit_cp, "NAME")
-```
-
-Kitsap County has `r nrow(transit_cp)` Public libraries; however, only `r nrow(transit_cp2)` are outside of city limits. Here they are shown with a 30 minute walking distance.
-
-```{r plot_transit}
+transits2 <- getWalkingPolygonInCounty(transits, "NAME")
+the_transits <- unique(st_set_geometry(transits2, NULL)$id)
+loop_transits <- st_as_sf(transits2) %>% mutate(tomatch = id)
+
+# combine data into tibble for this report
+#These regions and pedestrian generators are:
+kds <- "Kitsap County GIS"
+kt <- "Kitsap Transit Website"
+wd <- "Walking distance polygons from Mapbox"
+breakdowns <- tribble(
+  ~Name,~pointdf,~polydf,~idcolpoly,~items,~data_source,
+  "Commissioner Districts",NA,"commdists","NAME","the_commdists",kds,
+  "UGAs",NA,"uga_no_city","GMA_JURISD","the_ugas",kds,
+  "LAMIRDs",NA,"lamird","PLAN_AREA","the_lamirds",kds,
+  "Public Schools","pub_schools","pub_schools2","id","the_schools",paste(kds,wd),
+  "Public Libraries","libraries","libraries2","id","the_libraries",paste(kds,wd),
+  "Transit Centers","transits","transits2","id","the_transits",paste(kt,wd),
+  )
+
+```
+
+# Purpose
+
+Kitsap County Pedestrian Facilities Metrics/Evaluation Report will:
+
+1. Provide a complete evaluation of pedestrian facilities across the county down to the segment and side of road level
+2. Present an aggregated quality rating for pedestrian facilities in specific areas
+3. Target missing facilities critical to connecting communities or to connecting residential to local services, recreation, and shops
+4. Show measure of progress (over time) of pedestrian facilities improvements (_starting with the second report_)
+
+## This Draft
+
+This draft example was prepared by the Kitsap County Non-Motorized Community Advisory Committee Special Committee on Pedestrian Metrics.
+It is NOT considered an authoritative source of these data and is NOT an official report of the county.
+
+# Introduction
+
+This report provides a complete evaluation of pedestrian facilities in Kitsap County comparing the actual facilities to a minimum adequate pedestrian facility level.
+The minimum adequate pedestrian facility level is either a paved shoulder or a sidewalk of a certain minimum width - depending on the character of the road.
+
+The intended audience is the people of Kitsap County and law/decision makers determining how to target limited funds for transportation development.
+
+Not all road segments will be evaluated. This report excludes road segments that are:
+
+ - Inside of cities (Bremerton, Port Orchard, etc.)^[Draft excludes segments that intersect with the Kitsap County incorporated city limits polygons[@kcgis_cities]]
+ - Local easements^[Draft assuming Function Class 100 or 110 in Road Centerline Data]
+ - Highways^[Draft Assuming Function Class 2 in Road Centerline Data]
+ - Speed limit > 50 MPH^[**these are not currently identified because the road centerline data does not include speed limit**]
+
+```{r kitsapmap, cache = TRUE}
+
+#| fig.align = "center"
+#| fig.width = 12
+#| fig.height = 8
+#| out.width = '100%'
+
+library(tidyverse)
+library(sf)
+
+kitsap_all_roads <- kitsap +
+  geom_sf(data = clean_roadcl, aes(color = road_class)) + 
+  road_col +
+  theme_void() +
+  labs(title = "Roads in Kitsap County",
+       subtitle = paste0(nice_comma(sum(st_length(clean_roadcl))/5280),
+                         " centerline miles"),
+       #caption = "Data source: Kitsap GIS",
+       color = "")
+
+kitsap_roads_no_cities <- kitsap + 
+  geom_sf(data = cities["NAME"], 
+          aes(fill="NAME"),
+          fill="#808080",
+          alpha = 0.5,
+          show.legend = FALSE) +
+  geom_sf(data = ncdf, aes(color = road_class), show.legend = FALSE) + 
+  road_col +
+  theme_void() +
+  labs(title = "No Cities, Easements, & Highways",
+       subtitle = paste0(nice_comma(sum(st_length(ncdf))/5280),
+                         " centerline miles"),
+       caption = "Data source: Kitsap GIS")
+
+library(patchwork)
+
+kitsap_all_roads + kitsap_roads_no_cities +
+  plot_layout(guides = 'collect') & theme(legend.position = 'bottom')
+
+detach("package:patchwork", unload=TRUE)
+```
+
+# Data and Methods
+
+## Data {#datasection}
+
+Kitsap County maintains data on roads, sidewalks, and shoulders.
+These data all use the Kitsap County Public Works Road Log Id to identify road segments.
+Combined, these data allow for assessment of the quality of pedestrian facilities on Kitsap County roads.
+In addition, Kitsap County maintains data on the geographic representations of the cities, UGAs, LAMIRDs, and some key points of interest (like schools and parks) within the County.[@kcgis]
+
+These are the roads that will be evaluated.
+```{r}
+library(tidyverse)
+library(sf)
+
+groupcl <- ncdf %>%  
+  st_set_geometry(., NULL) %>% 
+  group_by(road_class) %>% 
+  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
+            segments = nice_comma(n(),0), 
+            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            .groups = "drop") 
+df <- ncdf %>% #clean_roadcl %>%  
+  st_set_geometry(., NULL) %>% 
+  group_by(road_class = "Total") %>% 
+  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
+            segments = nice_comma(n(),0), 
+            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            .groups = "drop") 
+groupcl <- bind_rows(groupcl, df)
+
+colnames(groupcl) <- c("Road Classification", 
+                       "Count of Roads", "Count of Segments", 
+                       "Length in Miles") 
+knitr::kable(groupcl, 
+             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")  %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+These are the sidewalk and shoulders also outside of Cities.
+Single sides shown.
+This is one example where we know the data are incomplete; for example, Silverdale Way is not in a city and is known to have sidewalks on both sides.^[For the specific example of Silverdale Way, there are three entries in the sidewalk data with the Road Log Id of 19515 that matches Silverdale Way, and they all indicate sidewalks on the left side of the road.
+Further note, all three have a segment ID of 0, so they will not match any of the segments in the roads data if we match on Road Log Id and Segment ID - an entirely different problem with the data that is not isolated to Silverdale Way.
+`r round(sum(nc_sw_sh$SEGMENT_ID ==0)/nrow(nc_sw_sh)*100,1)`% of the sidewalk data do not have segment ids.] For this draft example report it is okay; if the initial report is completed with the incomplete data, it will present a poorer overall condition of pedestrian facilities than is in place.
+
+```{r}
+df <- sidewalks_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+df2 <- sidewalks_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+
+df <- bind_rows(df, df2)
+
+knitr::kable(df, 
+             caption = "Miles of Sidewalks by Width of Sidewalk (single side)") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+```{r}
+df <- shoulders_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+df2 <- shoulders_nc %>%  
+  st_set_geometry(., NULL) %>%
+  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
+  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
+  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
+            .groups = "drop") %>%
+  select(-WIDTH)
+
+df <- bind_rows(df, df2)
+
+knitr::kable(df, 
+             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+### Combining the Data
+
+In order to do this evaluation, we must combine the roads data (which has information on the class and setting of roads) with the sidewalk and shoulder data (with information about the pedestrian facilities).
+However, the most obvious methods of combination fail.
+
+#### Matching on Road Log Id is good but insufficient
+
+All of the data sets appear to use the same set of road log IDs from public works.
+Unfortunately, knowing that there is a shoulder or sidewalk somewhere on a road does not get to how much of that road has the sidewalk or shoulder.^[As roads sometimes change names and road log IDs, there are also some breaks between the road data and sidewalk/shoulder data.
+This was noted for Madrone Ave in Manchester - given the same road log ID as Main St when the road bends.]
+
+This diagnostic and ugly plot is intended to show the coverage of road log id across the sidewalk and shoulder data sets compared to the raod data.
+While the sidewalk and shoulder data are more sparse, the range is consistent.
+
+```{r}
+sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
+                  aes(x = as.numeric(row.names(nc_sw_sh)), y = RD_LOG_ID)) + 
+  geom_point() +
+  labs(title = "Sidewalk and Shoulder Data",
+       x = "") +
+  expand_limits(y = c(0,100000))
+road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
+                  aes(x = as.numeric(row.names(ncdf)), y = RD_LOG_ID)) + 
+  geom_point() +
+  labs(title = "Roads Data",
+       x = "")
+
+library(patchwork)
+
+sw_plot + road_plot
+
+detach("package:patchwork", unload=TRUE)
+rm(sw_plot)
+rm(road_plot)
+```
+
+```{r}
+match_road_id <- st_set_geometry(sidewalks,NULL) %>%
+  mutate(RD_LOG_ID = as.integer(ROADLOGID)) %>%
+  left_join(select(st_set_geometry(clean_roadcl,NULL),RD_LOG_ID,FULL_NAME)) %>%
+  group_by(RD_LOG_ID, ROADNAME, SIDE_OF_RO, WIDTH) %>%
+  summarise(n = sum(!is.na(FULL_NAME)), .groups = "drop") %>% 
+  filter(n == 0, WIDTH > 0)
+```
+
+The most obvious smaller match is the segment id. However, 
+**NOTE: We cannot actually use the SEGMENT ID in this draft report because it does not have the same meaning in the sidewalk and shoulder data as it has in the road data.
+The following plots are diagnostic but show that the sidewalks and shoulders data do not have segment ids covering the same spectrum as the road data.
+
+```{r}
+sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
+                  aes(x = as.numeric(row.names(nc_sw_sh)), y = SEGMENT_ID)) + 
+  geom_point() +
+  labs(title = "Sidewalk and Shoulder Data",
+       x = "")+
+  expand_limits(y = c(0,20000))
+road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
+                  aes(x = as.numeric(row.names(ncdf)), y = SEGMENT_ID)) + 
+  geom_point() +
+  labs(title = "Roads Data",
+       x = "")
+
+library(patchwork)
+
+sw_plot + road_plot
+
+detach("package:patchwork", unload=TRUE)
+rm(sw_plot)
+rm(road_plot)
+```
+
+The segment ids are clearly not from the same space, but the road log ids are consistent.^[Mostly.] However, we cannot combine on road log id alone - because the roads change character and the type of pedestrian facilities over space.
+We combine instead on geometry.
+
+For now, we are making a space around a sidewalk or shoulder geometry, a buffer, to capture the nearby area where the road would be, and seeing if there are any road segments that intersect.
+If the road segment intersects and has the same road log id, we keep it.
+
+```{r combine-rd-sw-sh}
+# combine the sidewalk and shoulder and roads data
+x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 600)
+x$RD_LOG_ID <- 43809
+comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
+                                     SEGMENT_ID_R = SEGMENT_ID,
+                                     FULL_NAME_R = FULL_NAME, 
+                                     LENGTH_R = LENGTH,
+                                     road_class, RUCODE), 
+                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
+  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
+  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
+  mutate(distance = as.integer(st_length(.)))
+
+```
+
+## Method of Evaluation
+
+Road segments are the basis for evaluation because they are the smallest unit in common across the data and can be grouped together to form trips.
+Actually matching up the data for roads, shoulders, and sidewalks will prove to be a more complicated task than it appears at first blush.
+As noted in the [Data section](#datasection), the Segment IDs in the shoulder and sidewalk data do not cover the range of segments as in the road data.
+Here, we join based on the road log id and beginning and ending mileage point
+
+Each road segment is evaluated individually.
+For each road segment, its adequate pedestrian facility level is based on its classification and the Kitsap County Road Standards Table 3.3 and 3.4.[@kc_road_standards]^[revisions of road standards should be followed by review and reevaluation of these criteria]
+
+```{r tbl_standards}
+# Road Standards Tables 3.3 and 3.4
+library(tidyverse)
+library(kableExtra)
+road_standards_given <- tribble(
+  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, 
+  "Urban", "Local Road", NA, 5, NA,
+  "Urban", "Local Sub-collector", NA, 5, NA,
+  "Urban", "Arterial", NA, 6, NA,
+  "Urban", "Collector", NA, 6, NA,
+  "Rural", "Local Road", NA, NA, 3,
+  "Rural", "Local Sub-collector", NA, NA, 4,
+  "Rural", "Collector", "ADT 400-750", NA, 3,
+  "Rural", "Collector", "ADT 751-1000", NA, 4,
+  "Rural", "Collector", "DHV 100-200", NA, 6,
+  "Rural", "Collector", "DHV > 200", NA, 8,
+  "Rural", "Minor Arterial", "DHV < 100", NA, 4,
+  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,
+  "Rural", "Minor Arterial", "DHV > 200", NA, 8,
+  "Rural", "Principal Arterial", "DHV <200", NA, 6,
+  "Rural", "Principal Arterial", "DHV > 200", NA, 8,
+)
+
+knitr::kable(road_standards_given,
+             caption = "Summary of Tables 3.3 and 3.4 in Kitsap County Road Standards [@kc_road_standards]")  %>%
+   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT.
+Note: sub-collector, ADT, and DHV are not defined in the publicly available road data.") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+Because of all of the items that are not available in the road data at this time, this draft makes some assumptions to allow for an evaluation.
+When the information is known, a more accurate evaluation can be made.^[Roads with unknown setting - RUCODE is 0, we match them to the Urban road standards.
+This may be incorrect.]
+
+```{r tbl_standards_simple}
+# Road Standards Tables 3.3 and 3.4 translated for data we have
+library(tidyverse)
+library(kableExtra)
+road_standards <- tribble(
+  ~setting, ~class, ~sidewalk, ~shoulder,
+  "Urban", "Local Access", 5, NA,
+  "Urban", "Minor Collector",6, NA,
+  "Urban", "Major Collector",6, NA,
+  "Urban", "Minor Arterial", 6, NA,
+  "Urban", "Principal Arterial", 6, NA,
+  "Rural", "Local Access", NA, 3,
+  "Rural", "Minor Collector",NA, 3,
+  "Rural", "Major Collector",NA, 4,
+  "Rural", "Minor Arterial", NA, 4,
+  "Rural", "Principal Arterial", NA, 6,
+)
+
+knitr::kable(road_standards,
+             caption = "Simplified Assumptions to Allow for Evaluation")  %>%
+   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+Road segments that are explicitly excluded from the evaluation will be scored NA and not included in a summary of pedestrian metrics in an area (they will neither improve nor pull down the summary score).
+
+Pedestrian facilities on road segments will be scored based on the sum of the score of the sides. A side of the road for a segment will be scored:
+
+- 5 if it has a sidewalk or shoulder of adequate size (based on road classification & standards) or a sidewalk is present where a shoulder is the standard
+- 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
+- 1 if the sidewalk or shoulder is present and is narrower than AASHTO (**NOTE to subcommittee - do we think we should use a 1 if there is a shoulder when there should be a sidewalk?**)
+- 0 if none present or no data
+
+Therefore, segments with sidewalk or shoulder of adequate size (based on road classification) on both sides of road would receive a 10.
+Segments with no data or with no sidewalk or shoulder would receive a 0.
+
+Any given area score will be a weighted average of the scores of the segments within the area. Weighting is by centerline linear feet per segment. A simple example is an area composed of 4 segments.
+
+```{r}
+example_tbl <- tribble(
+  ~Segment, ~"Segment Score", ~"Segment Centerline Linear Feet",
+  "A", 6, 100,
+  "B", 0, 10,
+  "C",10,50,
+  "D",8,200,
+  "Total",7.5 ,360,
+)
+
+knitr::kable(example_tbl) %>%
+kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
+              full_width = F, 
+              position = "float_left")
+```
+
+The Total Score would be the segment scores multiplied by their linear feet and then divided by the total linear feet of the area (360).
+For this example, we get a area score of 7.5 $$ 6 x 100 + 0 x 10 + 10 x 50 + 8 x 200 = 2700$$ $$2700/360 = 7.5$$
+
+```{r}
+score_interpretation <- tribble(
+  ~score_at_least, ~interpretation,
+  10, "Good",
+  5, "Adequate",
+  3, "Minimum",
+  1, "Poor",
+  0.001, "Very Poor",
+  0, "Nothing"
+)
+
+getScoreInterpretation <- function(x){
+  first(score_interpretation$interpretation[
+        score_interpretation$score_at_least <= x]
+    )
+}
+
+df <- score_interpretation %>%
+  mutate(`Score At Least` = ifelse(score_at_least == 0.001, "> 0", 
+                                   nice_comma(score_at_least,0))) %>%
+  select(`Score At Least`, Interpretation = interpretation)
+
+knitr::kable(df,
+             caption = "Interpretation of Scores for Segments and for Areas")  %>%
+   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
+                full_width = F)
+```
+
+# Results
+First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.
+
+## Overall
+
+```{r compare_standards}
+# take the road cl data for the roads under evaluation
+# remove geometry
+# merge with standards to get requirement based on setting (RUCODE) and class (road_class)
+# merge with sidewalk and shoulder data to get what exists
+
+library(tidyverse)
+library(sf)
+
+road_standards <- road_standards %>%
+  rename(standard_sidewalk = sidewalk, 
+         standard_shoulder = shoulder, 
+         road_class = class)
+
+getSideScore <- function(standard_sidewalk, standard_shoulder,
+                         sidewalk, shoulder) {
+  ifelse(is.na(standard_sidewalk), # dealing with shoulder
+         case_when(
+           !is.na(sidewalk) ~ 5, # there is a sidewalk present when a shoulder is the standard
+           shoulder >= standard_shoulder ~ 5,
+           shoulder < standard_shoulder & shoulder > 4 ~ 3,
+           shoulder < standard_shoulder & shoulder > 0 ~ 1,
+           TRUE ~ 0
+         ),
+         case_when(
+           sidewalk >= standard_sidewalk ~ 5,
+           sidewalk < standard_sidewalk & sidewalk > 4 ~ 3,
+           sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
+           TRUE ~ 0
+         )
+  )
+}
+
+comb_stand <- comb %>% 
+  st_set_geometry(., NULL) %>% 
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown")),
+         ped_type = sub("\\:.*", "", TYPE),
+         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
+  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), 
+              names_from = nn, 
+              values_from = WIDTH,
+              values_fn = mean) %>%
+  left_join(road_standards)
+
+comb_score <- comb_stand %>%
+  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_left, shoulder_left),
+         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
+                                   sidewalk_right, shoulder_right),
+         score = score_left + score_right)
+
+```
+
+```{r}
+
+comb_score2 <- st_set_geometry(ncdf, NULL) %>% 
+  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
+         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
+  left_join(comb_score) %>%
+  mutate(the_score = ifelse(is.na(score), 0, score),
+         the_length = ifelse(is.na(distance), LENGTH, distance),
+         w_score = the_score*the_length)
+
+comb_total_score <- sum(comb_score2$w_score)/sum(comb_score2$the_length)
+
+comb_score_by_setting <- comb_score2 %>%
+  group_by(setting) %>%
+  summarise(score = sum(w_score)/sum(the_length),
+            .groups = "drop") %>%
+  rowwise() %>%
+  mutate(interpretation = getScoreInterpretation(score))
+
+comb_score_by_setting_class <- comb_score2 %>%
+  group_by(setting, road_class) %>%
+  summarise(score = sum(w_score)/sum(the_length),
+            .groups = "drop") %>%
+  rowwise() %>%
+  mutate(interpretation = getScoreInterpretation(score))
+```
+
+The overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, `r getScoreInterpretation(comb_total_score)`.
+
+Looking by the same breakdown that we have for standards, we can see where the scores are highest and lowest.
+
+```{r tbl_combscorestandard}
+knitr::kable(comb_score_by_setting_class %>%
+               rename(Setting = setting,
+                      Class = road_class,
+                      Score = score,
+                      Interpretation = interpretation) %>%
+               mutate(Score = ifelse(Score == 0, "0", nice_comma(Score, 1))),
+             caption = "Weighted Scores by Standard Group")  %>%
+   footnote(general = "Unknown setting is scored as if it is Urban."
+                      ) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+```{r map_score}
+
+comb_map <- st_as_sf(ncdf) %>% 
+  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
+         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>% 
+  left_join(comb_score) %>%
+  rowwise() %>%
+  mutate(interpretation = getScoreInterpretation(score))
+
+score_col <- scale_color_manual(values = c("Good"="#4d9221",
+                                           "Adequate"="#a1d76a",
+                                           "Minimum" = "#e6f5d0",
+                                           "Poor" = "#fde0ef",
+                                           "Very Poor" = "#e9a3c9",
+                                           "Nothing" = "#c51b7d"),
+                                na.value = "#808080")
+comb_map$interpretation <- factor(comb_map$interpretation, 
+                                  levels =  score_interpretation$interpretation)
 kitsap + 
-  geom_sf(data = transit_cp2, aes(fill = id), 
-          lwd = 0, alpha = 0.3) +
-  geom_sf(data = transit_cp, aes(color = 1), color = "black", show.legend = FALSE) +
+  geom_sf(data = cities["NAME"], 
+          aes(fill="NAME"),
+          fill="#808080",
+          alpha = 0.5,
+          show.legend = FALSE) +
+  geom_sf(data = ncdf, aes(color = 1), color = "#808080", show.legend = FALSE) + 
+  geom_sf(data = comb_map, aes(color = interpretation), show.legend = "line") + 
+  score_col +
   theme_void() +
-  labs(title = 'Kitsap County Transit Centers and Transfer Points',
-       subtitle = "Centers not labeled with color are not in the unincorporated county",
-       caption = "Data source: Kitsap Transit Website\n
-                  Walking distance polygons from Mapbox",
-       fill = "Transit Centers")
-
-```
-
-```{r loop_transit,include=FALSE,echo=FALSE, cache = TRUE}
-the_transits <- unique(st_set_geometry(transit_cp2, NULL)$id)
-loop_df <- st_as_sf(transit_cp2) %>% mutate(tomatch = id)
-
-score_breakdowns <- score_breakdowns %>%
-  bind_rows(tibble(Area = the_transits, Score = NA))
-
-expanded_child <- lapply(
-    the_transits,
-    function(x) {
-      knitr::knit_expand(file = "child_single_geo.Rmd",
-                                   current_geo = x )
+  labs(title = "Calculated Pedestrian Facility Score",
+       caption = "Data source: Kitsap GIS, scores calculated",
+       color = "Score Interpretation")
+
+```
+
+```{r}
+# More detailed information
+# knitr::kable(
+#   comb_score2 %>%
+#   group_by(setting, road_class, score) %>% 
+#   summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
+#             segments = nice_comma(n()), 
+#             #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
+#             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
+#             .groups = "drop") 
+#   )
+```
+
+```{r}
+ggplot(filter(comb_score2, setting != "Unknown"), 
+       aes(x = as.factor(score), y = the_length/5280))+
+  geom_col() +
+  labs(x = "score based on data available", 
+       y = "centerline miles of road") +
+  theme_minimal() +
+  facet_grid(setting ~ road_class)
+```
+
+If we remove the segments where we do not have information about sidewalks and shoulders at all, we can see that for those, we are generally finding higher scores for rural roads - with paved shoulders.
+
+```{r}
+ggplot(filter(comb_score2, setting != "Unknown", !is.na(score)), 
+       aes(x = as.factor(score), y = LENGTH/5280))+
+  geom_col() +
+  labs(x = "score based on data available", 
+       y = "centerline miles of road") +
+  theme_minimal() +
+  facet_grid(setting ~ road_class)
+```
+
+Some of the segments with score of 0 are because they have shoulders when they should have sidewalks.
+If you are familiar with Central Kitsap, you might recognize this kind of situation on Provost/Old Frontier - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.
+Here's a few example segments.
+
+```{r}
+
+roads_0 <- comb_score2 %>%
+  filter(score == 0) %>%
+  select(setting, road_class, FULL_NAME_R, sidewalk_right:standard_sidewalk)
+
+knitr::kable(head(roads_0) %>%
+               rename(Setting = setting,
+                      Class = road_class,
+                      Name = FULL_NAME_R,
+                      `sidewalk R` = sidewalk_right,
+                      `sidewalk L` = sidewalk_left,
+                      `shoulder R` = shoulder_right,
+                      `shoulder L` = shoulder_left,
+                      `standard sidewalk` = standard_sidewalk)) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+
+```
+
+Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road.
+A few details for roads where at least one segment has a 10 are shown here.
+```{r}
+roads_10 <- comb_score2 %>%
+  filter(score == 10) %>%
+  group_by(setting, road_class, FULL_NAME_R) %>%
+  summarise(segments = nice_comma(n(),0), 
+            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
+            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
+            .groups = "drop") %>%
+  rename(Setting = setting,
+         Class = road_class,
+         Name = FULL_NAME_R,
+         `Segments w 10` = segments,
+         `Segment Length` = road_len,
+         `Ped Fac Length` = dist)
+
+knitr::kable(roads_10) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```
+
+Some road segments are missing entirely from the centerline road data - for example, the portion of Ridgetop that goes from the end of Myhre and curves back to Silverdale Way.
+These missing road segments on Ridgetop are mostly complete streets with wide sidewalks.
+Here, showing a map with all roads with the name *RIDGETOP*, you can only see the part going from Silverdale Way to Myhre.^[For completeness, we also checked the road log id, and it has the same data.] These segments, and any others like them, missing from the data decrease the overall pedestrian facilities score.
+
+```{r}
+kitsap +
+  geom_sf(data = filter(clean_roadcl, STREETNAME == "RIDGETOP"), aes(color = road_class)) + 
+  road_col +
+  theme_void() +
+  labs(title = "STREETNAME == RIDGETOP",
+       color = "")
+```
+
+## Smaller Breakdowns of the County
+
+We can also look at regions and pedestrian generators within the County.
+For these, we identify the roads and any sidewalks or shoulders.^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]
+
+Regions are already defined by an area.
+For pedestrian generators, we identify an area that is within 30 minute walking distance of the facility and evaluate roads therein.^[30 minute walking distance calculated using the mapboxapi function `mb_isochrone`.
+Isochrones use map data from Mapbox and OpenStreetMap and their data sources.
+To learn more, visit https://www.mapbox.com/about/maps/ and http://www.openstreetmap.org/copyright.[@mapbox]]
+
+These regions and pedestrian generators are:
+
+```{r, results = 'asis'}
+cat(paste0("- ", breakdowns$Name, collapse = "\n"))
+```
+
+For each smaller region or pedestrian generator type, we show a map of Kitsap County (sans incorporated cities) indicating locations of the regions of that type.
+
+Then, for each individual region, we show the pedestrian facilities on the left side and the right side of the road and calculate scores.
+
+```{r}
+# this starts the tibble that will hold ALL the scores
+score_breakdowns <- tibble(Area = "Kitsap County", Score = comb_total_score)
+```
+
+```{r bdlooping, include=FALSE, echo=FALSE}
+
+expanded_region <- lapply(
+    seq_along(1:nrow(breakdowns)),
+    function(irow) {
+      knitr::knit_expand(file = "child_region.Rmd",
+                                   i = irow )
     })
 
-parsed_child <- knitr::knit_child(text = unlist(expanded_child))
-```
-
-`r parsed_child`
-
-## Conclusion
+parsed_region <- knitr::knit_child(text = unlist(expanded_region))
+```
+
+`r parsed_region`
+
+# Conclusion
 
@@
 Smaller breakdowns of the county fare differently.
+Here, we show a table for each region type.
+Any region with no segments getting more than a 0, "Nothing" show up with red text.
+The best region of that type will have bold text, and any regions with a score of **Adequate** will have green text.
 
@@
 getScoreBreakdownForArea <- function(the_area) {
-  knitr::kable(score_breakdowns %>% 
+  # get name
+  the_area_name <- breakdowns$Name[breakdowns$items == the_area]
+  the_area_df <- score_breakdowns %>% 
                  filter(Area %in% eval(as.symbol(the_area))) %>%
-                 mutate(Score = nice_comma(Score,2) 
-                                           ),
-             caption = paste("Scores for",
-                             str_to_title(sub("the_"," ",the_area))
-                             )
+                 mutate(s = Score,
+                        Score = ifelse(Score == 0,
+                                       nice_comma(Score,0),
+                                       nice_comma(Score,2))) %>%
+                 arrange(Area)
+  knitr::kable( select(the_area_df, -s),
+                caption = paste("Scores for", the_area_name )
              ) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
+  row_spec(which(the_area_df$s == 0), color = "red") %>%
+  row_spec(which(the_area_df$s == max(the_area_df$s)), bold = T) %>%
+  row_spec(which(the_area_df$Interpretation == "Adequate"), bold = T, color = "green")
 }
@@
 score_breakdowns <- read_csv(paste0(this_dir,"score_breakdowns.csv"))
-res <- lapply(c("the_commdists","the_ugas", "the_lamirds", "the_schools", "the_libraries", "the_transits"), function(x) {
+res <- lapply(breakdowns$items, function(x) {
   knitr::knit_child(text = c(
-    '## `r str_to_title(gsub("_"," ",x))`',
     '',
@@
  * missing part of Ridgetop Blvd in Silverdale where sidewalks are known to exist
+ * missing data on sidewalks that are known to exist
  * unmatched road log identifiers between road and sidewalk and shoulder data
@@
 
-In addition, many road segments have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**
+In addition, many road segments have shoulders when their standards suggest they should have sidewalks.
+Is the presence of a shoulder, though not adequate, really the same as **nothing?** - that is how the scoring currently calculates.
+
+One cautionary note - while this method evaluates pedestrian facilities, it is entirely focused on the presence of facilities of adequate width and separation from cars.
+It is not taking into consideration the quality of those facilities.
+Disrepair (cracks, uneven surfaces) and encroaching vegetation can greatly reduce the usability of pedestrian facilities.
+Quality of facilities is beyond the scope of this evaluation.
 
@@
 
-## References
+# References

### 2022-10-17 Draft-Ped-Metrics-Example.Rmd
updates to capture score breakdowns in conclusion

@@
   html_document:
+    number_sections: true
     toc: true
+    toc_float: true
     toc_depth: 3
@@
 As noted in the [Data section](#datasection), the Segment IDs in the shoulder and sidewalk data do not cover the range of segments as in the road data.
-Here, we join based on the road log id and beginning and ending mileage point
+Here, we join based on the geometries of the roads and sidewalks/shoulders.^[We intersect the road data with a buffered geometry of the sidewalk or shoulder and then remove any matching roads that do not have the same road log id.]
 
@@
 
+This example shows how the aggregation works.
+With a weighted score of 7.5, we have to determine how that should be interpreted.
+We developed a score interpretation.
+Generally, we want segments to score a 10, right?
+But if it has at least a 5, it is adequate.
+
 ```{r}
@@
 First, we calculate the score overall for the county and then break it down for other smaller units of area in our community.
+Every score shown, except when explicitly for a single segment, is a weighted score based on the length of segments of road and their respective pedestrian facilities (sidewalks or shoulders) as shown in the example above.
 
@@
 Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road.
-A few details for roads where at least one segment has a 10 are shown here.
+A few details for a few roads where at least one segment has a 10 are shown here.^[This table is only showing the first few entries because it is a long table.]
 ```{r}
@@
 
-knitr::kable(roads_10) %>%
+knitr::kable(head(roads_10)) %>%
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

### 2022-10-17 child_single_geo_in_region.Rmd
updates to capture score breakdowns in conclusion

@@
 ```
+
 ```{r}

### 2022-10-18 Draft-Ped-Metrics-Example.Rmd
change the substitute standards for rural collectors

@@
 
-1. Provide a complete evaluation of pedestrian facilities across the county down to the segment and side of road level
+1. Provide a complete evaluation of pedestrian facilities across the unincorporated county down to the segment and side of road level
 2. Present an aggregated quality rating for pedestrian facilities in specific areas
@@
 This draft example was prepared by the Kitsap County Non-Motorized Community Advisory Committee Special Committee on Pedestrian Metrics.
-It is NOT considered an authoritative source of these data and is NOT an official report of the county.
+It is NOT considered an authoritative source of these data and is NOT an official report of the county and should NOT be used for any decision-making.
 
@@
 
-This report provides a complete evaluation of pedestrian facilities in Kitsap County comparing the actual facilities to a minimum adequate pedestrian facility level.
+This report provides a complete evaluation of pedestrian facilities in unincorporated Kitsap County comparing the actual facilities to a minimum adequate pedestrian facility level.
 The minimum adequate pedestrian facility level is either a paved shoulder or a sidewalk of a certain minimum width - depending on the character of the road.
@@
 The intended audience is the people of Kitsap County and law/decision makers determining how to target limited funds for transportation development.
+The information should be used alongside other valuable inputs such as from the community advisory committees, (e.g.
+Non-motorized, Accessibility, Community Development).
 
@@
 
- - Inside of cities (Bremerton, Port Orchard, etc.)^[Draft excludes segments that intersect with the Kitsap County incorporated city limits polygons[@kcgis_cities]]
+ - Inside of incorporated cities (Bremerton, Port Orchard, etc.)^[Draft excludes segments that intersect with the Kitsap County incorporated city limits polygons[@kcgis_cities]]
  - Local easements^[Draft assuming Function Class 100 or 110 in Road Centerline Data]
  - Highways^[Draft Assuming Function Class 2 in Road Centerline Data]
- - Speed limit > 50 MPH^[**these are not currently identified because the road centerline data does not include speed limit**]
+ - Speed limit > 50 MPH^[**these are not currently identified because the road centerline data does not include speed limit, but no county road should have a speed limit exceeding this.**]
 
@@
   "Rural", "Local Access", NA, 3,
-  "Rural", "Minor Collector",NA, 3,
-  "Rural", "Major Collector",NA, 4,
+  "Rural", "Minor Collector",NA, 4,
+  "Rural", "Major Collector",NA, 6,
   "Rural", "Minor Arterial", NA, 4,
@@
 Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road.
-A few details for a few roads where at least one segment has a 10 are shown here.^[This table is only showing the first few entries because it is a long table.]
+A few details for a few roads where at least one segment has a 10 are shown here.^[This table is only showing the first few entries because it is a long table.] **Debbie - these would be the subset of segments that "meet the standard"**
 ```{r}

### 2022-10-31 Child-Metrics-Geo.Rmd
many updates not tracked

```{r}
x <- as.integer({{x}})
current_geo <- loop_df$tomatch[x]
current_geo_title <- ifelse(b_name %in% c("LAMIRDs"),
                            paste(current_geo, sub("s","",b_name)),
                            current_geo)
this_geo <- getGeoSFForRegion(loop_df,current_geo) 
this_comb_score <- getCombSFForRegion(this_geo)
has_roads <- TRUE # our assumption
if (is.logical(this_comb_score)) {
  has_roads <- FALSE
}
this_comb_total_score <- 0

```

#### `r current_geo_title`

```{r, results = 'asis', eval=!has_roads}
cat("This area does not have known pedestrian facilities.")
```

```{r, eval=has_roads}
this_comb <- getCombMapForRegion(comb_map, this_geo)
write_csv(this_comb_score, paste0(this_dir, 'region_score_ref/',b_name, x,'.csv'))

this_comb_total_score <- getAggregateScore(this_comb_score)
rm(this_comb_score)
```

```{r}
# add the score to the score breakdowns
if (!exists(deparse(substitute(this_comb_total_score)))){
  this_comb_total_score <- 0
}
score_breakdowns$Score[score_breakdowns$Area == current_geo] <- this_comb_total_score
```

```{r, eval=has_roads & has_points}
  # see if this geo has a point (is it an area around a place, like a school)
this_point <- st_as_sf(b_pointdf) %>%
      filter(NAME == this_geo$tomatch)
```

```{r, results = 'asis', eval=has_roads}
cat("This area has pedestrian facilities identified in the data; they are shown on the map for the left side and the right side of the road.")
```

```{r map-{{i}}{{x}}-sides, eval=has_roads}
#| fig.width = 14,
#| fig.height = 8,
#| out.width = '100%',
#| dpi = 96
#|

p1 <- getMapForRegionInCounty(this_geo, current_geo) +
    labs(title = paste("Pedestrian Facilities:",current_geo_title))
r <- getMapFacilitiesSideForRegion(this_geo, "right") +
    {if(has_points) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)} 
l <- getMapFacilitiesSideForRegion(this_geo, "left") +
    {if(has_points) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)}

  library(patchwork)

  p1 + r + l +
    plot_layout(guides = 'collect') & theme(legend.position = 'bottom')

  detach("package:patchwork", unload=TRUE)
```

The overall score for pedestrian facilities in `r current_geo_title` is **`r nice_comma(this_comb_total_score)`**, **`r getScoreInterpretation(this_comb_total_score)`.**

```{r map-{{i}}{{x}}-scores, eval=has_roads}
#| out.width = '100%',
#| dpi = 96

  p1 <- getMapForRegionInCounty(this_geo, current_geo) +
    labs(title = paste("Calculated Pedestrian Facility Score:",current_geo_title))
  p2 <- getMapScoreSummaryForRegion(this_geo, this_comb, current_geo) +
    labs(title = "") +
    {if(has_points) geom_sf(data = this_point, 
                         aes(color = 1), color = "black", show.legend = FALSE)}

  library(patchwork)

  p1 + p2 +
    plot_layout(guides = 'collect') & theme(legend.position = 'bottom')

  detach("package:patchwork", unload=TRUE)
```

```{r kbl-{{i}}{{x}}, eval=has_roads}

getKableScoreSummaryForMap(this_comb)
```

### 2022-10-31 Child-Metrics-Region.Rmd
many updates not tracked

--- 2022-10-10 child_region.Rmd
@@
-[comment]: # creates a section for a smaller geo passed with details
-[comment]: # loops over a number i referencing the geo in breakdowns tibble
-[comment]: # called from Draft-Ped-Metrics-Example.Rmd
-
 ```{r bdsetup-{{i}}}
@@
                                   nrow(b_polydf),
-                                  " are outside of city limits. Here they are shown with a 30 minute walking distance")
+                                  " are outside of city limits. Here they are shown with a 15 minute walking distance")
                            ),".")
@@
 subtitle_text <- ifelse(b_name %ni% c("Commissioner Districts","LAMIRDs"),
-                        "only showing those in the unincorporated county","")
+                        "only showing the unincorporated county","")
 caption_text <- paste("Data source:",breakdowns$data_source[i])
@@
   labs(title = title_text,
-       subtitle = subtitle_text,
+       #subtitle = subtitle_text,
        caption = caption_text,
@@
 
-```{r bdloop-{{i}}, include=FALSE, echo=FALSE, cache = TRUE}
+```{r bdloop-{{i}}, include=FALSE, echo=FALSE}
 
 expanded_child <- lapply(
-    b_items,
-    function(x) {
-      knitr::knit_expand(file = "child_single_geo_in_region.Rmd",
-                                current_geo = x )
+  seq_along(1:length(b_items)),
+    function(xrow) {
+      knitr::knit_expand(file = "Child-Metrics-Geo.Rmd",
+                                x = xrow )
     })

### 2022-10-31 Draft-Ped-Metrics-Example-Alternate.Rmd
many updates not tracked

--- 2022-10-18 Draft-Ped-Metrics-Example.Rmd
@@
   select(RD_LOG_ID,SEGMENT_ID,BMP,EMP,SIDE_OF_RO, WIDTH, LENGTH, TYPE)
+
+# SHLDR_SURF 
+# ACP - paved
+# BST - bituminous surface (chip seal), like paved
+# GRV - gravel
+# PAP - porous asphalt
 shoulders_nc <- st_intersection(shoulders, st_buffer(county_no_cities,0)) %>%
   st_as_sf(.) %>%
-  filter(SHLDR_WIDT > 0, SHLDR_SURF == "ACP", DISTANCE_F > 0) %>%
+  filter(SHLDR_WIDT > 0, 
+         SHLDR_SURF %in% c("ACP", "BST", "PAP"), # these are the "paved" ones
+         DISTANCE_F > 0) %>%
   mutate(LENGTH = as.integer(st_length(.)),
@@
 - 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
-- 1 if the sidewalk or shoulder is present and is narrower than AASHTO (**NOTE to subcommittee - do we think we should use a 1 if there is a shoulder when there should be a sidewalk?**)
+- *change in alternate* 2 if a shoulder is present of adequate width when a sidewalk should be present (based on road classification & standards)
+- *change in alternate* 1 if there is any sidewalk or shoulder but not wide enough
 - 0 if none present or no data
@@
            sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
+           shoulder >= standard_sidewalk ~ 2, # add for shoulder exist
+           shoulder < standard_sidewalk & shoulder > 0 ~ 1,
            TRUE ~ 0

### 2022-10-31 Draft-Ped-Metrics-Example_w_Gravel.Rmd
many updates not tracked

--- 2022-10-31 Draft-Ped-Metrics-Example.Rmd
@@
 ---
-title: "DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report"
+title: "DRAFT W GRAVEL! Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report"
 date: "`r Sys.Date()`"
@@
   filter(SHLDR_WIDT > 0, 
-         SHLDR_SURF %in% c("ACP", "BST", "PAP"), # these are the "paved" ones
+         SHLDR_SURF %in% c("ACP", "BST", "PAP", "GRV"), # these are the "paved" ones
          DISTANCE_F > 0) %>%
   mutate(LENGTH = as.integer(st_length(.)),
-         TYPE = paste0("Paved Shoulder: ", SHLDR_WIDT," ft"),
+         TYPE = paste0("Shoulder: ", SHLDR_WIDT," ft"),
          RD_LOG_ID = as.integer(ROAD_LOG_I),
@@
                                            "Sidewalk: 8 ft"="#449B75",
-                                           "Paved Shoulder: 1 ft"="#56A255",
-                                           "Paved Shoulder: 2 ft"="#7E6E85",
-                                           "Paved Shoulder: 3 ft"="#AC5782",
-                                           "Paved Shoulder: 4 ft"="#E3712B", 
-                                           "Paved Shoulder: 5 ft"="#FFA10D",
-                                           "Paved Shoulder: 6 ft"="#FFE528",
-                                           "Paved Shoulder: 7 ft"="#E1C62F",
-                                           "Paved Shoulder: 8 ft"="#B16C29",
-                                           "Paved Shoulder: 10 ft"="#C66764",
-                                           "Paved Shoulder: 12 ft"="#F17EB4",
-                                           "Paved Shoulder: 15 ft"="#CB8CAD",
-                                           "Paved Shoulder: 17 ft"="#999999"))
+                                           "Shoulder: 1 ft"="#56A255",
+                                           "Shoulder: 2 ft"="#7E6E85",
+                                           "Shoulder: 3 ft"="#AC5782",
+                                           "Shoulder: 4 ft"="#E3712B", 
+                                           "Shoulder: 5 ft"="#FFA10D",
+                                           "Shoulder: 6 ft"="#FFE528",
+                                           "Shoulder: 7 ft"="#E1C62F",
+                                           "Shoulder: 8 ft"="#B16C29",
+                                           "Shoulder: 10 ft"="#C66764",
+                                           "Shoulder: 12 ft"="#F17EB4",
+                                           "Shoulder: 15 ft"="#CB8CAD",
+                                           "Shoulder: 17 ft"="#999999"))
 nc_sw_sh$TYPE <- factor(nc_sw_sh$TYPE, levels = c("Sidewalk: 3 ft",
@@
                                                   "Sidewalk: 8 ft" ,
-                                                  "Paved Shoulder: 1 ft",
-                                                  "Paved Shoulder: 2 ft",
-                                                  "Paved Shoulder: 3 ft",
-                                                  "Paved Shoulder: 4 ft", 
-                                                  "Paved Shoulder: 5 ft",
-                                                  "Paved Shoulder: 6 ft",
-                                                  "Paved Shoulder: 7 ft",
-                                                  "Paved Shoulder: 8 ft",
-                                                  "Paved Shoulder: 10 ft",
-                                                  "Paved Shoulder: 12 ft",
-                                                  "Paved Shoulder: 15 ft",
-                                                  "Paved Shoulder: 17 ft"))
+                                                  "Shoulder: 1 ft",
+                                                  "Shoulder: 2 ft",
+                                                  "Shoulder: 3 ft",
+                                                  "Shoulder: 4 ft", 
+                                                  "Shoulder: 5 ft",
+                                                  "Shoulder: 6 ft",
+                                                  "Shoulder: 7 ft",
+                                                  "Shoulder: 8 ft",
+                                                  "Shoulder: 10 ft",
+                                                  "Shoulder: 12 ft",
+                                                  "Shoulder: 15 ft",
+                                                  "Shoulder: 17 ft"))
 ```

### 2022-10-31 Draft-Ped-Metrics.Rmd
many updates not tracked

--- 2022-10-18 Draft-Ped-Metrics-Example.Rmd
@@
 ---
-title: "DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report"
+title: "Example Kitsap County Pedestrian Facilities Evaluation Report"
 date: "`r Sys.Date()`"
@@
 
-nice_comma <- function(x, dec = 1){
-  format(round(as.numeric(x), dec), digits = NULL, nsmall=dec, big.mark=",") 
-}
-'%ni%' <- function(x,y)!('%in%'(x,y))
-
 this_dir <- '~/Documents/non-motorized/'
@@
 library(sf)
+source(paste0(this_dir,'functions/reportFunctions.R'))
 ```
@@
 
-source(paste0(this_dir,'functions/getCleanRoadCenterLine.R'))
-library(tidyverse)
-library(sf)
+source(paste0(this_dir,'functions/getReportData.R'))
 
-outline <- sf::read_sf(paste0(this_dir,"data/outline"))
-outline$plain = 1
-kitsap <- ggplot() + 
-  geom_sf(data = outline["plain"], alpha = 0.5) + 
-  theme_void()
-
-cities <- sf::read_sf(paste0(this_dir,"data/cities"))
-
-# try to make a kitsap without cities
-county_no_cities <- st_difference(outline, st_union(cities))
-
-uga <- sf::read_sf(paste0(this_dir,"data/uga"))
-uga$isuga = 1
-lamird <- sf::read_sf(paste0(this_dir,"data/lamird"))
-lamird$islamird = 1
-community <- sf::read_sf(paste0(this_dir,"data/community"))
-schools <- sf::read_sf(paste0(this_dir,"data/schools"))
-
-# start plotting roads
-# want always to see same color for same type of road, so
-#['#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf']
-# NOTE: Also make this a factor so that it is sorted properly!!
-road_col <- scale_color_manual(values = c("Freeway / Expressway" = "#e41a1c",
-                                          "Principal Arterial" = "#377eb8",
-                                          "Minor Arterial" = '#4daf4a',
-                                          "Major Collector" = "#984ea3",
-                                          "Minor Collector" ='#ff7f00',
-                                          "Local Access" = '#a65628',
-                                          "Rural Local Access (e.g. easement or other type)" = '#f781bf',
-                                          "Urban Local Access (e.g. easement or other type)" = '#ffff33'
-                                          ))
-clean_roadcl$road_class <- factor(clean_roadcl$road_class, levels = c("Freeway / Expressway",
-                                                                      "Principal Arterial",
-                                                                      "Minor Arterial",
-                                                                      "Major Collector",
-                                                                      "Minor Collector",
-                                                                      "Local Access",
-                                                                      "Rural Local Access (e.g. easement or other type)",
-                                                                      "Urban Local Access (e.g. easement or other type)"))
-
-# exclude Local Access EASEMENTS & Highway/Freeway
-road_df <-st_as_sf(clean_roadcl) %>%
-  filter(FCLASS %ni% c(100,110,1,2), RD_LOG_ID > 0) %>%
-  mutate(LENGTH = as.integer(st_length(.)))
-
-# need to make sure that this is right
-
-ncdf <- st_intersection(road_df, st_buffer(county_no_cities,0)) %>%
-  mutate(LENGTH = as.integer(st_length(.)))
+cities <- getDataCities()
+county_no_cities <- getDataCountyNoCities()
+kitsap <- getMapKitsap()
+clean_roadcl <- getDataRoad()
+ncdf <- getDataRoadNoCities(clean_roadcl, county_no_cities)
+sidewalks_nc <- getDataSidewalks(county_no_cities)
+shoulders_nc <- getDataShoulders(county_no_cities)
+nc_sw_sh <- getDataFacilities(sidewalks_nc,shoulders_nc)
 
@@
 
-```{r swshdata, cache = TRUE}
-library(tidyverse)
-library(sf)
-shoulders <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SHOULDERS")
-sidewalks <- sf::read_sf("./data/KitsapShapefiles03_04_2022/SIDEWALKS")
+```{r moredata, cache = TRUE}
 
-sidewalks_nc <- st_intersection(sidewalks, st_buffer(county_no_cities,0)) %>%
-  st_as_sf(.) %>%
-  filter(WIDTH > 0, LENG_FT > 0) %>%
-  mutate(LENGTH = as.integer(st_length(.)),
-         TYPE = paste0("Sidewalk: ",WIDTH, " ft"),
-         RD_LOG_ID = as.integer(ROADLOGID),
-         SEGMENT_ID = as.integer(SEG_ID)) %>%
-  select(RD_LOG_ID,SEGMENT_ID,BMP,EMP,SIDE_OF_RO, WIDTH, LENGTH, TYPE)
-shoulders_nc <- st_intersection(shoulders, st_buffer(county_no_cities,0)) %>%
-  st_as_sf(.) %>%
-  filter(SHLDR_WIDT > 0, SHLDR_SURF == "ACP", DISTANCE_F > 0) %>%
-  mutate(LENGTH = as.integer(st_length(.)),
-         TYPE = paste0("Paved Shoulder: ", SHLDR_WIDT," ft"),
-         RD_LOG_ID = as.integer(ROAD_LOG_I),
-         SEGMENT_ID = as.integer(SEG_ID)) %>%
-  select(RD_LOG_ID,SEGMENT_ID,BMP,EMP,SIDE_OF_RO, WIDTH = SHLDR_WIDT, LENGTH, TYPE)
-nc_sw_sh <- bind_rows(sidewalks_nc, shoulders_nc)
+comb_score2 <- getCombScore2(ncdf, nc_sw_sh)
 
-# NOTE: Also make this a factor so that it is sorted properly!!
-sw_sh_col <- scale_color_manual(values = c("Sidewalk: 3 ft"="#E41A1C",
-                                           "Sidewalk: 4 ft"="#874F6F",
-                                           "Sidewalk: 5 ft"="#3881B0",
-                                           "Sidewalk: 8 ft"="#449B75",
-                                           "Paved Shoulder: 1 ft"="#56A255",
-                                           "Paved Shoulder: 2 ft"="#7E6E85",
-                                           "Paved Shoulder: 3 ft"="#AC5782",
-                                           "Paved Shoulder: 4 ft"="#E3712B", 
-                                           "Paved Shoulder: 5 ft"="#FFA10D",
-                                           "Paved Shoulder: 6 ft"="#FFE528",
-                                           "Paved Shoulder: 7 ft"="#E1C62F",
-                                           "Paved Shoulder: 8 ft"="#B16C29",
-                                           "Paved Shoulder: 10 ft"="#C66764",
-                                           "Paved Shoulder: 12 ft"="#F17EB4",
-                                           "Paved Shoulder: 15 ft"="#CB8CAD",
-                                           "Paved Shoulder: 17 ft"="#999999"))
-nc_sw_sh$TYPE <- factor(nc_sw_sh$TYPE, levels = c("Sidewalk: 3 ft",
-                                                  "Sidewalk: 4 ft",
-                                                  "Sidewalk: 5 ft",
-                                                  "Sidewalk: 8 ft" ,
-                                                  "Paved Shoulder: 1 ft",
-                                                  "Paved Shoulder: 2 ft",
-                                                  "Paved Shoulder: 3 ft",
-                                                  "Paved Shoulder: 4 ft", 
-                                                  "Paved Shoulder: 5 ft",
-                                                  "Paved Shoulder: 6 ft",
-                                                  "Paved Shoulder: 7 ft",
-                                                  "Paved Shoulder: 8 ft",
-                                                  "Paved Shoulder: 10 ft",
-                                                  "Paved Shoulder: 12 ft",
-                                                  "Paved Shoulder: 15 ft",
-                                                  "Paved Shoulder: 17 ft"))
+comb_total_score <- getAggregateScore(comb_score2)
+
+comb_map <- getCombMapSFFromScoreAndNCDF(ncdf, comb_score2) 
 ```
 
-```{r gwfunction}
-getWalkingPolygonInCounty <- function(locdf, idcol) {
-  # finds the items in locdf that are in the target area
-  # then finds the polygon around them
-  # then limits that again to the area in the target (because the polygon may go into a city)
-  if (!is.na(st_crs(locdf)$epsg)) { # the locdf has a epsg and cnc is NA
-    cnc <- st_transform(county_no_cities, st_crs(locdf)$epsg)
-    fdf <- st_intersection(locdf, st_buffer(cnc,0)) 
-  } else {
-    fdf <- st_intersection(locdf, st_buffer(county_no_cities,0)) 
-  }
-  fdf2 <- mapboxapi::mb_isochrone(fdf,
-                                  time = 30, "walking", id_column=idcol)
-  cnc <- st_transform(county_no_cities, st_crs(fdf2)$epsg)
-  st_intersection(fdf2, st_buffer(cnc,0)) 
-}
-```
+<p style="font-family: times, serif; font-size:24pt; font-weight:bold; font-style:italic">
+    Executive Summary
+</p>
 
-```{r smallerdata}
-# this section sets up the data that will be used for subdivisions of the county
-# get any data not set up yet
-commdists <- sf::read_sf(paste0(this_dir,"data/commdist"))
-commdists <- st_intersection(commdists, st_buffer(county_no_cities,0)) 
-commdists <- st_as_sf(commdists) %>% 
-  mutate(NAME = ifelse(DISTRICT == "1", "1 North",
-                ifelse(DISTRICT == "2", "2 South", "3 Central")))
-the_commdists <- unique(st_set_geometry(commdists, NULL)$NAME)
-loop_commdists <- st_as_sf(commdists) %>% mutate(tomatch = NAME)
+<div style="font-family: times, serif; font-size:14pt;  font-style:italic">
+This is an example of the first in a proposed series of regular Unincorporated Kitsap County Pedestrian Facilities Evaluation Reports.
 
-uga_no_city <- st_as_sf(uga) %>%
-  filter(substring(GMA_JURISD, 1, 4) != "City")
-the_ugas <- unique(st_set_geometry(uga_no_city, NULL)$GMA_JURISD)
-loop_ugas <- st_as_sf(uga_no_city) %>% mutate(tomatch = GMA_JURISD)
+In this example, pedestrian facilities on Unincorporated Kitsap County roads score **`r nice_comma(comb_total_score)` out of 10** which is evaluated as **`r getScoreInterpretation(comb_total_score)`**.
+This overall score is weighted by length of centerline roadway and the presence and width of road shoulders or sidewalks compared to the Kitsap County Road Standards.
 
-the_lamirds <- unique(st_set_geometry(lamird, NULL)$PLAN_AREA)
-loop_lamirds <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)
+Details and scores by Commissioner Districts as well as other smaller regions of the county are included in this report.
 
-pub_schools <- schools %>%
-  filter(TYPE == "PUBLIC")
-pub_schools2 <- getWalkingPolygonInCounty(pub_schools, "NAME")
-the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
-loop_schools <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)
-
-libraries <- community %>%
-  filter(TYPE == "PUBLIC LIBRARY")
-libraries2 <- getWalkingPolygonInCounty(libraries, "NAME")
-the_libraries <- unique(st_set_geometry(libraries2, NULL)$id)
-loop_libraries <- st_as_sf(libraries2) %>% mutate(tomatch = id)
-
-transit_cp <- tribble(
-  ~NAME, ~ADDRESS,
-  "Bremerton Transportation Center (Ferry Terminal)","10 Washington Ave, Bremerton 98337",
-  "Bainbridge Island Ferry Terminal","270 Olympic Drive SE, Bainbridge 98110",
-  "Wheaton Way Transit Center","3915 Wheaton Way, Bremerton 98310",
-  "West Bremerton Transit Center","540 Bruenn Ave, Bremerton 98312",
-  "Silverdale Transit Center","Greaves Way & Kitsap Mall Blvd, Silverdale 98383",
-  "North Viking Transit Center","21992 Viking Ave NW, Poulsbo 98370",
-  "Kingston Ferry Terminal","11264 State Route 104, Kingston 98346",
-  "Port Orchard Ferry Dock","73 Sidney Ave, Port Orchard 98366",
-  "Southworth Ferry Terminal","11564 SE State Hwy. 160, Southworth 98386",
-  "Annapolis Ferry Dock","1076 Bay Street, Port Orchard 98366",
-  "Hwy. 305 & Suquamish Way","16003 WA-305, Poulsbo, Washington 98370",
-  "McWilliams Park & Ride","1601 NE McWilliams Rd, Bremerton, Washington 98311",
-  "Suquamish Park & Ride","18829 Division Ave NE, Suquamish, Washington 98392",
-  "Port Orchard Wal-Mart","3497 Bethel Rd SE, Port Orchard 98366",
-  "George's Corner Park & Ride","27618 Hansville Rd NE, Kingston 98346",
-  "Gateway Fellowship Park & Ride","18901 8th Ave NE, Poulsbo 98370",
-  "Miller Bay & Indianola","23404 Miller Bay Rd NE Poulsbo, Washington 98370",
-  "St. Gabriel's Church","1150 Mitchell Ave SE, Port Orchard 98366"
-)
-# geocode the addresses into geocode = c(lon, lat)
-transit_cp <- transit_cp %>% #tmaptools::geocode_OSM(transit_cp$ADDRESS)
-  rowwise() %>%
-  mutate(geocode = list(set_names(mapboxapi::mb_geocode(ADDRESS),
-                        c("lon","lat")))) %>%
-  unnest_wider(geocode)
-
-# convert the lat lon into a point geometry
-transits <- st_as_sf(transit_cp, 
-                      coords = c(x = "lon", y = "lat"), 
-                      crs = 4326) 
-transits2 <- getWalkingPolygonInCounty(transits, "NAME")
-the_transits <- unique(st_set_geometry(transits2, NULL)$id)
-loop_transits <- st_as_sf(transits2) %>% mutate(tomatch = id)
-
-# combine data into tibble for this report
-#These regions and pedestrian generators are:
-kds <- "Kitsap County GIS"
-kt <- "Kitsap Transit Website"
-wd <- "Walking distance polygons from Mapbox"
-breakdowns <- tribble(
-  ~Name,~pointdf,~polydf,~idcolpoly,~items,~data_source,
-  "Commissioner Districts",NA,"commdists","NAME","the_commdists",kds,
-  "UGAs",NA,"uga_no_city","GMA_JURISD","the_ugas",kds,
-  "LAMIRDs",NA,"lamird","PLAN_AREA","the_lamirds",kds,
-  "Public Schools","pub_schools","pub_schools2","id","the_schools",paste(kds,wd),
-  "Public Libraries","libraries","libraries2","id","the_libraries",paste(kds,wd),
-  "Transit Centers","transits","transits2","id","the_transits",paste(kt,wd),
-  )
-
-```
+We are hopeful that future investments will improve the quality of pedestrian facilities across the county, especially around community facilities, like schools and libraries, and in areas designated for urban service - in the Urban Growth Areas.
+</div>
 
@@
 
-```{r kitsapmap, cache = TRUE}
-
-#| fig.align = "center"
-#| fig.width = 12
-#| fig.height = 8
-#| out.width = '100%'
+```{r kitsapmap}
+#| fig.align = "center", 
+#| fig.width = 12,
+#| fig.height = 8, 
+#| out.width = '100%',
+#| dpi = 96
 
@@
 
-kitsap_all_roads <- kitsap +
-  geom_sf(data = clean_roadcl, aes(color = road_class)) + 
-  road_col +
-  theme_void() +
-  labs(title = "Roads in Kitsap County",
-       subtitle = paste0(nice_comma(sum(st_length(clean_roadcl))/5280),
-                         " centerline miles"),
-       #caption = "Data source: Kitsap GIS",
-       color = "")
-
-kitsap_roads_no_cities <- kitsap + 
-  geom_sf(data = cities["NAME"], 
-          aes(fill="NAME"),
-          fill="#808080",
-          alpha = 0.5,
-          show.legend = FALSE) +
-  geom_sf(data = ncdf, aes(color = road_class), show.legend = FALSE) + 
-  road_col +
-  theme_void() +
-  labs(title = "No Cities, Easements, & Highways",
-       subtitle = paste0(nice_comma(sum(st_length(ncdf))/5280),
-                         " centerline miles"),
-       caption = "Data source: Kitsap GIS")
-
-library(patchwork)
-
-kitsap_all_roads + kitsap_roads_no_cities +
-  plot_layout(guides = 'collect') & theme(legend.position = 'bottom')
-
-detach("package:patchwork", unload=TRUE)
+getKitsapRoadMap(kitsap, clean_roadcl, ncdf) 
 ```
@@
 ```{r}
-library(tidyverse)
-library(sf)
 
-groupcl <- ncdf %>%  
-  st_set_geometry(., NULL) %>% 
-  group_by(road_class) %>% 
-  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
-            segments = nice_comma(n(),0), 
-            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
-            .groups = "drop") 
-df <- ncdf %>% #clean_roadcl %>%  
-  st_set_geometry(., NULL) %>% 
-  group_by(road_class = "Total") %>% 
-  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
-            segments = nice_comma(n(),0), 
-            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
-            .groups = "drop") 
-groupcl <- bind_rows(groupcl, df)
-
-colnames(groupcl) <- c("Road Classification", 
-                       "Count of Roads", "Count of Segments", 
-                       "Length in Miles") 
-knitr::kable(groupcl, 
-             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")  %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+getKableRoadMiles(ncdf)
 ```
@@
 ```{r}
-df <- sidewalks_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-df2 <- sidewalks_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-
-df <- bind_rows(df, df2)
-
-knitr::kable(df, 
-             caption = "Miles of Sidewalks by Width of Sidewalk (single side)") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+getKableFacilityMiles(sidewalks_nc, "Sidewalk", "The minimum sidewalk in the Kitsap County Road Standards is 5 ft")
 ```
@@
 ```{r}
-df <- shoulders_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = paste(WIDTH, "ft")) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-df2 <- shoulders_nc %>%  
-  st_set_geometry(., NULL) %>%
-  mutate(`Width in Feet` = "Total", WIDTH = 100) %>%
-  group_by(WIDTH, `Width in Feet` ) %>% # road_type) %>%
-  summarise(miles = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
-            .groups = "drop") %>%
-  select(-WIDTH)
-
-df <- bind_rows(df, df2)
-
-knitr::kable(df, 
-             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+getKableFacilityMiles(shoulders_nc, "Shoulder","The minimum shoulder in the Kitsap County Road Standards is 3 ft")
 ```
@@
 In order to do this evaluation, we must combine the roads data (which has information on the class and setting of roads) with the sidewalk and shoulder data (with information about the pedestrian facilities).
-However, the most obvious methods of combination fail.
-
-#### Matching on Road Log Id is good but insufficient
-
-All of the data sets appear to use the same set of road log IDs from public works.
-Unfortunately, knowing that there is a shoulder or sidewalk somewhere on a road does not get to how much of that road has the sidewalk or shoulder.^[As roads sometimes change names and road log IDs, there are also some breaks between the road data and sidewalk/shoulder data.
-This was noted for Madrone Ave in Manchester - given the same road log ID as Main St when the road bends.]
-
-This diagnostic and ugly plot is intended to show the coverage of road log id across the sidewalk and shoulder data sets compared to the raod data.
-While the sidewalk and shoulder data are more sparse, the range is consistent.
-
-```{r}
-sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
-                  aes(x = as.numeric(row.names(nc_sw_sh)), y = RD_LOG_ID)) + 
-  geom_point() +
-  labs(title = "Sidewalk and Shoulder Data",
-       x = "") +
-  expand_limits(y = c(0,100000))
-road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
-                  aes(x = as.numeric(row.names(ncdf)), y = RD_LOG_ID)) + 
-  geom_point() +
-  labs(title = "Roads Data",
-       x = "")
-
-library(patchwork)
-
-sw_plot + road_plot
-
-detach("package:patchwork", unload=TRUE)
-rm(sw_plot)
-rm(road_plot)
-```
-
-```{r}
-match_road_id <- st_set_geometry(sidewalks,NULL) %>%
-  mutate(RD_LOG_ID = as.integer(ROADLOGID)) %>%
-  left_join(select(st_set_geometry(clean_roadcl,NULL),RD_LOG_ID,FULL_NAME)) %>%
-  group_by(RD_LOG_ID, ROADNAME, SIDE_OF_RO, WIDTH) %>%
-  summarise(n = sum(!is.na(FULL_NAME)), .groups = "drop") %>% 
-  filter(n == 0, WIDTH > 0)
-```
-
-The most obvious smaller match is the segment id. However, 
-**NOTE: We cannot actually use the SEGMENT ID in this draft report because it does not have the same meaning in the sidewalk and shoulder data as it has in the road data.
-The following plots are diagnostic but show that the sidewalks and shoulders data do not have segment ids covering the same spectrum as the road data.
-
-```{r}
-sw_plot <- ggplot(st_set_geometry(nc_sw_sh,NULL), 
-                  aes(x = as.numeric(row.names(nc_sw_sh)), y = SEGMENT_ID)) + 
-  geom_point() +
-  labs(title = "Sidewalk and Shoulder Data",
-       x = "")+
-  expand_limits(y = c(0,20000))
-road_plot <- ggplot(st_set_geometry(ncdf,NULL), 
-                  aes(x = as.numeric(row.names(ncdf)), y = SEGMENT_ID)) + 
-  geom_point() +
-  labs(title = "Roads Data",
-       x = "")
-
-library(patchwork)
-
-sw_plot + road_plot
-
-detach("package:patchwork", unload=TRUE)
-rm(sw_plot)
-rm(road_plot)
-```
-
-The segment ids are clearly not from the same space, but the road log ids are consistent.^[Mostly.] However, we cannot combine on road log id alone - because the roads change character and the type of pedestrian facilities over space.
-We combine instead on geometry.
-
-For now, we are making a space around a sidewalk or shoulder geometry, a buffer, to capture the nearby area where the road would be, and seeing if there are any road segments that intersect.
-If the road segment intersects and has the same road log id, we keep it.
-
-```{r combine-rd-sw-sh}
-# combine the sidewalk and shoulder and roads data
-x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 600)
-x$RD_LOG_ID <- 43809
-comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
-                                     SEGMENT_ID_R = SEGMENT_ID,
-                                     FULL_NAME_R = FULL_NAME, 
-                                     LENGTH_R = LENGTH,
-                                     road_class, RUCODE), 
-                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
-  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
-  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
-  mutate(distance = as.integer(st_length(.)))
-
-```
+Here, we join based on the geometries of the roads and sidewalks/shoulders.^[We intersect the road data with a buffered geometry of the sidewalk or shoulder and then remove any matching roads that do not have the same road log id.] Because segments are different for roads and sidewalk shoulder data, a single segment of road centerline may match several segments of sidewalk or shoulder data, and less commonly vice versa.
 
 ## Method of Evaluation
-
-Road segments are the basis for evaluation because they are the smallest unit in common across the data and can be grouped together to form trips.
-Actually matching up the data for roads, shoulders, and sidewalks will prove to be a more complicated task than it appears at first blush.
-As noted in the [Data section](#datasection), the Segment IDs in the shoulder and sidewalk data do not cover the range of segments as in the road data.
-Here, we join based on the geometries of the roads and sidewalks/shoulders.^[We intersect the road data with a buffered geometry of the sidewalk or shoulder and then remove any matching roads that do not have the same road log id.]
 
@@
 # Road Standards Tables 3.3 and 3.4
-library(tidyverse)
-library(kableExtra)
-road_standards_given <- tribble(
-  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, 
-  "Urban", "Local Road", NA, 5, NA,
-  "Urban", "Local Sub-collector", NA, 5, NA,
-  "Urban", "Arterial", NA, 6, NA,
-  "Urban", "Collector", NA, 6, NA,
-  "Rural", "Local Road", NA, NA, 3,
-  "Rural", "Local Sub-collector", NA, NA, 4,
-  "Rural", "Collector", "ADT 400-750", NA, 3,
-  "Rural", "Collector", "ADT 751-1000", NA, 4,
-  "Rural", "Collector", "DHV 100-200", NA, 6,
-  "Rural", "Collector", "DHV > 200", NA, 8,
-  "Rural", "Minor Arterial", "DHV < 100", NA, 4,
-  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,
-  "Rural", "Minor Arterial", "DHV > 200", NA, 8,
-  "Rural", "Principal Arterial", "DHV <200", NA, 6,
-  "Rural", "Principal Arterial", "DHV > 200", NA, 8,
-)
 
-knitr::kable(road_standards_given,
-             caption = "Summary of Tables 3.3 and 3.4 in Kitsap County Road Standards [@kc_road_standards]")  %>%
-   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT.
-Note: sub-collector, ADT, and DHV are not defined in the publicly available road data.") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+getKableRoadStandardsGiven()
 ```
@@
 Because of all of the items that are not available in the road data at this time, this draft makes some assumptions to allow for an evaluation.
-When the information is known, a more accurate evaluation can be made.^[Roads with unknown setting - RUCODE is 0, we match them to the Urban road standards.
+When the information is known, a more accurate evaluation can be made.^[Roads with unknown setting - RUCODE is 0 are evaluated to the Urban road standards.
 This may be incorrect.]
@@
 # Road Standards Tables 3.3 and 3.4 translated for data we have
-library(tidyverse)
-library(kableExtra)
-road_standards <- tribble(
-  ~setting, ~class, ~sidewalk, ~shoulder,
-  "Urban", "Local Access", 5, NA,
-  "Urban", "Minor Collector",6, NA,
-  "Urban", "Major Collector",6, NA,
-  "Urban", "Minor Arterial", 6, NA,
-  "Urban", "Principal Arterial", 6, NA,
-  "Rural", "Local Access", NA, 3,
-  "Rural", "Minor Collector",NA, 4,
-  "Rural", "Major Collector",NA, 6,
-  "Rural", "Minor Arterial", NA, 4,
-  "Rural", "Principal Arterial", NA, 6,
-)
-
-knitr::kable(road_standards,
-             caption = "Simplified Assumptions to Allow for Evaluation")  %>%
-   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+getKableRoadStandards()
 ```
@@
 - 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
-- 1 if the sidewalk or shoulder is present and is narrower than AASHTO (**NOTE to subcommittee - do we think we should use a 1 if there is a shoulder when there should be a sidewalk?**)
+- 2 if the standard is a sidewalk but a shoulder of adequate size (based on road classification and standards) is present
+- 1 if a sidewalk or shoulder is present and is narrower than AASHTO 
 - 0 if none present or no data
+
+Note that in the absence of ADT data, our interpretation of the AASHTO minimum is limited - we assume it is a 4 foot shoulder for this example.
+However, AASHTO minimums can be up to 8 feet under certain conditions.
 
@@
 ```{r}
-example_tbl <- tribble(
-  ~Segment, ~"Segment Score", ~"Segment Centerline Linear Feet",
-  "A", 6, 100,
-  "B", 0, 10,
-  "C",10,50,
-  "D",8,200,
-  "Total",7.5 ,360,
-)
-
-knitr::kable(example_tbl) %>%
-kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
-              full_width = F, 
-              position = "float_left")
+getKableExampleAggregateScore()
 ```
@@
 ```{r}
-score_interpretation <- tribble(
-  ~score_at_least, ~interpretation,
-  10, "Good",
-  5, "Adequate",
-  3, "Minimum",
-  1, "Poor",
-  0.001, "Very Poor",
-  0, "Nothing"
-)
 
-getScoreInterpretation <- function(x){
-  first(score_interpretation$interpretation[
-        score_interpretation$score_at_least <= x]
-    )
-}
+getScoreInterpretationKable()
+```
 
-df <- score_interpretation %>%
-  mutate(`Score At Least` = ifelse(score_at_least == 0.001, "> 0", 
-                                   nice_comma(score_at_least,0))) %>%
-  select(`Score At Least`, Interpretation = interpretation)
-
-knitr::kable(df,
-             caption = "Interpretation of Scores for Segments and for Areas")  %>%
-   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums") %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
-                full_width = F)
-```
+**NOTE TO SPECIAL COMMITTEE: a single segment can never have a score between 8 and 10.
+And an aggregation of an area could have a score of 9.9 and still show up as Adequate - it only gets a Good at a 10.
+Should we consider interpreting anything over an 8 as Good?**
 
@@
 
-```{r compare_standards}
-# take the road cl data for the roads under evaluation
-# remove geometry
-# merge with standards to get requirement based on setting (RUCODE) and class (road_class)
-# merge with sidewalk and shoulder data to get what exists
+The overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, `r getScoreInterpretation(comb_total_score)`.
 
-library(tidyverse)
-library(sf)
+There are road segments with ratings of Nothing (0) through Good (10).
 
-road_standards <- road_standards %>%
-  rename(standard_sidewalk = sidewalk, 
-         standard_shoulder = shoulder, 
-         road_class = class)
+A map of the county with roads colored based on their scores shows the distribution of pedestrian facilities. Urban Growth Areas are highlighted.
 
-getSideScore <- function(standard_sidewalk, standard_shoulder,
-                         sidewalk, shoulder) {
-  ifelse(is.na(standard_sidewalk), # dealing with shoulder
-         case_when(
-           !is.na(sidewalk) ~ 5, # there is a sidewalk present when a shoulder is the standard
-           shoulder >= standard_shoulder ~ 5,
-           shoulder < standard_shoulder & shoulder > 4 ~ 3,
-           shoulder < standard_shoulder & shoulder > 0 ~ 1,
-           TRUE ~ 0
-         ),
-         case_when(
-           sidewalk >= standard_sidewalk ~ 5,
-           sidewalk < standard_sidewalk & sidewalk > 4 ~ 3,
-           sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
-           TRUE ~ 0
-         )
-  )
-}
-
-comb_stand <- comb %>% 
-  st_set_geometry(., NULL) %>% 
-  mutate(setting = ifelse(RUCODE == 1, "Rural", 
-                          ifelse(RUCODE == 2, "Urban", "Unknown")),
-         ped_type = sub("\\:.*", "", TYPE),
-         nn = paste(sub("paved ","",tolower(ped_type)), tolower(SIDE_OF_RO),sep ="_")) %>%
-  pivot_wider(id_cols = c(-RD_LOG_ID:-TYPE,-ped_type), 
-              names_from = nn, 
-              values_from = WIDTH,
-              values_fn = mean) %>%
-  left_join(road_standards)
-
-comb_score <- comb_stand %>%
-  mutate(score_left = getSideScore(standard_sidewalk, standard_shoulder, 
-                                   sidewalk_left, shoulder_left),
-         score_right = getSideScore(standard_sidewalk, standard_shoulder, 
-                                   sidewalk_right, shoulder_right),
-         score = score_left + score_right)
+```{r map_score_overall}
+#| fig.align = "center", 
+#| fig.width = 10, 
+#| fig.height = 8, 
+#| out.width = '100%',
+#| dpi = 96
+#| 
+getMapScoreSummaryOverall(kitsap, uga, cities, comb_map)
 
@@
 
-comb_score2 <- st_set_geometry(ncdf, NULL) %>% 
-  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
-         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
-  mutate(setting = ifelse(RUCODE == 1, "Rural", 
-                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
-  left_join(comb_score) %>%
-  mutate(the_score = ifelse(is.na(score), 0, score),
-         the_length = ifelse(is.na(distance), LENGTH, distance),
-         w_score = the_score*the_length)
-
-comb_total_score <- sum(comb_score2$w_score)/sum(comb_score2$the_length)
-
-comb_score_by_setting <- comb_score2 %>%
-  group_by(setting) %>%
-  summarise(score = sum(w_score)/sum(the_length),
-            .groups = "drop") %>%
-  rowwise() %>%
-  mutate(interpretation = getScoreInterpretation(score))
-
-comb_score_by_setting_class <- comb_score2 %>%
-  group_by(setting, road_class) %>%
-  summarise(score = sum(w_score)/sum(the_length),
-            .groups = "drop") %>%
-  rowwise() %>%
-  mutate(interpretation = getScoreInterpretation(score))
+getKableScoreSummaryForMap(comb_map)
 ```
 
-The overall score for pedestrian facilities in Kitsap County is `r nice_comma(comb_total_score)`, `r getScoreInterpretation(comb_total_score)`.
-
-Looking by the same breakdown that we have for standards, we can see where the scores are highest and lowest.
+### By Setting and Road Class
+Looking by the same breakdown that we have for standards - by setting and road class, we can see where the scores are highest and lowest.
 
 ```{r tbl_combscorestandard}
-knitr::kable(comb_score_by_setting_class %>%
-               rename(Setting = setting,
-                      Class = road_class,
-                      Score = score,
-                      Interpretation = interpretation) %>%
-               mutate(Score = ifelse(Score == 0, "0", nice_comma(Score, 1))),
-             caption = "Weighted Scores by Standard Group")  %>%
-   footnote(general = "Unknown setting is scored as if it is Urban."
-                      ) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+getKableScoreSummaryForStandards(comb_score2)
 ```
 
-```{r map_score}
+### By Setting Alone
 
-comb_map <- st_as_sf(ncdf) %>% 
-  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
-         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
-  mutate(setting = ifelse(RUCODE == 1, "Rural", 
-                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>% 
-  left_join(comb_score) %>%
-  rowwise() %>%
-  mutate(interpretation = getScoreInterpretation(score))
+The character of an area contributes greatly to whether we would expect pedestrians. More pedestrians are expected in urban settings than rural.
 
-score_col <- scale_color_manual(values = c("Good"="#4d9221",
-                                           "Adequate"="#a1d76a",
-                                           "Minimum" = "#e6f5d0",
-                                           "Poor" = "#fde0ef",
-                                           "Very Poor" = "#e9a3c9",
-                                           "Nothing" = "#c51b7d"),
-                                na.value = "#808080")
-comb_map$interpretation <- factor(comb_map$interpretation, 
-                                  levels =  score_interpretation$interpretation)
-kitsap + 
-  geom_sf(data = cities["NAME"], 
-          aes(fill="NAME"),
-          fill="#808080",
-          alpha = 0.5,
-          show.legend = FALSE) +
-  geom_sf(data = ncdf, aes(color = 1), color = "#808080", show.legend = FALSE) + 
-  geom_sf(data = comb_map, aes(color = interpretation), show.legend = "line") + 
-  score_col +
-  theme_void() +
-  labs(title = "Calculated Pedestrian Facility Score",
-       caption = "Data source: Kitsap GIS, scores calculated",
-       color = "Score Interpretation")
-
+```{r tbl_combscoresetting}
+getKableScoreSummaryForSetting(comb_score2)
 ```
 
-```{r}
-# More detailed information
-# knitr::kable(
-#   comb_score2 %>%
-#   group_by(setting, road_class, score) %>% 
-#   summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
-#             segments = nice_comma(n()), 
-#             #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
-#             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
-#             .groups = "drop") 
-#   )
-```
+It may be useful to consider that for the unincorporated county, the roads that are designated as in an Urban setting are mostly inside of the Urban Growth Areas - these areas are identified to develop to an urban service standard, like a city.
+**With the data and methods used in this example report, these areas are not amenable to pedestrians - contrasting what we might desire in our urban environment.**
 
-```{r}
-ggplot(filter(comb_score2, setting != "Unknown"), 
-       aes(x = as.factor(score), y = the_length/5280))+
-  geom_col() +
-  labs(x = "score based on data available", 
-       y = "centerline miles of road") +
-  theme_minimal() +
-  facet_grid(setting ~ road_class)
-```
-
-If we remove the segments where we do not have information about sidewalks and shoulders at all, we can see that for those, we are generally finding higher scores for rural roads - with paved shoulders.
-
-```{r}
-ggplot(filter(comb_score2, setting != "Unknown", !is.na(score)), 
-       aes(x = as.factor(score), y = LENGTH/5280))+
-  geom_col() +
-  labs(x = "score based on data available", 
-       y = "centerline miles of road") +
-  theme_minimal() +
-  facet_grid(setting ~ road_class)
-```
-
-Some of the segments with score of 0 are because they have shoulders when they should have sidewalks.
-If you are familiar with Central Kitsap, you might recognize this kind of situation on Provost/Old Frontier - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.
-Here's a few example segments.
-
-```{r}
-
-roads_0 <- comb_score2 %>%
-  filter(score == 0) %>%
-  select(setting, road_class, FULL_NAME_R, sidewalk_right:standard_sidewalk)
-
-knitr::kable(head(roads_0) %>%
-               rename(Setting = setting,
-                      Class = road_class,
-                      Name = FULL_NAME_R,
-                      `sidewalk R` = sidewalk_right,
-                      `sidewalk L` = sidewalk_left,
-                      `shoulder R` = shoulder_right,
-                      `shoulder L` = shoulder_left,
-                      `standard sidewalk` = standard_sidewalk)) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-
-```
-
-Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road.
-A few details for a few roads where at least one segment has a 10 are shown here.^[This table is only showing the first few entries because it is a long table.] **Debbie - these would be the subset of segments that "meet the standard"**
-```{r}
-roads_10 <- comb_score2 %>%
-  filter(score == 10) %>%
-  group_by(setting, road_class, FULL_NAME_R) %>%
-  summarise(segments = nice_comma(n(),0), 
-            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
-            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
-            .groups = "drop") %>%
-  rename(Setting = setting,
-         Class = road_class,
-         Name = FULL_NAME_R,
-         `Segments w 10` = segments,
-         `Segment Length` = road_len,
-         `Ped Fac Length` = dist)
-
-knitr::kable(head(roads_10)) %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
-```
-
-Some road segments are missing entirely from the centerline road data - for example, the portion of Ridgetop that goes from the end of Myhre and curves back to Silverdale Way.
-These missing road segments on Ridgetop are mostly complete streets with wide sidewalks.
-Here, showing a map with all roads with the name *RIDGETOP*, you can only see the part going from Silverdale Way to Myhre.^[For completeness, we also checked the road log id, and it has the same data.] These segments, and any others like them, missing from the data decrease the overall pedestrian facilities score.
-
-```{r}
-kitsap +
-  geom_sf(data = filter(clean_roadcl, STREETNAME == "RIDGETOP"), aes(color = road_class)) + 
-  road_col +
-  theme_void() +
-  labs(title = "STREETNAME == RIDGETOP",
-       color = "")
-```
-
-## Smaller Breakdowns of the County
-
+## Smaller Regions
 We can also look at regions and pedestrian generators within the County.
@@
 Regions are already defined by an area.
-For pedestrian generators, we identify an area that is within 30 minute walking distance of the facility and evaluate roads therein.^[30 minute walking distance calculated using the mapboxapi function `mb_isochrone`.
+For pedestrian generators, we identify an area that is within a 15 minute walking distance of the facility and evaluate roads therein.^[A 15 minute walking distance is somewhere between 1/4 and 3/4 mile varying by walking pace.
+Walking distance calculated using the mapboxapi function `mb_isochrone`.
 Isochrones use map data from Mapbox and OpenStreetMap and their data sources.
@@
 
-Then, for each individual region, we show the pedestrian facilities on the left side and the right side of the road and calculate scores.
+Then, for each individual region, we show the evaluated pedestrian facilities and calculate scores.
 
@@
     function(irow) {
-      knitr::knit_expand(file = "child_region.Rmd",
+      knitr::knit_expand(file = "Child-Metrics-Region.Rmd",
                                    i = irow )
@@
 
-However, there are a couple of known data irregularities, such as:
-
- * missing part of Ridgetop Blvd in Silverdale where sidewalks are known to exist
- * missing data on sidewalks that are known to exist
- * unmatched road log identifiers between road and sidewalk and shoulder data
-
- And there are likely unknown data irregularities.
-
-In addition, many road segments have shoulders when their standards suggest they should have sidewalks.
-Is the presence of a shoulder, though not adequate, really the same as **nothing?** - that is how the scoring currently calculates.
-
-One cautionary note - while this method evaluates pedestrian facilities, it is entirely focused on the presence of facilities of adequate width and separation from cars.
+While this method evaluates pedestrian facilities, it is entirely focused on the presence of facilities of adequate width and separation from cars.
 It is not taking into consideration the quality of those facilities.
@@
 
-### Notes to the special committee drafting this
-We have a few to dos on this before we move forward.
+*Important reader note:  The authors of this example report are aware of errors (mostly of omission) in the sidewalk and shoulder data that we were able to obtain from the County.
+The County does not officially maintain these records, so we are not surprised that they are out of date.
+As far as we can tell, these errors skew the scoring downwards - actual pedestrian facilities should evaluate better than shown here.
+Discriminating readers will see obvious gaps, such as Silverdale Way with "Nothing" when it has sidewalks, and Ridgetop Blvd with "Nothing" when it has mostly complete streets.*
 
-1. Be sure that our criteria make sense to us
-2. Consider how the scores should be interpreted - does the interpretation fit?
-3. Consider if we are missing something critical
+**DO NOT USE THESE NUMBERS OR GRAPHS FOR ANY OFFICIAL PURPOSE**
 

### 2022-10-31 child_single_geo_in_region.Rmd
many updates not tracked

@@
 ```
+
+Overall, for {{current_geo}}, the pedestrian facilities are evaluated as so.
+
+```{r, eval=has_both}
+# problem, we probably want to color the geometry of the sw/sh segments, but how to get the roads
+# that don't have a match and set them at 0
+# this is the challenge now
+this_comb_map <- st_as_sf(this_roads) %>% 
+  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
+         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
+  mutate(setting = ifelse(RUCODE == 1, "Rural", 
+                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>% 
+  left_join(this_comb_score2) %>%
+  rowwise() %>%
+  mutate(interpretation = getScoreInterpretation(the_score))
+
+this_comb_map$interpretation <- factor(this_comb_map$interpretation, 
+                                  levels =  score_interpretation$interpretation)
+this_comb_map <-  this_comb_map %>%
+  group_by(interpretation) %>%  # ".x" is refers to the current group:
+  group_modify(~ st_union(.x) %>% as_tibble()) %>%
+  ungroup() %>%
+  st_as_sf()
+
+ggplot() +
+  geom_sf(data = this_geo, aes(fill = 1), alpha = 0.1, lwd = 0, show.legend = FALSE) +
+  #geom_sf(data = this_roads, aes(color = 1), color = "#808080", show.legend = FALSE) + 
+  geom_sf(data = this_comb_map, aes(color = interpretation), show.legend = "line") + 
+  score_col +
+  theme_void() +
+  labs(title = paste("Calculated Pedestrian Facility Score:","{{current_geo}}"),
+       caption = "Data source: Kitsap GIS, scores calculated",
+       color = "Score Interpretation")
+
+```
+
+```{r, eval=has_both}
+
+knitr::kable(
+  this_comb_map %>% 
+    mutate(LENGTH = as.integer(st_length(.))) %>%
+    st_set_geometry(., NULL) %>% 
+    group_by(interpretation) %>%
+    summarise(Miles = nice_comma(sum(LENGTH)/5280,1),
+              .groups = "drop") %>%
+    bind_rows(this_comb_map %>% 
+                mutate(LENGTH = as.integer(st_length(.))) %>%
+                st_set_geometry(., NULL) %>% 
+                mutate(interpretation = "Total") %>%
+                group_by(interpretation) %>%
+                summarise(Miles = nice_comma(sum(LENGTH)/5280,1),
+                          .groups = "drop")
+    ) %>%
+    rename(`Score Interpretation` = interpretation),
+    caption = "Miles here more than overall miles is indicative of multiple shoulder or sidewalk geometries matching the same portion of road centerline on the same side of the road.
+This kind of data error is expected given the state of the sidewalk and shoulder data."
+) %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
+```

### 2022-10-31 functions/getReportData.R

@@ -1,0 +1,280 @@
+# report functions to return main data
+
+library(tidyverse)
+library(sf)
+
+########
+# kitsap
+
+getDataCities <- function() {
+  sf::read_sf(paste0(this_dir,"data/cities"))
+}
+
+getMapKitsap <- function() {
+  outline <- sf::read_sf(paste0(this_dir,"data/outline"))
+  outline$plain = 1
+  ggplot() + 
+    geom_sf(data = outline["plain"], color = NA, alpha = 0.6) + 
+    theme_void()
+}
+
+getDataCountyNoCities <- function() {
+  cities <- getDataCities()
+  outline <- sf::read_sf(paste0(this_dir,"data/outline"))
+  st_difference(outline, st_union(cities))
+}
+
+
+########
+# handl road data
+
+getDataRoad <- function() {
+roadcl <- sf::read_sf("./data/roadcl")
+clean_roadcl <- roadcl %>%
+  filter(RD_LOG_ID != 0 & RD_LOG_ID != 200) %>% #not county roads
+  mutate(road_type = recode(HIERARCHY,
+                            `1` = "State HWY",
+                            `2` = "State HWY",
+                            `3` = "Collector/Arterial",
+                            `4` = "Local",
+                            `5` = "Easement",
+                            `6` = "Proposed",
+                            `7` = "Ramp",
+                            `8` = "Walking Path", # there should be 0
+                            .default = "Other"),
+         road_class = recode(FCLASS,
+                             `2` = "Freeway / Expressway",
+                             `3` = "Principal Arterial",
+                             `4` = "Minor Arterial",
+                             `5` = "Major Collector",
+                             `6` = "Minor Collector",
+                             `7` = "Local Access",
+                             `100` = "Rural Local Access (e.g. easement or other type)",
+                             `110` = "Urban Local Access (e.g. easement or other type)",
+                             .default = "Other")
+  )
+# FCLASS - Federal Functional Classifications as referenced in WSDOT document 
+# "Guidelines For Amending Functional Classification in Washington State" from October 2013
[…]

### 2022-10-31 functions/reportFunctions.R

@@ -1,0 +1,595 @@
+# functions for the document
+
+library(kableExtra)
+library(tidyverse)
+library(sf)
+
+nice_comma <- function(x, dec = 1){
+  format(round(as.numeric(x), dec), digits = NULL, nsmall=dec, big.mark=",") 
+}
+'%ni%' <- function(x,y)!('%in%'(x,y))
+
+
+###########################
+# number functions
+
+getAggregateScore <- function(the_df) {
+  # calculates with any df with w_score and the_length
+  
+  as.numeric(
+        sum(the_df$w_score, na.rm = TRUE)/
+          sum(the_df$the_length, na.rm=TRUE)
+        )
+}
+
+getFormattedAggregateScore <- function(the_df) {
+  # calculates with any df with w_score and the_length
+  
+  format(
+    round(
+      getAggregateScore(the_df),
+      1), 
+    digits = NULL, nsmall=1) 
+}
+
+getSideScore <- function(standard_sidewalk, standard_shoulder,
+                         sidewalk, shoulder) {
+  ifelse(is.na(standard_sidewalk), # dealing with shoulder
+         case_when(
+           !is.na(sidewalk) ~ 5, # sidewalk present when a shoulder is the standard
+           shoulder >= standard_shoulder ~ 5,
+           shoulder < standard_shoulder & shoulder >= 4 ~ 3, #aashto
+           shoulder < standard_shoulder & shoulder > 0 ~ 1,
+           TRUE ~ 0
+         ),
+         case_when(
+           sidewalk >= standard_sidewalk ~ 5,
+           sidewalk < standard_sidewalk & sidewalk > 4 ~ 3,
+           sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
+           shoulder >= standard_sidewalk ~ 2, # add for shoulder exist
+           shoulder < standard_sidewalk & shoulder > 0 ~ 1,
+           TRUE ~ 0
+         )
+  )
+}
+
+getScoreInterpretationTibble <- function(){
+  score_interpretation <- tribble(
+    ~score_at_least, ~interpretation,
+    10, "Good",
+    5, "Adequate",
+    3, "Minimum",
+    1, "Poor",
+    0.001, "Very Poor",
+    0, "Nothing"
+  )
+  score_interpretation
+}
+
+getRoadStandardsGivenTibble <- function(){
+  # these are standards from Kitsap Road Standards
+  road_standards_given <- tribble(
[…]
